### 2014-10-30 docs/ms.Rmd
Adding markdown file of the manuscript

---
output: pdf_document
---
##The transparency of colour reporting in plant and animal taxa

Daniel Noble^1^, Tom White^1^, Kate Umbers^1,2,5^, James O'Hanlon^1^, Rhiannon Dalrymple^3^, Daniel Zurek^4^

**Authors are currently not in order for publication

^1^ Department of Biological Sciences, Macquarie University, Sydney, NSW, Australia 
^2^ Department of Biological Sciences, University of Wollogong, Wollogong, NSW, Australia
^3^ School of BEEs, University of New South Wales, Sydney, NSW, Australia
^4^ Biological Sciences, University of Pennsylvania

^5^ Author to whom correspondance should be addressed:
Umbers?

### 2014-10-31 docs/ms.Rmd
Very rough outline of the manuscript. Lots of work that needs to be done. Refs need adding etc.

@@
 ---
-##The transparency of colour reporting in plant and animal taxa
+#The transparency of colour reporting in plant and animal taxa
 
-Daniel Noble^1^, Tom White^1^, Kate Umbers^1,2,5^, James O'Hanlon^1^, Rhiannon Dalrymple^3^, Daniel Zurek^4^
+Rhiannon L. Dalrymple^1,6^, Daniel W. A. Noble^2,6^, James C. O’Hanlon ^2,6^, Kate D. L. Umbers^3,4,6^, Thomas E. White^2,6^, Daniel B. Zurek^5,6^
 
-**Authors are currently not in order for publication
+^1^ Evolution & Ecology Research Centre, School of Biological, Earth and Environmental Sciences, University of New South Wales, Kensington, NSW, Australia, 2022 ^2^ Department of Biological Sciences, Macquarie University, North Ryde, NSW, Australia, 2109 ^3^ School of Biological Sciences, University of Wollongong, Wollongong, NSW, Australia 2252 ^4^ Centre for Evolutionary Biology, University of Western Australia, Crawley, WA, Australia 6008 ^5^ Department of Biological Sciences, University of Pittsburgh, Pittsburgh, PA, USA, 15260
 
-^1^ Department of Biological Sciences, Macquarie University, Sydney, NSW, Australia 
-^2^ Department of Biological Sciences, University of Wollogong, Wollogong, NSW, Australia
-^3^ School of BEEs, University of New South Wales, Sydney, NSW, Australia
-^4^ Biological Sciences, University of Pennsylvania
+^6^ Authors contributed equally and are presented in alphabetical order
 
-^5^ Author to whom correspondance should be addressed:
-Umbers?
+Corresponding Author:
+
+Keywords: #colsci, visual ecology, spectrophotometry, tetrahedral color space, quantum catch, pigment, structural color, thermoregulation, signaling, PAVO, colormetrics, visual modeling, melanism, sexual selection.
+
+##Abstract
+
+##Introduction
+Since Thayer (1909), Poulton (1890) and Wallace (YEAR) generations of biologists have endeavored to explain the mechanisms and functions of animal and plant coloration.
+For these biologists, identifying best practice in measuring color and color pattern and in determining how other species see colour has been a great challenge.
+Over the past 150 years biologists, both independently and in collaboration with physicists and chemists, have sought out and created methods that have enabled some remarkable discoveries (Lythgoe 1979 Shout out to Endler here?).
+These include, for example, the fundamental differences between human color vision and that of other animals (Bennett 1994) and the ubiquity of ultraviolet coloration in birds (Eaton & Lanyon 2003).
+
+  Today, the range of techniques used in research for measuring coloration are diverse and encompass spectrophotometry (also called spectroradiometry, spectroscopy ETC), photography, visual modeling.
+The application of spectrophotometry to questions in ecology and evolution revolutionized the field (Dyck 1966).
+Before the widespread use of spectrophotometry researchers assigned categorical color descriptions, arbitrary rankings or scores (e.g.
+Aldrich & James, 1991; Wheelwright & Janson, 1985) or matched paint swatches or munsell chip ratings based on their own visual assessment (e.g.
+Key and Day 1954, Burtt 1986).
+Although no doubt applied carefully, these techniques limited progress in field.
+For example, many researchers designed color assessment methods specifically for their projects (e.g.
+Alrich and James 1991) rendering data by different researchers incomparable.
+This was/is especially problematic when different measurement techniques are used on the same species (e,g, refXXX vs ref XXX).
+Also, human vision, blind to ultraviolet wavelengths, is inherently different to most other animals such as avian and reptilian tetrachromats and insect trichromats, (Bennett 1994, Douglas and Jeffrey 2014); human perception of color, and variation therein, is not only biologically irrelevant to most systems, but it can lead to erroneous conclusions (Bennett, Cuthill and Norris 1994, Endler 1990).
+
+Measuring reflectance spectra is now a reasonably inexpensive and convenient technique.
+Spectrophotometry allows with capture of quantitative and arguably repeatable data and has been rapidly taken up by researchers the field.
+However one drawback to spectrophotometry, is the ease with which erroneous data can be captured if attention is not paid to finer details of spectrum capture.
+For example, the angle from which spectra are taken (Santos & Lumeij, 2007), consistent calibration of light and dark standards, the distance between light source and object, the expectations of the researcher and the size, shape, glossiness or iridescence of the specimen (Andersson and Prager, 2006, Schaefer et al., 2008, Kemp, Reznick & Grether, 2008, Meadows et al 2011), can all contribute to potentially skewed data or alter results.
+Experience, guidance and objectivity are necessary to ensure data are consistent and representative of the subject and to ensure repeatability the conditions under which data were collected must be reported exactly.
+In conjunction with spectrophotometry, digital photography is increasingly being implemented in color quantification.
+High-resolution cameras are now inexpensive and portable, allow instantaneous sampling of multiple colour patches, and can produce highly repeatable images (McKay 2013).
+The main challenges for gathering and using color data from photographs are ensure correct standardization methods such as linearization and consistent lighting (Stevens et al.
+2007; ref Dariella Whitiyev 2014’s paper).
+Crucially, most digital cameras used in color science are consumer products designed for human visual color space and hardware modifications for ultraviolet and infrared imaging are necessary (e.g.
+Chiao et al.
+2009, McKay 2013), making photography complementary, rather than superior to spectrophotometry..
+
+  Expansion in the availability of objective methods for the measurement of color has been matched by advances in theory and analysis.
+In particular, the development of sensory models has produced significant insights into the way in which colorful stimuli are perceived (Bennett et al 1994, Chittka 1992, Endler & Mielke 2005, Vorobyev & Osorio 1998).
+Where the receptor sensitivities of an animal’s visual system are known, the physiological response of that animal’s visual system when viewing a given color can be predicted.
+Physiological sensory models of animal vision have been developed to model the perception of bees (e.g.
+Chittka, 1992), flies (e.g.
+Troje, 1993), birds (e.g.
+Endler & Mielke, 2005), moths (e.g.
+Chuang, Yang & Tso, 2007), frogs (REF), crabs (REF) and snakes (e.g.
+Maan & Cummings, 2012).
+
+Sensory models typically attempt to describe the subjective perception of color or brightness information as a function of an object’s reflectance, the ambient illumination, and a receiver’s sensory system (eg.
+Endler & Mielke 2005, Vorobyev et al.
+1998).
+Furthermore, an animal’s ability to distinguish between reflectance spectra can then be predicted based on the differential responses of a visual system to varying reflectance spectra.
+This approach enables researchers to move beyond purely quantitative comparisons of reflectance spectra and adopt biologically relevant perspectives of signal receivers to test/formulate hypotheses.
+
+Although relatively easy to implement and potentially powerful, sensory models are built on multiple assumptions about the way in which stimuli are visually and cognitively processed that can dramatically shape the results of a given analysis (Lind & Kelber 2009; Pike 2012).
+For example, Vorobyev et al.’s (1998; 2001) widely used receptor-noise model requires researchers to specify the form of receptor quantum catches (i.e.
+raw versus log-transformed), the existence of any color correction mechanism (e.g.
+von Kries), as well as the photoreceptor signal-to-noise ratios, their relative densities, and the type of noise predominating (i.e.
+neural and/or receptor noise).
+These data are only available for a small number of model species (e.g.
+honeybees) and thus assumptions and extrapolations are necessary when models are applied to related species.
+Unless the biological rationale underlying every parameter choice is clearly understood, justified, and reported, such analyses may become ineffective, especially when comparing results between studies (Pike 2012).
+
+  Since the repeatability and transparency of science is paramount, it is imperative that we regularly and constructively reflect on our own practices (Peng, 2009; Alsheikh-Ali et al.
+2011; Sandve et al YEAR).
+The potential for detrimental subjectivity in color measurement and the mathematical opacity of visual models for the average colour scienctist means that, as in any field, is it critical that methods are reported rigorously.
+Methods reporting should ideally have well defined standards such that sufficient detail is provided to allow comprehensive peer review, precise replication, cross-study comparison and data sharing (REF data sharing papers, Ask Rob Lanfear).
+The impetus for this paper was borne out of six months of participation in a journal club that focused on publications on coloration in biological systems.
+Based on our group discussions and on our personal experiences gathering and reporting color data, we were compelled to look into the literature to quantify  variation in how methods are reported.
+A preliminary investigation revealed sufficient variation among papers for us question the field’s consistency and completeness.
+Therefore, the aims of this paper are to (a) review methods reporting in papers published on colour in 2013 from 20 top tier journals and (b) provide a set of recommendations for best-practice in reporting on the most common methods.
+Our best hope is that this instructional provides a useful set of guidelines to improve the reproducibility and transparency of studies on color.
+
+##Ideal reporting in colour studies
+
+We developed a list of metrics we deemed critical to be reported for reproducibility of color methods and analysis loosely based on Andersson and Prager (2006) and Montgomerie (2006) (Table X).
+These included quantitative methods including spectrophotometric and photographic techniques (as in REF), but not qualitative methods such as the use of Munsell chips or paint swatches (as in REF).
+For visual modeling techniques we characterized a set of criteria required for reproducing the analysis.
+
+##What is reporting like in the literature? (what do people do?)
+
+To determine the current state of methods reporting in the field we searched for papers from 2013 in 20 journals (see Table X): We used the Boolean search terms colour*, color* or spectra* and searched for these in the title or abstract from each of the journal’s home pages.
+Our search returned a total of X papers in X journals.
+To review the large number of papers returned from the searches, journals were haphazardly divided up between the authors.
+Upon reading abstracts we determined which studies took empirical measures of colour (via photography or spectrophotometry) and discarded those that did not.
+In order to reduce the risk of observer bias in our scoring.
+Papers were then read and reassessed by a second author and a third, three times in total.
+Any discrepancies between the assessor scores and the inclusion of papers with more obscure or uncommon methodologies, were discussed and agreed upon.
+We then tabulated what details of the methods were reported against our criteria.

### 2014-10-31 docs/ms.Rmd
Changed line endings to have two spaces to start a new paragraph, rather than skipping a line

@@
 These include, for example, the fundamental differences between human color vision and that of other animals (Bennett 1994) and the ubiquity of ultraviolet coloration in birds (Eaton & Lanyon 2003).
-
 Today, the range of techniques used in research for measuring coloration are diverse and encompass spectrophotometry (also called spectroradiometry, spectroscopy ETC), photography, visual modeling.
@@
 Also, human vision, blind to ultraviolet wavelengths, is inherently different to most other animals such as avian and reptilian tetrachromats and insect trichromats, (Bennett 1994, Douglas and Jeffrey 2014); human perception of color, and variation therein, is not only biologically irrelevant to most systems, but it can lead to erroneous conclusions (Bennett, Cuthill and Norris 1994, Endler 1990).
-
 Measuring reflectance spectra is now a reasonably inexpensive and convenient technique.
@@
 2009, McKay 2013), making photography complementary, rather than superior to spectrophotometry..
-
 Expansion in the availability of objective methods for the measurement of color has been matched by advances in theory and analysis.
@@
 Maan & Cummings, 2012).
-
 Sensory models typically attempt to describe the subjective perception of color or brightness information as a function of an object’s reflectance, the ambient illumination, and a receiver’s sensory system (eg.
@@
 This approach enables researchers to move beyond purely quantitative comparisons of reflectance spectra and adopt biologically relevant perspectives of signal receivers to test/formulate hypotheses.
-
 Although relatively easy to implement and potentially powerful, sensory models are built on multiple assumptions about the way in which stimuli are visually and cognitively processed that can dramatically shape the results of a given analysis (Lind & Kelber 2009; Pike 2012).
@@
 Unless the biological rationale underlying every parameter choice is clearly understood, justified, and reported, such analyses may become ineffective, especially when comparing results between studies (Pike 2012).
-
 Since the repeatability and transparency of science is paramount, it is imperative that we regularly and constructively reflect on our own practices (Peng, 2009; Alsheikh-Ali et al.

### 2014-11-02 docs/ms.Rmd
Just messing around with r chunks and inline code within asbtract.
Changes the R markdown file

--- 2014-10-31 docs/ms.Rmd
@@
 ##Abstract
+```{r, eval = TRUE, echo = FALSE}
+x <- rnorm(200, 0, 1)
+x
+```
+
+`r round(mean(x), digits = 2)`
 
@@
 These include, for example, the fundamental differences between human color vision and that of other animals (Bennett 1994) and the ubiquity of ultraviolet coloration in birds (Eaton & Lanyon 2003).
+Today, the range of techniques used in research for measuring coloration are diverse and encompass spectrophotometry (also called spectroradiometry, spectroscopy ETC), photography, visual modeling.
 
-  Today, the range of techniques used in research for measuring coloration are diverse and encompass spectrophotometry (also called spectroradiometry, spectroscopy ETC), photography, visual modeling.
 The application of spectrophotometry to questions in ecology and evolution revolutionized the field (Dyck 1966).
@@
 Chiao et al.
-2009, McKay 2013), making photography complementary, rather than superior to spectrophotometry..
+2009, McKay 2013), making photography complementary, rather than superior to spectrophotometry.
 
@@
 Methods reporting should ideally have well defined standards such that sufficient detail is provided to allow comprehensive peer review, precise replication, cross-study comparison and data sharing (REF data sharing papers, Ask Rob Lanfear).
-The impetus for this paper was borne out of six months of participation in a journal club that focused on publications on coloration in biological systems.
-Based on our group discussions and on our personal experiences gathering and reporting color data, we were compelled to look into the literature to quantify  variation in how methods are reported.
-A preliminary investigation revealed sufficient variation among papers for us question the field’s consistency and completeness.
-Therefore, the aims of this paper are to (a) review methods reporting in papers published on colour in 2013 from 20 top tier journals and (b) provide a set of recommendations for best-practice in reporting on the most common methods.
+The aims of this paper are to (a) review methods reporting in papers published on colour in 2013 from 20 top tier journals and (b) provide a set of recommendations for best-practice in reporting on the most common methods.
 Our best hope is that this instructional provides a useful set of guidelines to improve the reproducibility and transparency of studies on color.

### 2014-11-02 R/analysis.r
Added the analysis.r file. This was old and must be a conflict with Tom.

@@ -1,0 +1,28 @@
+#-------------------------------------------------------------------------------------------------#
+# Title: Quality of colour reporting 
+# Authors: Daniel Noble, Tom White, Kate Umbers, James O'Hanlon, Rhiannon Dalrymple, Daniel Zurek
+# Description: Analysis of reporting throughout a series of "colour" journals. Mainly summary 
+#			   stats and graphs of the data. 
+#-------------------------------------------------------------------------------------------------#
+
+# Clear workspace
+rm(list=ls())
+
+## Packages. Coded to simplify multiple package installation and loading.
+packages <- c("car", "ggplot2")
+install.packages(packages)
+sapply(packages, function(x) library(x, character.only = TRUE, verbose = FALSE))
+
+## Set working dir. path.expand gives a bit of flexibility
+setwd(path.expand("~/Dropbox/Share\ Folder/colsci_rep/"))
+
+## Load data 
+coldat <- read.csv("data/col_data.csv", header = TRUE)
+str(coldat)
+
+## Process data. Remove irrelevant variables for analaysis. 
+# Remove justification columns, which are just simply for our use and complicate analysis
+col_just   <- colnames(coldat)[c(grep("_just", colnames(coldat)))]
+coldat     <- coldat[,! colnames(coldat) %in% col_just][-c(35:36)] # Additioally remove details on analysis.
+
+## Things we need to check. The studies categorized as "multiple" Need to make clear what the values (0 or 1) correspond to camera or specs. 

### 2014-11-02 docs/ms.Rmd

@@
 ##Abstract
-```{r, eval = TRUE, echo = FALSE}
-x <- rnorm(200, 0, 1)
-x
-```
-
-`r round(mean(x), digits = 2)`
 
@@
 For these biologists, identifying best practice in measuring color and color pattern and in determining how other species see colour has been a great challenge.
-Over the past 150 years biologists, both independently and in collaboration with physicists and chemists, have sought out and created methods that have enabled some remarkable discoveries (Lythgoe 1979 Shout out to Endler here?).
-These include, for example, the fundamental differences between human color vision and that of other animals (Bennett 1994) and the ubiquity of ultraviolet coloration in birds (Eaton & Lanyon 2003).
-Today, the range of techniques used in research for measuring coloration are diverse and encompass spectrophotometry (also called spectroradiometry, spectroscopy ETC), photography, visual modeling.
+Over the past 150 years biologists, both independently and in collaboration with physicists and chemists, have developed a range of techniques used to measure and analyse animal and plant coloration including spectrophotometry (also called spectroradiometry, spectroscopy ETC), photography and visual modeling.
 
@@
 Measuring reflectance spectra is now a reasonably inexpensive and convenient technique.
-Spectrophotometry allows with capture of quantitative and arguably repeatable data and has been rapidly taken up by researchers the field.
+Spectrophotometry allows the capture of quantitative and arguably repeatable data and has been rapidly taken up by researchers the field.
 However one drawback to spectrophotometry, is the ease with which erroneous data can be captured if attention is not paid to finer details of spectrum capture.
 For example, the angle from which spectra are taken (Santos & Lumeij, 2007), consistent calibration of light and dark standards, the distance between light source and object, the expectations of the researcher and the size, shape, glossiness or iridescence of the specimen (Andersson and Prager, 2006, Schaefer et al., 2008, Kemp, Reznick & Grether, 2008, Meadows et al 2011), can all contribute to potentially skewed data or alter results.
-Experience, guidance and objectivity are necessary to ensure data are consistent and representative of the subject and to ensure repeatability the conditions under which data were collected must be reported exactly.
+Experience, guidance and objectivity are necessary to ensure data are consistent and representative of the subject and to ensure repeatability, conditions under which data were collected must be reported exactly.
+
 In conjunction with spectrophotometry, digital photography is increasingly being implemented in color quantification.
@@
 
+###Table 1
+```{r, eval = TRUE, echo = FALSE}
+tab1 <- read.csv("/Users/danielnoble/Dropbox/6_Share_Folder/colsci_rep/output/tables/table1.csv", header = TRUE)
+table1 <- xtable(tab1)
+print(table1, type = "pdf")
+```
 ##What is reporting like in the literature? (what do people do?)
 
-To determine the current state of methods reporting in the field we searched for papers from 2013 in 20 journals (see Table X): We used the Boolean search terms colour*, color* or spectra* and searched for these in the title or abstract from each of the journal’s home pages.
-Our search returned a total of X papers in X journals.
+To determine the current state of methods reporting in the field we searched for papers from 2013 in 20 journals (see Table X): We used the Boolean search terms colour*, color* or spectra* and searched for these in the title or abstract from each of the journal home pages.
+Our search returned a total of X papers.
 To review the large number of papers returned from the searches, journals were haphazardly divided up between the authors.

### 2014-11-03 R/analysis.r
Very minor changes. Just thinking about the data more, but now need to run to meeting so didn't get far

@@ -23,6 +23,6 @@ str(coldat)
 ## Process data. Remove irrelevant variables for analaysis. 
 # Remove justification columns, which are just simply for our use and complicate analysis
 col_just   <- colnames(coldat)[c(grep("_just", colnames(coldat)))]
-coldat     <- coldat[,! colnames(coldat) %in% col_just][-c(35:36)] # Additioally remove details on analysis.
+coldat     <- coldat[,! colnames(coldat) %in% col_just][-c(2:13, 35:36)] # Additioally remove details on analysis and details on papers, but keep paper number.
 
 ## Things we need to check. The studies categorized as "multiple" Need to make clear what the values (0 or 1) correspond to camera or specs. 

### 2014-11-03 R/knit_doc.r
Added knitR script to knit ms

@@ -1,0 +1,6 @@
+
+#Read necessary functions and files
+rm(list=ls())
+setwd("./docs")
+knit2pdf("ms.Rmd")
+setwd("..")

### 2014-11-03 R/analysis.r
Minor changes to doc and waiting for KU to get back to me about analysis type col in data

@@ -21,7 +21,7 @@ coldat <- read.csv("data/col_data.csv", header = TRUE)
 str(coldat)
 
 ## Process data. Remove irrelevant variables for analaysis. 
-# Remove justification columns, which are just simply for our use and complicate analysis
+## Remove justification columns, which are just simply for our use and complicate analysis
 col_just   <- colnames(coldat)[c(grep("_just", colnames(coldat)))]
 coldat     <- coldat[,! colnames(coldat) %in% col_just][-c(2:13, 35:36)] # Additioally remove details on analysis and details on papers, but keep paper number.
 

### 2014-11-03 docs/ms.Rmd
Removed table 1 code, well blocked out at least, as it was giving errors when knitting. Added in journal names in italics

@@
 ```{r, eval = TRUE, echo = FALSE}
-tab1 <- read.csv("/Users/danielnoble/Dropbox/6_Share_Folder/colsci_rep/output/tables/table1.csv", header = TRUE)
-table1 <- xtable(tab1)
-print(table1, type = "pdf")
+#tab1 <- read.csv("/Users/danielnoble/Dropbox/6_Share_Folder/colsci_rep/output/tables/table1.csv", header = TRUE)
+#table1 <- xtable(tab1)
+#print(table1, type = "pdf")
 ```
@@
 
-To determine the current state of methods reporting in the field we searched for papers from 2013 in 20 journals (see Table X): We used the Boolean search terms colour*, color* or spectra* and searched for these in the title or abstract from each of the journal home pages.
+To determine the current state of methods reporting in the field we searched for papers from 2013 in 20 journals (**Animal Behaviour**,**Behavioral Ecology**,**Journal of Experimental Biology**,**BIology Letters**, **Proceedings of the Royal Society B: Biological Sciences**,**Biological Journal of the Linnean Society**,**Ethology**, **PLoS ONE**, ): We used the Boolean search terms colour*, color* or spectra* and searched for these in the title or abstract from each of the journal home pages.
 Our search returned a total of X papers.
@@
 In order to reduce the risk of observer bias in our scoring.
-Papers were then read and reassessed by a second author and a third, three times in total.
+Papers were then read and reassessed by two additional authors (three times in total).
 Any discrepancies between the assessor scores and the inclusion of papers with more obscure or uncommon methodologies, were discussed and agreed upon.

### 2014-11-03 docs/ms.Rmd
Opps forgot to save

@@
 
-To determine the current state of methods reporting in the field we searched for papers from 2013 in 20 journals (**Animal Behaviour**,**Behavioral Ecology**,**Journal of Experimental Biology**,**BIology Letters**, **Proceedings of the Royal Society B: Biological Sciences**,**Biological Journal of the Linnean Society**,**Ethology**, **PLoS ONE**, ): We used the Boolean search terms colour*, color* or spectra* and searched for these in the title or abstract from each of the journal home pages.
+To determine the current state of methods reporting in the field we searched for papers from 2013 in 20 journals (_Animal Behaviour_, _Behavioral Ecology_, _Journal of Experimental Biology_, _Biology Letters_, _Proceedings of the Royal Society B: Biological Sciences_, _Biological Journal of the Linnean Society_, _Ethology_, _The American Naturalist_, _Ecology Letters_, _PLoS ONE_, ): We used the Boolean search terms colour*, color* or spectra* and searched for these in the title or abstract from each of the journal home pages.
 Our search returned a total of X papers.

### 2014-11-04 R/analysis.r
Updated the analysis script and created a knitR script so that it can be used with sublime. knit the ms.md file first then use pandoc ms.md -o ms.pdf to convert to the desired output file

@@ -26,3 +26,45 @@ col_just   <- colnames(coldat)[c(grep("_just", colnames(coldat)))]
 coldat     <- coldat[,! colnames(coldat) %in% col_just][-c(2:13, 35:36)] # Additioally remove details on analysis and details on papers, but keep paper number.
 
 ## Things we need to check. The studies categorized as "multiple" Need to make clear what the values (0 or 1) correspond to camera or specs. 
+
+#--------------------------------------------- ANALYSIS -----------------------------------------#
+## How many studies utilized specs (1), cameras (2), both (3)?
+studytype <- table(coldat$stud_type)
+
+specs <- as.numeric(studytype[1])
+cam   <- as.numeric(studytype[2])
+both  <- as.numeric(studytype[3])
+
+## Of the studies using a spec, what proportion report the make and model?
+specModel <- (sum(coldat$spec_rep, na.rm = TRUE)/(specs+both))*100
+
+## Of the studies using a spec, what proportion report the make and model?
+camModel  <- (sum(coldat$cam_model, na.rm = TRUE)/(cam+both))*100
+
+## Of the studies using a spec, what proportion report the light source. Note that column needs to be extracted because it combines cameras and specs
+
+DatSpec   <- subset(coldat, stud_type == 1 |stud_type == 3)
+specLight <- (sum(DatSpec$light_source)/nrow(DatSpec))*100
+
+## Of the studies using a cam, what proportion report the light source. Note that column needs to be extracted because it combines cameras and specs
+
+DatCam    <- subset(coldat, stud_type == 2 |stud_type == 3)
+CamLight  <- (sum(DatCam$light_source, na.rm = TRUE)/nrow(DatCam))*100
+
+## Of Studies spec'ing what proportion report a dark standard
+drk_std   <- (sum(DatSpec$dark_std)/(nrow(DatSpec)))*100
+#white_std <- sum(DatSpec$white_std)/(nrow(DatSpec))
+
+## What proportion of studies specing report the num of specs averaged
+
+specAVG <- (sum(DatSpec$specpix_avg)/nrow(DatSpec))*100
+
+## What proportion of studies using a camera report the num of pixs averaged
+
+camAVG <- (sum(DatCam$specpix_avg)/nrow(DatCam))*100
+
+## What proportion of spec studies reported 1) the integration time, spec angle and distance?
+
[…]

### 2014-11-04 R/knit_doc.r

@@ -1,6 +1,11 @@
 
-#Read necessary functions and files
+packages <- c("knitr", "markdown")
+install.packages(packages)
+sapply(packages, function(x) library(x, character.only = TRUE))
+
+#Read necessary functions and files. There are two functions knit2pdf and knit2html which will produce documents directly, but there are problems with pdf and also problems rendering the markdown file when complete. Best to just use knit first then take md file and convert in ternimal. 
+
 rm(list=ls())
 setwd("./docs")
-knit2pdf("ms.Rmd")
+knit("ms.Rmd")
 setwd("..")

### 2014-11-04 docs/ms.Rmd
Began writing the results section and also spent most of day solving problems in Rmd coversion with pndoc.

--- 2014-11-02 docs/ms.Rmd
@@
 ---
-output: pdf_document
+fontsize: 12pt
+geometry: margin=0.75in
+output: html_document
 ---
+
 #The transparency of colour reporting in plant and animal taxa
@@
 
-We developed a list of metrics we deemed critical to be reported for reproducibility of color methods and analysis loosely based on Andersson and Prager (2006) and Montgomerie (2006) (Table X).
+We developed a list of metrics we deemed critical to be reported for reproducibility of color methods and analysis loosely based on Andersson and Prager (2006) and Montgomerie (2006) (Table 1 and 2).
 These included quantitative methods including spectrophotometric and photographic techniques (as in REF), but not qualitative methods such as the use of Munsell chips or paint swatches (as in REF).
-For visual modeling techniques we characterized a set of criteria required for reproducing the analysis.
+For visual modeling techniques we characterized a set of criteria required for reproducing the analysis (Table 2).
 
-###Table 1
-```{r, eval = TRUE, echo = FALSE}
-tab1 <- read.csv("/Users/danielnoble/Dropbox/6_Share_Folder/colsci_rep/output/tables/table1.csv", header = TRUE)
-table1 <- xtable(tab1)
-print(table1, type = "pdf")
-```
-##What is reporting like in the literature? (what do people do?)
+##What is reporting like in the literature?
 
-To determine the current state of methods reporting in the field we searched for papers from 2013 in 20 journals (see Table X): We used the Boolean search terms colour*, color* or spectra* and searched for these in the title or abstract from each of the journal home pages.
-Our search returned a total of X papers.
+To determine the current state of methods reporting in the field we searched for papers from 2013 in 23 top color science journals (_Animal Behaviour_, _Behavioral Ecology_, _Journal of Experimental Biology_, _Biology Letters_, _Proceedings of the Royal Society B: Biological Sciences_, _Biological Journal of the Linnean Society_, _Ethology_, _The American Naturalist_, _Ecology Letters_, _Journal of Evolutionary Biology_, _American Journal of Botany_, _Evology and Evolution_, _Oikos_, _Ecology_, _Journal of Ecology_, _Behavioral Ecology and Sociobiology_, _Current Zoology_, _Naturwissenschaften_, _New Phytologist_, _PLoS ONE_, _Evolution_, _Functional Ecology_, _Ethology_).
+For each of the journals, we used the Boolean search terms 'colour*', 'color*' or 'spectra*' and searched for these in the title or abstract from each of the journal home pages.
+We read the titles and abstracts of each paper and included only those papers that explicitly used either a spectrophotometer or camera to quantify colour signals.
+We excluded papers that were: 1) review papers; 2) methodological papers; 3) papers quantifying colour pattern; and 4) papers taking microspectrophotometric measurments of retinal absorbance.
 To review the large number of papers returned from the searches, journals were haphazardly divided up between the authors.
-Upon reading abstracts we determined which studies took empirical measures of colour (via photography or spectrophotometry) and discarded those that did not.
-In order to reduce the risk of observer bias in our scoring.
-Papers were then read and reassessed by a second author and a third, three times in total.
-Any discrepancies between the assessor scores and the inclusion of papers with more obscure or uncommon methodologies, were discussed and agreed upon.
+In order to reduce the risk of observer bias in our scoring, papers were read and reassessed by a total of three authors.
+Any discrepancies between the assessor scores were discussed and agreed upon prior to analysis.
 We then tabulated what details of the methods were reported against our criteria.
+
+Our search returned a total of 216 papers, however, only `r nrow(coldat)` met our criteria and were included in the final analysis.
+Overall, `r specs+both` studies used a spectrophotometer to measure color, while `r cam+both` used a camera.
+Of these studies, `r both` quantified color using both a spectrophotometer and a camera.
+Reporting of the make and model of the spectrophotometer and camera was high (Figure 1 - Specs = `r round(specModel, digits = 2)`% and Cameras = `r round(camModel, digits = 2)`%), however, only `r round(CamLight, digits = 2)`%  and `r round(specLight, digits = 2)`% of camera and spec studies reported a light source.
+In addition, the number reflectance spectra and pixels averaged were only reported in `r round(specAVG, digits = 2)`% and `r round(camAVG, digits = 2)`% of the studies, respectively.
+Surprisingly, of the spec studies, dark standards were only reported in `r round(drk_std, digits = 2)`% of the studies, while white standards were reported in %.
+Studies quantifying spectral reflectance also often did not report the integration time of the spectrophotmeter (`r round(intTime, digits = 2)`%), the angle (`r round(angle, digits = 2)`%) or the distance (`r `%) the probe was from the color patch.

### 2014-11-04 R/analysis.r
Updated analysis file with some more analysis

@@ -44,6 +44,8 @@ camModel  <- (sum(coldat$cam_model, na.rm = TRUE)/(cam+both))*100
 ## Of the studies using a spec, what proportion report the light source. Note that column needs to be extracted because it combines cameras and specs
 
 DatSpec   <- subset(coldat, stud_type == 1 |stud_type == 3)
+
+# Note that we need to change the rows where spec and cemera differ
 specLight <- (sum(DatSpec$light_source)/nrow(DatSpec))*100
 
 ## Of the studies using a cam, what proportion report the light source. Note that column needs to be extracted because it combines cameras and specs
@@ -53,7 +55,7 @@ CamLight  <- (sum(DatCam$light_source, na.rm = TRUE)/nrow(DatCam))*100
 
 ## Of Studies spec'ing what proportion report a dark standard
 drk_std   <- (sum(DatSpec$dark_std)/(nrow(DatSpec)))*100
-#white_std <- sum(DatSpec$white_std)/(nrow(DatSpec))
+white_stdspec <- (sum(DatSpec$white_stdspec)/nrow(DatSpec))*100
 
 ## What proportion of studies specing report the num of specs averaged
 
@@ -62,6 +64,13 @@ specAVG <- (sum(DatSpec$specpix_avg)/nrow(DatSpec))*100
 ## What proportion of studies using a camera report the num of pixs averaged
 
 camAVG <- (sum(DatCam$specpix_avg)/nrow(DatCam))*100
+
+## How do studies analyse colour?
+analysisType <- table(coldat$analysis_type)
+
+## How about just spec studies because they can do visual modelling, or colorimetric
+
+analysisTypeSpec <- table(DatSpec$analysis_type)
 
 ## What proportion of spec studies reported 1) the integration time, spec angle and distance?
 

### 2014-11-04 docs/ms.Rmd
Wrote more on analysis section

@@
 geometry: margin=0.75in
-output: html_document
+output: pdf_document
 ---
@@
 Reporting of the make and model of the spectrophotometer and camera was high (Figure 1 - Specs = `r round(specModel, digits = 2)`% and Cameras = `r round(camModel, digits = 2)`%), however, only `r round(CamLight, digits = 2)`%  and `r round(specLight, digits = 2)`% of camera and spec studies reported a light source.
-In addition, the number reflectance spectra and pixels averaged were only reported in `r round(specAVG, digits = 2)`% and `r round(camAVG, digits = 2)`% of the studies, respectively.
-Surprisingly, of the spec studies, dark standards were only reported in `r round(drk_std, digits = 2)`% of the studies, while white standards were reported in %.
-Studies quantifying spectral reflectance also often did not report the integration time of the spectrophotmeter (`r round(intTime, digits = 2)`%), the angle (`r round(angle, digits = 2)`%) or the distance (`r `%) the probe was from the color patch.
+In addition, the number of reflectance spectra and pixels averaged were only reported in `r round(specAVG, digits = 2)`% and `r round(camAVG, digits = 2)`% of the studies, respectively.
+Surprisingly, of the spec studies, dark standards were only reported in `r round(drk_std, digits = 2)`% of the studies, while white standards were reported in `r round(white_stdspec, digits = 2)`%.
+Studies quantifying spectral reflectance also often did not report the integration time of the spectrophotmeter (`r round(intTime, digits = 2)`%) or the angle (`r round(angle, digits = 2)`%) or the distance (`r round(dist, digits = 2)`%) the probe was from the color patch.
+
+Analysis varied across studies but, overall `r analysisType[1]` analyzed their data by generating colormetric (brightness, hue, chroma) variables whereas `r analysisType[4]` used visual models, `r analysisType[2]` used multiple analysis types (e.g.
+visual model and colourmetric) and `r analysisType[3]` did other forms of analysis (e.g.
+RGB, PCA's, GLMMs etc.).
+Of the studies quantifying colour using spectral reflectance curves `r analysisTypeSpec[1]` used colormetric measures, `r analysisTypeSpec[2]` used multiple analyses, `r analysisTypeSpec[4]` used visual models and `r analysisTypeSpec[3]` used other forms of analysis.

### 2014-11-06 R/analysis.r
Added more analysis on how studies are analysing their colour

@@ -77,3 +77,11 @@ analysisTypeSpec <- table(DatSpec$analysis_type)
 intTime <- (sum(DatSpec$int_time)/nrow(DatSpec))*100
 angle   <- (sum(DatSpec$spec_angle)/nrow(DatSpec))*100
 dist    <- (sum(DatSpec$spec_dist)/nrow(DatSpec))*100
+
+## Of colormetric studies how many defined their colormetrics?
+
+col_def <- sum(coldat$colmetric_def, na.rm = TRUE)/(length(coldat$colmetric_def[!is.na(coldat$colmetric_def)]))
+
+## How did studies fair in their reporting of visual models?
+
+DatSpec

### 2014-11-07 R/analysis.r
Added new analysis script

@@ -46,30 +46,30 @@ camModel  <- (sum(coldat$cam_model, na.rm = TRUE)/(cam+both))*100
 DatSpec   <- subset(coldat, stud_type == 1 |stud_type == 3)
 
 # Note that we need to change the rows where spec and cemera differ
-specLight <- (sum(DatSpec$light_source)/nrow(DatSpec))*100
+specLight <- (sum(DatSpec$light_source)/length(DatSpec$light_source))*100
 
 ## Of the studies using a cam, what proportion report the light source. Note that column needs to be extracted because it combines cameras and specs
 
 DatCam    <- subset(coldat, stud_type == 2 |stud_type == 3)
-CamLight  <- (sum(DatCam$light_source, na.rm = TRUE)/nrow(DatCam))*100
+CamLight  <- (sum(DatCam$light_source, na.rm = TRUE)/length(DatCam$light_source))*100
 
-## Of Studies spec'ing what proportion report a dark standard
-drk_std   <- (sum(DatSpec$dark_std)/(nrow(DatSpec)))*100
-white_stdspec <- (sum(DatSpec$white_stdspec)/nrow(DatSpec))*100
+## Of studies spec'ing what proportion report a dark standard
+drk_std       <- (sum(DatSpec$dark_std)/(length(DatSpec$dark_std)))*100
+white_stdspec <- (sum(DatSpec$white_stdspec)/length(DatSpec$white_stdspec))*100
 
 ## What proportion of studies specing report the num of specs averaged
 
-specAVG <- (sum(DatSpec$specpix_avg)/nrow(DatSpec))*100
+specAVG <- (sum(DatSpec$specpix_avg)/length(DatSpec$specpix_avg))*100
 
 ## What proportion of studies using a camera report the num of pixs averaged
 
-camAVG <- (sum(DatCam$specpix_avg)/nrow(DatCam))*100
+camAVG  <- (sum(DatCam$specpix_avg)/length(DatCam$specpix_avg))*100
 
+##--------------------------------------
 ## How do studies analyse colour?
 analysisType <- table(coldat$analysis_type)
 
 ## How about just spec studies because they can do visual modelling, or colorimetric
-
 analysisTypeSpec <- table(DatSpec$analysis_type)
 
 ## What proportion of spec studies reported 1) the integration time, spec angle and distance?
[…]

### 2014-11-07 docs/ms.Rmd
Update ms.Rmd
Added a bit more analysis
brought back two last paragaraphs in ms.Rmd. #188da13

@@
 
-#The transparency of colour reporting in plant and animal taxa
+#The transparency of methods-reporting in studies measuring biological colouration
 
@@
 Of the studies quantifying colour using spectral reflectance curves `r analysisTypeSpec[1]` used colormetric measures, `r analysisTypeSpec[2]` used multiple analyses, `r analysisTypeSpec[4]` used visual models and `r analysisTypeSpec[3]` used other forms of analysis.
+Studies using colormetric measurements defined their measure of brightness, hue and/or chroma in `r round(col_def, digits = 2)`% of the studies.
+Of the studies analysing their data using visual models, `r round(prop_vismod, digits = 2)`% reported the type of visual model they used while only `r round(prop_irrad, digits =2)`%  reported the type of irradiance, `r round(prop_spvismod, digits = 2)`% reported the species used in their visual model and `r round(propvis_mod_param, digits = 2)`% made explicit mention of the parameters they used (i.e.
+cone densities, neural vs quantum catchs etc.) in their methods.

### 2014-11-07 R/analysis.r
Added figure 1 and some small additonal code

@@ -80,7 +80,7 @@ dist    <- (sum(DatSpec$spec_dist)/nrow(DatSpec))*100
 
 ## Of colormetric studies how many defined their colormetrics?
 
-col_def <- sum(coldat$colmetric_def, na.rm = TRUE)/(length(coldat$colmetric_def[!is.na(coldat$colmetric_def)]))
+col_def <- (sum(coldat$colmetric_def, na.rm = TRUE)/(length(coldat$colmetric_def[!is.na(coldat$colmetric_def)])))*100
 
 ## How did studies fair in their reporting of visual models?
 prop_irrad <- (sum(DatSpec$irrad_type[! is.na(DatSpec$irrad_type)])/length(DatSpec$irrad_type[! is.na(DatSpec$irrad_type)]))*100 # Note that there are a few studies under "multiple" that need to be checked because irrad_type is NA and yet the visual model they report is not given. We should check these.
@@ -91,9 +91,19 @@ prop_spvismod <- (sum(DatSpec$vis_mod_sp[! is.na(DatSpec$vis_mod_sp)])/length(Da
 
 propvis_mod_param <- (sum(DatSpec$vis_mod_param[! is.na(DatSpec$vis_mod_param)])/length(DatSpec$vis_mod_param[! is.na(DatSpec$vis_mod_param)]))*100
 
+## Number of studies referencing other work.
+ref_stud <- (sum(coldat$prev_pub)/length(coldat$prev_pub))*100
 
 ##------------------------------------ Figure 1--------------------------------------##
 # Figure of proportions on the hardware/software used
 
 props   <- c(specModel, camModel, specLight, CamLight, drk_std, white_stdspec, specAVG, camAVG)
-N_props <- c((specs+both), (cam+both), length(DatSpec$light_source), length(DatCam$light_source), )
+N_props <- c((specs+both), (cam+both), length(DatSpec$light_source), length(DatCam$light_source), length(DatSpec$dark_std), length(DatSpec$white_stdspec), length(DatSpec$specpix_avg), length(DatCam$specpix_avg))
+names   <- paste("C", seq(1:8), sep = "")
+dat     <- t(data.frame(props))
+colnames(dat) <- names
+
+barplot(dat, ylim = c(0,100), ylab = "Percentages of studies reporting", xlab = "Criteria", col = "gray") -> bp.out
+abline(h = 0, lwd = 2)
+text(N_props, x = bp.out, y = 98)
+

### 2014-11-07 docs/ms.Rmd
Added figure 1 chunk to ms and also finished results section. Still need to add one more figure, plus fix up fig 1 and also add the additonal visual modelling stuff that Tom did in the text.

@@
 cone densities, neural vs quantum catchs etc.) in their methods.
+It is important to note that papers often made reference to previous work for various methods (`r ref_stud`%) and so these figures are slightly liberal given that referenced work may have comprehensively covered some of these criteria.
+Despite this, many of the studies that did not reference previous work were lacking in their reporting of many aspects of our criteria suggesting the need for more careful and comprehensive reporting within studies.
+
+```{r, eval = TRUE, echo = FALSE}
+
+barplot(dat, ylim = c(0,100), ylab = "Percentages of studies reporting", xlab = "Criteria", col = "gray") -> bp.out
+abline(h = 0, lwd = 2)
+text(N_props, x = bp.out, y = 98)
+
+```

### 2014-11-07 docs/ms.Rmd
Got example ref working (see abstract) - I'll do the rest when it's not Friday night. Also moved styling to .sty file

@@
----
-fontsize: 12pt
-geometry: margin=0.75in
-output: pdf_document
----
-
-#The transparency of methods-reporting in studies measuring biological colouration
+# The transparency of methods-reporting in studies measuring biological colouration
 
@@
 
-##Abstract
+## Abstract
 
-##Introduction
+TEST REFERENCE [@endler_comparing_2005; @endler_convergent_1982]
+
+## Introduction
 Since Thayer (1909), Poulton (1890) and Wallace (YEAR) generations of biologists have endeavored to explain the mechanisms and functions of animal and plant coloration.
@@
 
-##Ideal reporting in colour studies
+## Ideal reporting in colour studies
 
@@
 
-##What is reporting like in the literature?
+## What is reporting like in the literature?
 
@@
 ```
+
+## References

### 2014-11-07 R/analysis.r
Completed, rather crappily, the figure and analysis section. Codes crap and can be cleaned but works

@@ -13,9 +13,6 @@ packages <- c("car", "ggplot2")
 install.packages(packages)
 sapply(packages, function(x) library(x, character.only = TRUE, verbose = FALSE))
 
-## Set working dir. path.expand gives a bit of flexibility
-setwd(path.expand("~/Dropbox/Share\ Folder/colsci_rep/"))
-
 ## Load data 
 coldat <- read.csv("data/col_data.csv", header = TRUE)
 str(coldat)
@@ -83,6 +80,9 @@ dist    <- (sum(DatSpec$spec_dist)/nrow(DatSpec))*100
 col_def <- (sum(coldat$colmetric_def, na.rm = TRUE)/(length(coldat$colmetric_def[!is.na(coldat$colmetric_def)])))*100
 
 ## How did studies fair in their reporting of visual models?
+
+vismod_type <- table(DatSpec$vis_mod_type)
+
 prop_irrad <- (sum(DatSpec$irrad_type[! is.na(DatSpec$irrad_type)])/length(DatSpec$irrad_type[! is.na(DatSpec$irrad_type)]))*100 # Note that there are a few studies under "multiple" that need to be checked because irrad_type is NA and yet the visual model they report is not given. We should check these.
 
 prop_vismod <- (sum(DatSpec$vis_mod[! is.na(DatSpec$vis_mod)])/length(DatSpec$vis_mod[! is.na(DatSpec$vis_mod)]))*100
@@ -91,19 +91,50 @@ prop_spvismod <- (sum(DatSpec$vis_mod_sp[! is.na(DatSpec$vis_mod_sp)])/length(Da
 
 propvis_mod_param <- (sum(DatSpec$vis_mod_param[! is.na(DatSpec$vis_mod_param)])/length(DatSpec$vis_mod_param[! is.na(DatSpec$vis_mod_param)]))*100
 
+prop_adap <- (sum(DatSpec$vis_mod_adapt[! is.na(DatSpec$vis_mod_adapt)])/length(DatSpec$vis_mod_adapt[! is.na(DatSpec$vis_mod_adapt)]))*100
+
+prop_qcatch <- (sum(DatSpec$vis_mod_qcatch[! is.na(DatSpec$vis_mod_qcatch)])/length(DatSpec$vis_mod_qcatch[! is.na(DatSpec$vis_mod_qcatch)]))*100
+
[…]

### 2014-11-07 docs/ms.Rmd
Finalised writing of results with updated dataset from Tom and added the figure

--- 2014-11-07 docs/ms.Rmd
@@
 Studies using colormetric measurements defined their measure of brightness, hue and/or chroma in `r round(col_def, digits = 2)`% of the studies.
-Of the studies analysing their data using visual models, `r round(prop_vismod, digits = 2)`% reported the type of visual model they used while only `r round(prop_irrad, digits =2)`%  reported the type of irradiance, `r round(prop_spvismod, digits = 2)`% reported the species used in their visual model and `r round(propvis_mod_param, digits = 2)`% made explicit mention of the parameters they used (i.e.
+
+Most studies using visual models analysed their data using receptor noise models (`r vismod_type[3]` of `r sum(vismod_type)` studies) while others used hexagon (`r vismod_type[1]` studies) or tcs (`r vismod_type[4]` studies).
+Case specific models were used in `r vismod_type[2]` cases.
+Overall, `r round(prop_irrad, digits =2)`% reported the type of irradiance, `r round(prop_spvismod, digits = 2)`% the species used in their visual model and `r round(propvis_mod_param, digits = 2)`% made explicit mention of the parameters they used (i.e.
 cone densities, neural vs quantum catchs etc.) in their methods.
+The background used in visual models was reported `r round(prop_bkg, digits = 2)`% of the time.
+The type of visual noise authors assumed was reported in `r round(prop_noise, digits = 2)`% of the studies.
+In addition, explicit mention of the type of quantum catche (`r round(prop_qcatch, digits = 2)`%) or whether photoreceptor adaptation (`r round(prop_adap, digits = 2)`%) was accounted for in visual models was not common.
+
+While some of the values reported above seem alarming, it is important to note that papers often made reference to previous work for various methods (`r round(ref_stud, digits = 2)`%) and so these figures are slightly liberal given that referenced work may have comprehensively covered some of these criteria.
+Despite this, many of the studies that did not reference previous work were lacking in their reporting of many aspects of our criteria suggesting the need for more comprehensive reporting of data collection and analysis methods within studies.
+
+```{r, barplot, eval = TRUE, echo = FALSE, fig.ext = "pdf", fig.cap = "Figure 1"}
+
+par(mfrow=c(1,2))
+props   <- c(specModel, camModel, specLight, CamLight, drk_std, white_stdspec, specAVG, camAVG)
+N_props <- c((specs+both), (cam+both), length(DatSpec$light_source), length(DatCam$light_source), length(DatSpec$dark_std), length(DatSpec$white_stdspec), length(DatSpec$specpix_avg), length(DatCam$specpix_avg))
+
+names   <- c("Spec", "Cam", "Spec L", "Pix Avg"," Cam L"," Wt Std", "Spec Avg", "Drk Std")
+dat     <- t(arrange(data.frame(props, N_props), props, decreasing = TRUE))
+colnames(dat) <- names
+
+barplot(dat[1,], ylim = c(0,100), ylab = "Percentage of studies reporting criteria", xlab = "Criteria", col = "gray", space = 0.30, cex.names = 0.72, mgp = c(2.5,0.5,0), cex.axis = 1.2, cex.lab = 1.5) -> bp.out
+abline(h = 0, lwd = 2)
+text(dat[2,], x = bp.out, y = 98)
+text(paste(round(dat[1,], digits = 0), "%", sep = ""), x = bp.out, y = round(dat[1,], digits = 0)*0.5)
+mtext("a)", adj = -0.10, padj = -1, cex = 2)
+
+props_analy   <- c(prop_irrad, prop_spvismod, propvis_mod_param, prop_adap, prop_qcatch, prop_bkg, prop_noise)
+
+N_props_analy <- c(length(DatSpec$irrad_type[!
+is.na(DatSpec$irrad_type)]), length(DatSpec$vis_mod_sp[!
+is.na(DatSpec$vis_mod_sp)]), length(DatSpec$vis_mod_param[!
+is.na(DatSpec$vis_mod_param)]), length(DatSpec$vis_mod_adapt[!
+is.na(DatSpec$vis_mod_adapt)]), length(DatSpec$vis_mod_qcatch[!
+is.na(DatSpec$vis_mod_qcatch)]), length(DatSpec$vis_mod_bkg[!
+is.na(DatSpec$vis_mod_bkg)]), length(DatSpec$vis_mod_noise_type[!
+is.na(DatSpec$vis_mod_noise_type)]))
+
+dat_analy     <- t(arrange(data.frame(props_analy, N_props_analy), props_analy, decreasing = TRUE))
+
+names_analy   <- c("SpRep", "Bkg", "Irrad", "Qcatch","Noise"," ChrAdapt", "ModPara")
+colnames(dat_analy) <- names_analy
+
+barplot(dat_analy[1,], ylim = c(0,100), ylab = "Percentage of studies reporting criteria", xlab = "Criteria", col = "gray", space = 0.30, cex.names = 0.72, mgp = c(2.5,0.5,0), cex.axis = 1.2, cex.lab = 1.5) -> bp.out
+abline(h = 0, lwd = 2)
+text(dat_analy[2,], x = bp.out, y = 98)
+text(paste(round(dat_analy[1,], digits = 0), "%", sep = ""), x = bp.out, y = round(dat_analy[1,], digits = 0)*0.5)
+mtext("b)", adj = -0.10, padj = -1, cex = 2)
+
+```

### 2014-11-10 docs/ms.Rmd
Fixed issues with compiling manuscript. Head conflict was still in file and cauing problems so I deleted this and everything compiles perfectly.

@@
 The type of visual noise authors assumed was reported in `r round(prop_noise, digits = 2)`% of the studies.
-In addition, explicit mention of the type of quantum catche (`r round(prop_qcatch, digits = 2)`%) or whether photoreceptor adaptation (`r round(prop_adap, digits = 2)`%) was accounted for in visual models was not common.
+In addition, explicit mention of the type of quantum catche (`r round(prop_qcatch, digits = 2)`% of studies) or whether photoreceptor adaptation (`r round(prop_adap, digits = 2)`%) was accounted for in visual models was not common.
 

### 2014-11-10 R/analysis.r
Added plyr package to required packages

@@ -9,7 +9,7 @@
 rm(list=ls())
 
 ## Packages. Coded to simplify multiple package installation and loading.
-packages <- c("car", "ggplot2")
+packages <- c("car", "ggplot2", "plyr")
 install.packages(packages)
 sapply(packages, function(x) library(x, character.only = TRUE, verbose = FALSE))
 

### 2014-11-10 R/analysis.r
Fixed up figure 1 and made it analysis script generate the figure in output folder to be attached/uploaded

@@ -106,7 +106,9 @@ ref_stud <- (sum(coldat$prev_pub)/length(coldat$prev_pub))*100
 
 ##------------------------------------ Figure 1--------------------------------------##
 # Figure of proportions on the hardware/software reported
-pdf(height = 9.45, width = 19)
+setwd(paste(getwd(), "/output/figures", sep = ""))
+
+pdf(height = 6.30, width = 12.40)
 par(mfrow=c(1,2))
 props   <- c(specModel, camModel, specLight, CamLight, drk_std, white_stdspec, specAVG, camAVG)
 N_props <- c((specs+both), (cam+both), length(DatSpec$light_source), length(DatCam$light_source), length(DatSpec$dark_std), length(DatSpec$white_stdspec), length(DatSpec$specpix_avg), length(DatCam$specpix_avg))
@@ -119,7 +121,7 @@ barplot(dat[1,], ylim = c(0,100), ylab = "Percentage of studies reporting criter
 abline(h = 0, lwd = 2)
 text(dat[2,], x = bp.out, y = 98)
 text(paste(round(dat[1,], digits = 0), "%", sep = ""), x = bp.out, y = round(dat[1,], digits = 0)*0.5)
-mtext("a)", adj = -0.10, padj = -1, cex = 2)
+mtext("a)", adj = -0.15, padj = -1, cex = 2)
 
 
 props_analy   <- c(prop_irrad, prop_spvismod, propvis_mod_param, prop_adap, prop_qcatch, prop_bkg, prop_noise)
@@ -135,6 +137,8 @@ barplot(dat_analy[1,], ylim = c(0,100), ylab = "Percentage of studies reporting 
 abline(h = 0, lwd = 2)
 text(dat_analy[2,], x = bp.out, y = 98)
 text(paste(round(dat_analy[1,], digits = 0), "%", sep = ""), x = bp.out, y = round(dat_analy[1,], digits = 0)*0.5)
-mtext("b)", adj = -0.10, padj = -1, cex = 2)
+mtext("b)", adj = -0.15, padj = -1, cex = 2)
 
 dev.off()
+
+setwd("../..")

### 2014-11-10 R/analysis.r
Minor fix to label figure 1 output.

@@ -108,7 +108,7 @@ ref_stud <- (sum(coldat$prev_pub)/length(coldat$prev_pub))*100
 # Figure of proportions on the hardware/software reported
 setwd(paste(getwd(), "/output/figures", sep = ""))
 
-pdf(height = 6.30, width = 12.40)
+pdf(file = "figure1.pdf", height = 6.30, width = 12.40)
 par(mfrow=c(1,2))
 props   <- c(specModel, camModel, specLight, CamLight, drk_std, white_stdspec, specAVG, camAVG)
 N_props <- c((specs+both), (cam+both), length(DatSpec$light_source), length(DatCam$light_source), length(DatSpec$dark_std), length(DatSpec$white_stdspec), length(DatSpec$specpix_avg), length(DatCam$specpix_avg))

### 2014-11-10 docs/ms.Rmd
Wrote some really crappy concluding paragraph...I mean really crappy

@@
 
-```{r, barplot, eval = TRUE, echo = FALSE, fig.ext = "pdf", fig.cap = "Figure 1"}
+##Conclusions
 
-par(mfrow=c(1,2))
-props   <- c(specModel, camModel, specLight, CamLight, drk_std, white_stdspec, specAVG, camAVG)
-N_props <- c((specs+both), (cam+both), length(DatSpec$light_source), length(DatCam$light_source), length(DatSpec$dark_std), length(DatSpec$white_stdspec), length(DatSpec$specpix_avg), length(DatCam$specpix_avg))
-
-names   <- c("Spec", "Cam", "Spec L", "Pix Avg"," Cam L"," Wt Std", "Spec Avg", "Drk Std")
-dat     <- t(arrange(data.frame(props, N_props), props, decreasing = TRUE))
-colnames(dat) <- names
-
-barplot(dat[1,], ylim = c(0,100), ylab = "Percentage of studies reporting criteria", xlab = "Criteria", col = "gray", space = 0.30, cex.names = 0.72, mgp = c(2.5,0.5,0), cex.axis = 1.2, cex.lab = 1.5) -> bp.out
-abline(h = 0, lwd = 2)
-text(dat[2,], x = bp.out, y = 98)
-text(paste(round(dat[1,], digits = 0), "%", sep = ""), x = bp.out, y = round(dat[1,], digits = 0)*0.5)
-mtext("a)", adj = -0.10, padj = -1, cex = 2)
-
-props_analy   <- c(prop_irrad, prop_spvismod, propvis_mod_param, prop_adap, prop_qcatch, prop_bkg, prop_noise)
-
-N_props_analy <- c(length(DatSpec$irrad_type[!
-is.na(DatSpec$irrad_type)]), length(DatSpec$vis_mod_sp[!
-is.na(DatSpec$vis_mod_sp)]), length(DatSpec$vis_mod_param[!
-is.na(DatSpec$vis_mod_param)]), length(DatSpec$vis_mod_adapt[!
-is.na(DatSpec$vis_mod_adapt)]), length(DatSpec$vis_mod_qcatch[!
-is.na(DatSpec$vis_mod_qcatch)]), length(DatSpec$vis_mod_bkg[!
-is.na(DatSpec$vis_mod_bkg)]), length(DatSpec$vis_mod_noise_type[!
-is.na(DatSpec$vis_mod_noise_type)]))
-
-dat_analy     <- t(arrange(data.frame(props_analy, N_props_analy), props_analy, decreasing = TRUE))
-
-names_analy   <- c("SpRep", "Bkg", "Irrad", "Qcatch","Noise"," ChrAdapt", "ModPara")
-colnames(dat_analy) <- names_analy
-
-barplot(dat_analy[1,], ylim = c(0,100), ylab = "Percentage of studies reporting criteria", xlab = "Criteria", col = "gray", space = 0.30, cex.names = 0.72, mgp = c(2.5,0.5,0), cex.axis = 1.2, cex.lab = 1.5) -> bp.out
-abline(h = 0, lwd = 2)
-text(dat_analy[2,], x = bp.out, y = 98)
-text(paste(round(dat_analy[1,], digits = 0), "%", sep = ""), x = bp.out, y = round(dat_analy[1,], digits = 0)*0.5)
-mtext("b)", adj = -0.10, padj = -1, cex = 2)
-
-```
+Our objective in this paper was not to point to single studies and lay blame, but to highlight an important gap in reporting that can influence the reproducibility of colour research and cross taxa comparisons.
+We hope our critical evaluation of recent litearture has highlighted a need for more careful reporting of methological reporting of how colour was measured, but also how it was analysed and can guide how research is communicated improving on the transparency and reproducibility of colour-based science.
+We belive that part of the issue is simply that researchers seldom have clear insight and guidance on what might be important aspects of their analysis and methods to be reported and we hope our paper will help estbalish more uniform reporting guidelines.
 

### 2014-11-10 R/analysis.r
Wrote first draft of conclusions. I still need to fix figure as I missed a few important criteria. i.e. distance and angle

@@ -102,7 +102,9 @@ prop_noise <- (sum(DatSpec$vis_mod_noise_type[! is.na(DatSpec$vis_mod_noise_type
 ## Number of studies referencing other work.
 ref_stud <- (sum(coldat$prev_pub)/length(coldat$prev_pub))*100
 
+## Percentage of showing reflectance figures
 
+fig <- (sum(DatSpec$refl_fig)/length(DatSpec$refl_fig))*100
 
 ##------------------------------------ Figure 1--------------------------------------##
 # Figure of proportions on the hardware/software reported

### 2014-11-10 docs/ms.Rmd
First draft of the conclusions finished

@@
 
-Rhiannon L. Dalrymple^1,6^, Daniel W. A. Noble^2,6^, James C. O’Hanlon ^2,6^, Kate D. L. Umbers^3,4,6^, Thomas E. White^2,6^, Daniel B. Zurek^5,6^
+Thomas E. White^2,6^, Rhiannon L. Dalrymple^1,6^, Daniel W. A. Noble^2,6^, James C. O’Hanlon ^2,6^, Daniel B. Zurek^5,6^, Kate D. L. Umbers^3,4,6^
 
@@
 
-^6^ Authors contributed equally and are presented in alphabetical order
+^6^ Authors contributed equally and are presented in random order
 
@@
 
-Our objective in this paper was not to point to single studies and lay blame, but to highlight an important gap in reporting that can influence the reproducibility of colour research and cross taxa comparisons.
-We hope our critical evaluation of recent litearture has highlighted a need for more careful reporting of methological reporting of how colour was measured, but also how it was analysed and can guide how research is communicated improving on the transparency and reproducibility of colour-based science.
-We belive that part of the issue is simply that researchers seldom have clear insight and guidance on what might be important aspects of their analysis and methods to be reported and we hope our paper will help estbalish more uniform reporting guidelines.
+Our review of recent colour literarure highlights some major reporting gaps that will hinder the reproducibility of studies quantifying colour signals.
+While we recognize that journals have space limitations and that work may reference previous methological and analytical details, we strongly beleive that studies can do a much better job reporting details which are essential to the transparency and reproducibility of their work.
+When possible, we also recommend that authors justify data capture and analytical approaches - something few studies we investiagted expliclty stated.
+While all criteria are important some deserve special attanetion.
+In particular we found that colour scientists could do a much better job reporting light sources, number of pixel and spectra averaged used for analysis and the type of standards used.
+More improtantly, the distance and angle should be standard reporting for all studies using spectrophotomtry.
+These all have important consequences on downstream analysis and the reproducibility of colour signal quantification.
+As we all know, a picture can say 1000 words and we recommend the incoperation of spectral reflectance curves were possible as it provides insight on the colour signal, which is visually appealing and easy to understand by most readers.
+Surprisingly, we found that only `r round(fig, digits = 2)`% of studies using spectral reflectance data had an explicit figure in their paper showing these curves.
+Visual models were regularly used to analyse spectral data, something we welcome.
+However, there were many important aspects of the analysis that were seldom reported, inclding the type of quantum catch used, whether photoreceptor adaptation was accounted for and the type of neural noise used in receptor noise models.
+Photoreceptor densities should also be reported where possible even if using standard avian photorecptor densities.
+
+In conclusion, our objective in this paper was not to point to single studies and lay blame, but to highlight an important gap in reporting that can influence the reproducibility of colour research and cross taxa comparisons.
+Indeed, even in our own work (e.g.
+Umbers et al.
+2013) we have failed to report important details, which in resptrospec can lead to confusiion about how such data was collected and analysed.
+Nontheless, our critisim of reporting is meant to be constructive and is not to say that studies lacking improtant details are not excellent works, but it is only through critical appraisal and colleageal discussion that we can identify these gaps and constructively contribute to the growth of our field.
+Such assessments are particlularly important in an age where accountability, transparency and reproducibility of are utmost importance.
+We hope our critical evaluation of recent litearture has highlighted a need for more careful reporting of how colour signals have been measured and analysed.
+We beleive that part of the issue is simply that researchers seldom have clear guidelines to follow when undertaking research and miss what might be important aspects of their methods and analysis that need to be reported.
+We therefore hope that our paper will help estbalish more uniform reporting guidelines in the literature so that we can continue an exciting era of colour signal research in the life sciences.
 

### 2014-11-11 docs/ms.Rmd
Add in most other references

@@
 
-TEST REFERENCE [@endler_comparing_2005; @endler_convergent_1982]
-
 ## Introduction
-Since Thayer (1909), Poulton (1890) and Wallace (YEAR) generations of biologists have endeavored to explain the mechanisms and functions of animal and plant coloration.
+Since Thayer [-@thayer_concealing_1909], Poulton [-@poulton_colours_1890] and Wallace [-@wallace_natural_1891] generations of biologists have endeavored to explain the mechanisms and functions of animal and plant coloration.
 For these biologists, identifying best practice in measuring color and color pattern and in determining how other species see colour has been a great challenge.
@@
 
-The application of spectrophotometry to questions in ecology and evolution revolutionized the field (Dyck 1966).
-Before the widespread use of spectrophotometry researchers assigned categorical color descriptions, arbitrary rankings or scores (e.g.
-Aldrich & James, 1991; Wheelwright & Janson, 1985) or matched paint swatches or munsell chip ratings based on their own visual assessment (e.g.
-Key and Day 1954, Burtt 1986).
-Although no doubt applied carefully, these techniques limited progress in field.
-For example, many researchers designed color assessment methods specifically for their projects (e.g.
-Alrich and James 1991) rendering data by different researchers incomparable.
+The application of spectrophotometry to questions in ecology and evolution revolutionized the field [@dyck_determination_1966].
+Before the widespread use of spectrophotometry, researchers assigned categorical color descriptions, arbitrary rankings or scores [e.g.
+@aldrich_ecogeographic_1991; @wheelwright_colors_1985] or matched paint swatches or munsell chip ratings based on their own visual assessment [e.g.
+@burtt_analysis_1986; @key_temperature_1954].
+Although no doubt applied carefully, these techniques limited progress in the field.
+For example, many researchers designed color assessment methods specifically for their projects [e.g.
+@aldrich_ecogeographic_1991] rendering data by different researchers incomparable.
 This was/is especially problematic when different measurement techniques are used on the same species (e,g, refXXX vs ref XXX).
-Also, human vision, blind to ultraviolet wavelengths, is inherently different to most other animals such as avian and reptilian tetrachromats and insect trichromats, (Bennett 1994, Douglas and Jeffrey 2014); human perception of color, and variation therein, is not only biologically irrelevant to most systems, but it can lead to erroneous conclusions (Bennett, Cuthill and Norris 1994, Endler 1990).
+Also, human vision, insensitive to ultraviolet wavelengths, is inherently different to most other animals such as avian and reptilian tetrachromats, and insect trichromats [@bennett_sexual_1994](MORE).
+Human perception of color, and variation therein, is not only biologically irrelevant to most systems, but it can lead to erroneous conclusions [@bennett_sexual_1994; @endler_measurement_1990].
 
@@
 However one drawback to spectrophotometry, is the ease with which erroneous data can be captured if attention is not paid to finer details of spectrum capture.
-For example, the angle from which spectra are taken (Santos & Lumeij, 2007), consistent calibration of light and dark standards, the distance between light source and object, the expectations of the researcher and the size, shape, glossiness or iridescence of the specimen (Andersson and Prager, 2006, Schaefer et al., 2008, Kemp, Reznick & Grether, 2008, Meadows et al 2011), can all contribute to potentially skewed data or alter results.
+For example, the angle from which spectra are taken [@santos_strong_2007], consistent calibration of light and dark standards, the distance between light source and object, the expectations of the researcher and the size, shape, glossiness or iridescence of the specimen [@andersson_quantifying_2006; @kemp_ornamental_2008; @meadows_quantifying_2011; @schaefer_birds_2008], can all contribute to potentially skewed data or alter results.
 Experience, guidance and objectivity are necessary to ensure data are consistent and representative of the subject and to ensure repeatability, conditions under which data were collected must be reported exactly.
@@
 In conjunction with spectrophotometry, digital photography is increasingly being implemented in color quantification.
-High-resolution cameras are now inexpensive and portable, allow instantaneous sampling of multiple colour patches, and can produce highly repeatable images (McKay 2013).
-The main challenges for gathering and using color data from photographs are ensure correct standardization methods such as linearization and consistent lighting (Stevens et al.
-2007; ref Dariella Whitiyev 2014’s paper).
-Crucially, most digital cameras used in color science are consumer products designed for human visual color space and hardware modifications for ultraviolet and infrared imaging are necessary (e.g.
-Chiao et al.
-2009, McKay 2013), making photography complementary, rather than superior to spectrophotometry.
+High-resolution cameras are now inexpensive and portable, allow instantaneous sampling of multiple colour patches, and can produce highly repeatable images [@mckay_use_2013].
+The main challenges for gathering and using color data from photographs are ensure correct standardization methods such as linearization are consistent lighting [@stevens_using_2007].
+Crucially, most digital cameras used in color science are consumer products designed for human visual color space and hardware modifications for ultraviolet and infrared imaging are necessary [e.g.
+@chiao_visualization_2009; @mckay_use_2013], making photography complementary, rather than superior to spectrophotometry.
 
 Expansion in the availability of objective methods for the measurement of color has been matched by advances in theory and analysis.
-In particular, the development of sensory models has produced significant insights into the way in which colorful stimuli are perceived (Bennett et al 1994, Chittka 1992, Endler & Mielke 2005, Vorobyev & Osorio 1998).
+In particular, the development of sensory models has produced significant insights into the way in which colorful stimuli are perceived [@chittka_colour_1992; @endler_comparing_2005; @vorobyev_receptor_1998].
 Where the receptor sensitivities of an animal’s visual system are known, the physiological response of that animal’s visual system when viewing a given color can be predicted.
-Physiological sensory models of animal vision have been developed to model the perception of bees (e.g.
-Chittka, 1992), flies (e.g.
-Troje, 1993), birds (e.g.
-Endler & Mielke, 2005), moths (e.g.
-Chuang, Yang & Tso, 2007), frogs (REF), crabs (REF) and snakes (e.g.
-Maan & Cummings, 2012).
 
-Sensory models typically attempt to describe the subjective perception of color or brightness information as a function of an object’s reflectance, the ambient illumination, and a receiver’s sensory system (eg.
-Endler & Mielke 2005, Vorobyev et al.
-1998).
+Sensory models typically attempt to describe the subjective perception of color or brightness information as a function of an object’s reflectance, the ambient illumination, and a receiver’s sensory system [e.g.
+@endler_comparing_2005; @vorobyev_receptor_1998].
 Furthermore, an animal’s ability to distinguish between reflectance spectra can then be predicted based on the differential responses of a visual system to varying reflectance spectra.
 This approach enables researchers to move beyond purely quantitative comparisons of reflectance spectra and adopt biologically relevant perspectives of signal receivers to test/formulate hypotheses.
-Although relatively easy to implement and potentially powerful, sensory models are built on multiple assumptions about the way in which stimuli are visually and cognitively processed that can dramatically shape the results of a given analysis (Lind & Kelber 2009; Pike 2012).
-For example, Vorobyev et al.’s (1998; 2001) widely used receptor-noise model requires researchers to specify the form of receptor quantum catches (i.e.
-raw versus log-transformed), the existence of any color correction mechanism (e.g.
+Although relatively easy to implement and potentially powerful, sensory models are built on multiple assumptions about the way in which stimuli are visually and cognitively processed that can dramatically shape the results of a given analysis [@lind_avian_2009; @pike_preserving_2012].
+For example, Vorobyev et al.’s [-@vorobyev_receptor_1998; @vorobyev_colour_2001] widely used receptor-noise model requires researchers to specify the form of receptor quantum catches (i.e.
+raw versus log-transformed), the use of any color constancy mechanism (e.g.
 von Kries), as well as the photoreceptor signal-to-noise ratios, their relative densities, and the type of noise predominating (i.e.
 neural and/or receptor noise).
-These data are only available for a small number of model species (e.g.
-honeybees) and thus assumptions and extrapolations are necessary when models are applied to related species.
-Unless the biological rationale underlying every parameter choice is clearly understood, justified, and reported, such analyses may become ineffective, especially when comparing results between studies (Pike 2012).
+These data are only available for a small number of model species, such as honeybees, and so assumptions and extrapolations are necessary when models are applied to related species.
+Unless the biological rationale underlying every parameter choice is clearly understood, justified, and reported, such analyses may become ineffective, especially when comparing results between studies [@pike_preserving_2012].
 
-Since the repeatability and transparency of science is paramount, it is imperative that we regularly and constructively reflect on our own practices (Peng, 2009; Alsheikh-Ali et al.
-2011; Sandve et al YEAR).
+Since the repeatability and transparency of science is paramount, it is imperative that we regularly and constructively reflect on our own practices [@alsheikh_public_2011; @sandve_ten_2013].
 The potential for detrimental subjectivity in color measurement and the mathematical opacity of visual models for the average colour scienctist means that, as in any field, is it critical that methods are reported rigorously.
-Methods reporting should ideally have well defined standards such that sufficient detail is provided to allow comprehensive peer review, precise replication, cross-study comparison and data sharing (REF data sharing papers, Ask Rob Lanfear).
+Methods reporting should ideally have well defined standards such that sufficient detail is provided to allow comprehensive peer review, precise replication, cross-study comparison and data sharing [@alsheikh_public_2011](Ask Rob Lanfear).
 The aims of this paper are to (a) review methods reporting in papers published on colour in 2013 from 20 top tier journals and (b) provide a set of recommendations for best-practice in reporting on the most common methods.
-Our best hope is that this instructional provides a useful set of guidelines to improve the reproducibility and transparency of studies on color.
+Our hope is that this instructional provides a useful set of guidelines to improve the reproducibility and transparency of studies on color.
 
@@
 
-We developed a list of metrics we deemed critical to be reported for reproducibility of color methods and analysis loosely based on Andersson and Prager (2006) and Montgomerie (2006) (Table 1 and 2).
-These included quantitative methods including spectrophotometric and photographic techniques (as in REF), but not qualitative methods such as the use of Munsell chips or paint swatches (as in REF).
+We developed a list of metrics we deemed critical to be reported for reproducibility of color methods and analysis loosely based on Andersson and Prager [-@andersson_quantifying_2006] and Montgomerie [@montgomerie_analyzing_2006] (Table 1 and 2).
+These included quantitative methods including spectrophotometric and photographic techniques, but not qualitative methods such as the use of Munsell chips or paint swatches.
 For visual modeling techniques we characterized a set of criteria required for reproducing the analysis (Table 2).
@@
 In conclusion, our objective in this paper was not to point to single studies and lay blame, but to highlight an important gap in reporting that can influence the reproducibility of colour research and cross taxa comparisons.
-Indeed, even in our own work (e.g.
-Umbers et al.
-2013) we have failed to report important details, which in resptrospec can lead to confusiion about how such data was collected and analysed.
+Indeed, even in our own work [@umbers_bright_2013] we have failed to report important details, which in resptrospec can lead to confusion about how such data were collected and analysed.
 Nontheless, our critisim of reporting is meant to be constructive and is not to say that studies lacking improtant details are not excellent works, but it is only through critical appraisal and colleageal discussion that we can identify these gaps and constructively contribute to the growth of our field.

### 2014-11-12 docs/ms.Rmd
Updated the ms and re-wrote parts of the crappy paragraphs that I wrote. Also added in some important notes about annominity in the data....which reminds me we need to remove these details from the data.

@@
 Since Thayer [-@thayer_concealing_1909], Poulton [-@poulton_colours_1890] and Wallace [-@wallace_natural_1891] generations of biologists have endeavored to explain the mechanisms and functions of animal and plant coloration.
-For these biologists, identifying best practice in measuring color and color pattern and in determining how other species see colour has been a great challenge.
+For these biologists, identifying best practices in measuring color and color pattern and in determining how other species see colour has been a great challenge.
 Over the past 150 years biologists, both independently and in collaboration with physicists and chemists, have developed a range of techniques used to measure and analyse animal and plant coloration including spectrophotometry (also called spectroradiometry, spectroscopy ETC), photography and visual modeling.
@@
 @burtt_analysis_1986; @key_temperature_1954].
-Although no doubt applied carefully, these techniques limited progress in the field.
+Although when applied carefully they maybe useful, these techniques in general have limited progress in the field.
 For example, many researchers designed color assessment methods specifically for their projects [e.g.
@@
 
-Measuring reflectance spectra is now a reasonably inexpensive and convenient technique.
-Spectrophotometry allows the capture of quantitative and arguably repeatable data and has been rapidly taken up by researchers the field.
+Measuring reflectance spectra is now a reasonably inexpensive and convenient technique fo quantifying color signals.
+Spectrophotometry allows the capture of quantitative and arguably repeatable data and has been rapidly taken up by researchers in the field.
 However one drawback to spectrophotometry, is the ease with which erroneous data can be captured if attention is not paid to finer details of spectrum capture.
 For example, the angle from which spectra are taken [@santos_strong_2007], consistent calibration of light and dark standards, the distance between light source and object, the expectations of the researcher and the size, shape, glossiness or iridescence of the specimen [@andersson_quantifying_2006; @kemp_ornamental_2008; @meadows_quantifying_2011; @schaefer_birds_2008], can all contribute to potentially skewed data or alter results.
-Experience, guidance and objectivity are necessary to ensure data are consistent and representative of the subject and to ensure repeatability, conditions under which data were collected must be reported exactly.
+Experience, guidance and objectivity are necessary to ensure data are consistent and representative of the subject.
+In addition, to ensure repeatability conditions under which data were collected must be reported exactly.
 
@@
 High-resolution cameras are now inexpensive and portable, allow instantaneous sampling of multiple colour patches, and can produce highly repeatable images [@mckay_use_2013].
-The main challenges for gathering and using color data from photographs are ensure correct standardization methods such as linearization are consistent lighting [@stevens_using_2007].
+The main challenges for gathering and using color data from photographs are ensuring correct standardization methods, such as linearization and consistent lighting [@stevens_using_2007].
 Crucially, most digital cameras used in color science are consumer products designed for human visual color space and hardware modifications for ultraviolet and infrared imaging are necessary [e.g.
@@
 Methods reporting should ideally have well defined standards such that sufficient detail is provided to allow comprehensive peer review, precise replication, cross-study comparison and data sharing [@alsheikh_public_2011](Ask Rob Lanfear).
-The aims of this paper are to (a) review methods reporting in papers published on colour in 2013 from 20 top tier journals and (b) provide a set of recommendations for best-practice in reporting on the most common methods.
-Our hope is that this instructional provides a useful set of guidelines to improve the reproducibility and transparency of studies on color.
+The aims of this paper are to (a) provide a set of criteria for best-practices in reporting of the most common methods used to collect and analyse data and (b) assess how well we are doing in reporting these criteria.
+We review methods reporting in papers published on colour in 2013 from 23 top tier journals.
+Our hope is that this instructional provides a useful set of guidelines that colour scientists can follow that will improve the reproducibility and transparency of their work.
 
@@
 
-We developed a list of metrics we deemed critical to be reported for reproducibility of color methods and analysis loosely based on Andersson and Prager [-@andersson_quantifying_2006] and Montgomerie [@montgomerie_analyzing_2006] (Table 1 and 2).
-These included quantitative methods including spectrophotometric and photographic techniques, but not qualitative methods such as the use of Munsell chips or paint swatches.
+We developed a list of metrics we deemed critical to be reported for reproducibility of color methods and analysis.
+Our guidelines are loosely based on Andersson and Prager [-@andersson_quantifying_2006] and Montgomerie [@montgomerie_analyzing_2006] (Table 1 and 2).
+They included quantitative methods, such as spectrophotometric and photographic techniques, but not qualitative methods such as the use of Munsell chips or paint swatches.
 For visual modeling techniques we characterized a set of criteria required for reproducing the analysis (Table 2).
@@
 We then tabulated what details of the methods were reported against our criteria.
+The data along with analysis script can be found at: https://github.com/daniel1noble/colsci_rep.
+We have keep the papers used in our dataset annoymous to avoid accusing specific colleagues of inadvertant reporting.
+We feel that the issue of reporting is something attributable to the ambiguity and lack of guidelines in our field and not something specific to individual research groups.
 
@@
 
-Our review of recent colour literarure highlights some major reporting gaps that will hinder the reproducibility of studies quantifying colour signals.
+Our review of recent colour literarure highlights some important reporting gaps that will hinder the reproducibility of studies quantifying colour signals.
 While we recognize that journals have space limitations and that work may reference previous methological and analytical details, we strongly beleive that studies can do a much better job reporting details which are essential to the transparency and reproducibility of their work.
-When possible, we also recommend that authors justify data capture and analytical approaches - something few studies we investiagted expliclty stated.
 While all criteria are important some deserve special attanetion.
-In particular we found that colour scientists could do a much better job reporting light sources, number of pixel and spectra averaged used for analysis and the type of standards used.
+In particular, we found that colour scientists could do a much better job reporting light sources, the number of pixels and spectra averaged and the type of standards used.
 More improtantly, the distance and angle should be standard reporting for all studies using spectrophotomtry.
-These all have important consequences on downstream analysis and the reproducibility of colour signal quantification.
-As we all know, a picture can say 1000 words and we recommend the incoperation of spectral reflectance curves were possible as it provides insight on the colour signal, which is visually appealing and easy to understand by most readers.
-Surprisingly, we found that only `r round(fig, digits = 2)`% of studies using spectral reflectance data had an explicit figure in their paper showing these curves.
 Visual models were regularly used to analyse spectral data, something we welcome.
@@
 Photoreceptor densities should also be reported where possible even if using standard avian photorecptor densities.
+These all have important consequences on downstream analysis and the reproducibility of colour signal quantification.
+While these criteria are important additional information can also be extremely helpful to readers.
+As we all know, a picture in worth 1000 words and we recommend the incoperation of spectral reflectance curves were possible as it provides insight on the colour signal, which is visually appealing and easy to understand by most readers.
+Surprisingly, we found that only `r round(fig, digits = 2)`% of studies using spectral reflectance data had an explicit figure in their paper showing these curves.
+When possible, we also recommend that authors justify data capture and analytical approaches - something few studies we investiagted expliclty stated.
 
-In conclusion, our objective in this paper was not to point to single studies and lay blame, but to highlight an important gap in reporting that can influence the reproducibility of colour research and cross taxa comparisons.
-Indeed, even in our own work [@umbers_bright_2013] we have failed to report important details, which in resptrospec can lead to confusion about how such data were collected and analysed.
-Nontheless, our critisim of reporting is meant to be constructive and is not to say that studies lacking improtant details are not excellent works, but it is only through critical appraisal and colleageal discussion that we can identify these gaps and constructively contribute to the growth of our field.
+In conclusion, our objective in this paper was not to lay blame on single studies, but to highlight an important gap in reporting that can influence the reproducibility of colour research and cross taxa comparisons.
+Indeed, even in our own work [e.g.
+@umbers_bright_2013] we have failed to report important details, which in resptrospec can lead to confusion about how such data were collected and analysed.
+Nontheless, our critisim of reporting is meant to be constructive and is not to say that studies lacking such details are not excellent works in their own right.
+There is no doubt many papers we evaluated have made important contributions.
+Nontheless, it is only through critical appraisal and colleageal discussion that we can identify these gaps and constructively contribute to the growth of our field.
 Such assessments are particlularly important in an age where accountability, transparency and reproducibility of are utmost importance.
-We hope our critical evaluation of recent litearture has highlighted a need for more careful reporting of how colour signals have been measured and analysed.
-We beleive that part of the issue is simply that researchers seldom have clear guidelines to follow when undertaking research and miss what might be important aspects of their methods and analysis that need to be reported.
+We hope our critical evaluation of recent litearture has highlighted a need for more thorough reporting of how colour signals have been measured and analysed.
+We beleive that part of the issue is simply that researchers seldom have clear guidelines to follow when undertaking research and as such miss important aspects of their methods and analysis that should to be reported.
 We therefore hope that our paper will help estbalish more uniform reporting guidelines in the literature so that we can continue an exciting era of colour signal research in the life sciences.

### 2014-11-12 R/analysis.r
Corrected the figure in script

@@ -110,36 +110,38 @@ fig <- (sum(DatSpec$refl_fig)/length(DatSpec$refl_fig))*100
 # Figure of proportions on the hardware/software reported
 setwd(paste(getwd(), "/output/figures", sep = ""))
 
-pdf(file = "figure1.pdf", height = 6.30, width = 12.40)
+pdf(file = "figure1.pdf", height = 7, width = 15)
 par(mfrow=c(1,2))
-props   <- c(specModel, camModel, specLight, CamLight, drk_std, white_stdspec, specAVG, camAVG)
-N_props <- c((specs+both), (cam+both), length(DatSpec$light_source), length(DatCam$light_source), length(DatSpec$dark_std), length(DatSpec$white_stdspec), length(DatSpec$specpix_avg), length(DatCam$specpix_avg))
+props   <- c(specModel, camModel, specLight, CamLight, drk_std, white_stdspec, specAVG, camAVG, intTime, angle, dist)
+N_props <- c((specs+both), (cam+both), length(DatSpec$light_source), length(DatCam$light_source), length(DatSpec$dark_std), length(DatSpec$white_stdspec), length(DatSpec$specpix_avg), length(DatCam$specpix_avg), rep(nrow(DatSpec), 3))
 
-names   <- c("Spec", "Cam", "Spec L", "Pix Avg"," Cam L"," Wt Std", "Spec Avg", "Drk Std")
+names   <- c("Spec", "Cam", "SpecL", "PixAvg"," CamL"," WtSd", "SpecAvg",  "IntT", "Angle", "Dist", "DrkSd")
 dat     <- t(arrange(data.frame(props, N_props), props, decreasing = TRUE))
 colnames(dat) <- names
 
-barplot(dat[1,], ylim = c(0,100), ylab = "Percentage of studies reporting criteria", xlab = "Criteria", col = "gray", space = 0.30, cex.names = 0.72, mgp = c(2.5,0.5,0), cex.axis = 1.2, cex.lab = 1.5) -> bp.out
[…]

### 2014-11-12 docs/ms.Rmd
Tightening the intro - in progress

@@
-# The transparency of methods-reporting in studies measuring biological colouration
+# The transparency of methods-reporting in studies of biological colouration
 
@@
 ## Introduction
-Since Thayer [-@thayer_concealing_1909], Poulton [-@poulton_colours_1890] and Wallace [-@wallace_natural_1891] generations of biologists have endeavored to explain the mechanisms and functions of animal and plant coloration.
-For these biologists, identifying best practices in measuring color and color pattern and in determining how other species see colour has been a great challenge.
-Over the past 150 years biologists, both independently and in collaboration with physicists and chemists, have developed a range of techniques used to measure and analyse animal and plant coloration including spectrophotometry (also called spectroradiometry, spectroscopy ETC), photography and visual modeling.
+Generations of biologists have endeavored to explain the mechanisms, functions, and evolution of animal and plant coloration [@poulton_colours_1890; @thayer_concealing_1909; @wallace_natural_1891].
+Identifying best practices for measuring and analysing colors and patterns has been an ongoing challenge.
+Over the past 150 years biologists, both independently and in collaboration with physicists and chemists, have developed a range of techniques used to measure and analyse animal and plant coloration including spectrophotometry (variously known as spectroradiometry, spectroscopy, ?), photography and, more recently, visual modeling (CITES).
 
@@
 Before the widespread use of spectrophotometry, researchers assigned categorical color descriptions, arbitrary rankings or scores [e.g.
-@aldrich_ecogeographic_1991; @wheelwright_colors_1985] or matched paint swatches or munsell chip ratings based on their own visual assessment [e.g.
+@aldrich_ecogeographic_1991; @wheelwright_colors_1985] or matched paint swatches or munsell chip ratings based on subjective human assessment [e.g.
 @burtt_analysis_1986; @key_temperature_1954].
-Although when applied carefully they maybe useful, these techniques in general have limited progress in the field.
+Although potentially useful when used carefully, the persistence of these techniques in general has limited progress in the field.
 For example, many researchers designed color assessment methods specifically for their projects [e.g.
 @aldrich_ecogeographic_1991] rendering data by different researchers incomparable.
-This was/is especially problematic when different measurement techniques are used on the same species (e,g, refXXX vs ref XXX).
+This is especially problematic when different measurement techniques are used on the same species (e,g, refXXX vs ref XXX).
 Also, human vision, insensitive to ultraviolet wavelengths, is inherently different to most other animals such as avian and reptilian tetrachromats, and insect trichromats [@bennett_sexual_1994](MORE).
@@
 This approach enables researchers to move beyond purely quantitative comparisons of reflectance spectra and adopt biologically relevant perspectives of signal receivers to test/formulate hypotheses.
+
 Although relatively easy to implement and potentially powerful, sensory models are built on multiple assumptions about the way in which stimuli are visually and cognitively processed that can dramatically shape the results of a given analysis [@lind_avian_2009; @pike_preserving_2012].

### 2014-11-18 docs/ms.Rmd
Integrated Rhi's changes into ms.Rmd

--- 2014-11-11 docs/ms.Rmd
@@
 
-Keywords: #colsci, visual ecology, spectrophotometry, tetrahedral color space, quantum catch, pigment, structural color, thermoregulation, signaling, PAVO, colormetrics, visual modeling, melanism, sexual selection.
+Keywords: #colsci, spectrometry, photography, visual ecology, pigment, structural colour, visual modeling, methodology
 
 ## Abstract
+The colours of plants and animals have long fascinated biological scientists, but it is only recently that quantitative colour measurement techniques have become readily available.
+Spectrometry and digital photography are now relatively accessible and common techniques for the appraisal of biological colour, however, their rapid uptake has not been followed up by clear guidelines on which methodological details are crucial to report.
+We identified a number of details regarding measurement and analysis of colouration which have the potential to affect the outcomes of research.
+We reviewed reporting of relevant methodology in the 2013 literature on biological colour, and have outlined explicit recommendations for future research.
+In providing a constructive criticism of current practices, and identifying areas for improvement, we aim to better the accountability, transparency and reproducibility of colour research in the life sciences.
 
 ## Introduction
-Since Thayer [-@thayer_concealing_1909], Poulton [-@poulton_colours_1890] and Wallace [-@wallace_natural_1891] generations of biologists have endeavored to explain the mechanisms and functions of animal and plant coloration.
-For these biologists, identifying best practice in measuring color and color pattern and in determining how other species see colour has been a great challenge.
-Over the past 150 years biologists, both independently and in collaboration with physicists and chemists, have developed a range of techniques used to measure and analyse animal and plant coloration including spectrophotometry (also called spectroradiometry, spectroscopy ETC), photography and visual modeling.
+Since Thayer [-@thayer_concealing_1909], Poulton [-@poulton_colours_1890] and Wallace [-@wallace_natural_1891] generations of biologists have endeavored to explain the mechanisms and functions of animal and plant colouration.
+For scientists of this discipline, identifying best practice in measuring colour and colour pattern and in determining how other species see colour has been a great challenge.
+Over the past 150 years biologists, both independently and in collaboration with physicists and chemists, have developed a range of techniques used to measure and analyse biological colouration including spectrometry (also called spectroradiometry, spectrophotometry etc.), photography, and visual modeling.
 
-The application of spectrophotometry to questions in ecology and evolution revolutionized the field [@dyck_determination_1966].
-Before the widespread use of spectrophotometry, researchers assigned categorical color descriptions, arbitrary rankings or scores [e.g.
-@aldrich_ecogeographic_1991; @wheelwright_colors_1985] or matched paint swatches or munsell chip ratings based on their own visual assessment [e.g.
+The application of spectrometry to questions in ecology and evolution revolutionized the field [@dyck_determination_1966].
+Before the widespread use of spectrometry, researchers assigned categorical colour descriptions, arbitrary rankings or scores [e.g.
+@aldrich_ecogeographic_1991; @bailey_latitudinal_1978; @wheelwright_colors_1985] or matched paint swatches or munsell chip ratings based on their own visual assessment [e.g.
 @burtt_analysis_1986; @key_temperature_1954].
 Although no doubt applied carefully, these techniques limited progress in the field.
-For example, many researchers designed color assessment methods specifically for their projects [e.g.
-@aldrich_ecogeographic_1991] rendering data by different researchers incomparable.
-This was/is especially problematic when different measurement techniques are used on the same species (e,g, refXXX vs ref XXX).
-Also, human vision, insensitive to ultraviolet wavelengths, is inherently different to most other animals such as avian and reptilian tetrachromats, and insect trichromats [@bennett_sexual_1994](MORE).
-Human perception of color, and variation therein, is not only biologically irrelevant to most systems, but it can lead to erroneous conclusions [@bennett_sexual_1994; @endler_measurement_1990].
+For example, many researchers designed colour assessment methods specifically for their projects [e.g.
+@aldrich_ecogeographic_1991; @bailey_latitudinal_1978] rendering data by different researchers incomparable.
+This is especially problematic when different measurement techniques are used on the same species [e.g.
+the house finch _Carpodacus mexicanus_; @aldrich_origin_1978; @hill_female_1990].
+Also, human vision, insensitive to ultraviolet wavelengths, is inherently different to most other animals such as avian and reptilian tetrachromats, and insect trichromats [@bennett_sexual_1994; @douglas_spectral_2014].
+Human perception of colour is not only biologically irrelevant to most systems, but can lead to erroneous conclusions [@bennett_sexual_1994; @endler_measurement_1990].
 
 Measuring reflectance spectra is now a reasonably inexpensive and convenient technique.
-Spectrophotometry allows the capture of quantitative and arguably repeatable data and has been rapidly taken up by researchers the field.
-However one drawback to spectrophotometry, is the ease with which erroneous data can be captured if attention is not paid to finer details of spectrum capture.
-For example, the angle from which spectra are taken [@santos_strong_2007], consistent calibration of light and dark standards, the distance between light source and object, the expectations of the researcher and the size, shape, glossiness or iridescence of the specimen [@andersson_quantifying_2006; @kemp_ornamental_2008; @meadows_quantifying_2011; @schaefer_birds_2008], can all contribute to potentially skewed data or alter results.
-Experience, guidance and objectivity are necessary to ensure data are consistent and representative of the subject and to ensure repeatability, conditions under which data were collected must be reported exactly.
+Spectrometry allows the capture of quantitative and arguably repeatable data and has been rapidly taken up by researchers the field.
+However one drawback to this methodology, is the ease with which erroneous data can be collected if attention is not paid to finer details of spectrum capture.
+The angle from which spectra are taken [@santos_strong_2007], consistent calibration of light and dark standards, the distance between light source and object, the expectations of the researcher and the size, shape, glossiness or iridescence of the specimen [@andersson_quantifying_2006; @kemp_ornamental_2008; @meadows_quantifying_2011; @schaefer_birds_2008], can all contribute to potentially skewed data or alter results.
+Experience, guidance and objectivity are necessary to ensure data are consistent and representative of the subject.
+To ensure repeatability, the conditions under which data were collected must be reported exactly.
 
-In conjunction with spectrophotometry, digital photography is increasingly being implemented in color quantification.
+In conjunction with spectrometry, digital photography is increasingly being implemented in colour quantification.
 High-resolution cameras are now inexpensive and portable, allow instantaneous sampling of multiple colour patches, and can produce highly repeatable images [@mckay_use_2013].
-The main challenges for gathering and using color data from photographs are ensure correct standardization methods such as linearization are consistent lighting [@stevens_using_2007].
-Crucially, most digital cameras used in color science are consumer products designed for human visual color space and hardware modifications for ultraviolet and infrared imaging are necessary [e.g.
-@chiao_visualization_2009; @mckay_use_2013], making photography complementary, rather than superior to spectrophotometry.
+The main challenge for gathering and using colour data from photographs is ensuring correct standardization methods, such as linearization, and consistent lighting [@stevens_using_2007].
+Crucially, most digital cameras used in colour science are consumer products designed for human visual colour space, and hardware modifications for ultraviolet and infrared imaging are generally necessary [e.g.
+@stevens_colour_2013], making photography complementary, rather than superior to spectrometry.
 
-Expansion in the availability of objective methods for the measurement of color has been matched by advances in theory and analysis.
-In particular, the development of sensory models has produced significant insights into the way in which colorful stimuli are perceived [@chittka_colour_1992; @endler_comparing_2005; @vorobyev_receptor_1998].
-Where the receptor sensitivities of an animal’s visual system are known, the physiological response of that animal’s visual system when viewing a given color can be predicted.
+Expansion in the availability of objective methods for the measurement of colour has been matched by advances in theory and analysis.
+In particular, the development of sensory models has produced significant insights into the way in which colourful stimuli are perceived [@chittka_colour_1992; @endler_comparing_2005; @vorobyev_receptor_1998].
+Where the receptor sensitivities of an animal’s visual system are known, the physiological response of that animal’s visual system when viewing a given colour can be predicted.
 
-Sensory models typically attempt to describe the subjective perception of color or brightness information as a function of an object’s reflectance, the ambient illumination, and a receiver’s sensory system [e.g.
+Sensory models typically attempt to describe the subjective perception of colour or brightness information as a function of an object’s reflectance, the ambient illumination, and a receiver’s sensory system [e.g.
 @endler_comparing_2005; @vorobyev_receptor_1998].
-Furthermore, an animal’s ability to distinguish between reflectance spectra can then be predicted based on the differential responses of a visual system to varying reflectance spectra.
-This approach enables researchers to move beyond purely quantitative comparisons of reflectance spectra and adopt biologically relevant perspectives of signal receivers to test/formulate hypotheses.
+These models enable researchers to move beyond purely quantitative comparisons of reflectance spectra and adopt biologically relevant perspectives of signal receivers to test/formulate hypotheses.
 Although relatively easy to implement and potentially powerful, sensory models are built on multiple assumptions about the way in which stimuli are visually and cognitively processed that can dramatically shape the results of a given analysis [@lind_avian_2009; @pike_preserving_2012].
 For example, Vorobyev et al.’s [-@vorobyev_receptor_1998; @vorobyev_colour_2001] widely used receptor-noise model requires researchers to specify the form of receptor quantum catches (i.e.
-raw versus log-transformed), the use of any color constancy mechanism (e.g.
+raw versus log-transformed), the use of any colour constancy mechanism (e.g.
 von Kries), as well as the photoreceptor signal-to-noise ratios, their relative densities, and the type of noise predominating (i.e.
@@
 Since the repeatability and transparency of science is paramount, it is imperative that we regularly and constructively reflect on our own practices [@alsheikh_public_2011; @sandve_ten_2013].
-The potential for detrimental subjectivity in color measurement and the mathematical opacity of visual models for the average colour scienctist means that, as in any field, is it critical that methods are reported rigorously.
-Methods reporting should ideally have well defined standards such that sufficient detail is provided to allow comprehensive peer review, precise replication, cross-study comparison and data sharing [@alsheikh_public_2011](Ask Rob Lanfear).
-The aims of this paper are to (a) review methods reporting in papers published on colour in 2013 from 20 top tier journals and (b) provide a set of recommendations for best-practice in reporting on the most common methods.
-Our hope is that this instructional provides a useful set of guidelines to improve the reproducibility and transparency of studies on color.
+The potential for subjectivity in colour measurement and the mathematical opacity of visual models for the average colour scientist means that, as in any field, is it critical that methods are reported rigorously.
+Methods reporting should ideally have well defined standards such that sufficient detail is provided to allow comprehensive peer review, precise replication, cross-study comparison and data sharing [@alsheikh_public_2011].
+The aims of this paper were to (a) review methods reporting in papers published on colour in 2013 from 22 top tier journals and (b) provide a set of recommendations for best-practice in reporting on the most common methods.
+Our hope is that this paper will provide a useful set of guidelines to improve the reproducibility and transparency of studies on colour.
 
@@
 
-We developed a list of metrics we deemed critical to be reported for reproducibility of color methods and analysis loosely based on Andersson and Prager [-@andersson_quantifying_2006] and Montgomerie [@montgomerie_analyzing_2006] (Table 1 and 2).
-These included quantitative methods including spectrophotometric and photographic techniques, but not qualitative methods such as the use of Munsell chips or paint swatches.
+We developed a list of metrics we deemed critical to be reported for reproducibility of colour methods (Table 1) and analysis (Table 2), loosely based on Andersson and Prager [-@andersson_quantifying_2006] and Montgomerie [@montgomerie_analyzing_2006].
+These included quantitative methods including spectrophotometric and photographic techniques [as in @kemp_ornamental_2008; @chiao_visualization_2009] but not qualitative methods such as the use of Munsell chips or paint swatches [as in @hogstad_gloger_2009].
 For visual modeling techniques we characterized a set of criteria required for reproducing the analysis (Table 2).
@@
 
-To determine the current state of methods reporting in the field we searched for papers from 2013 in 23 top color science journals (_Animal Behaviour_, _Behavioral Ecology_, _Journal of Experimental Biology_, _Biology Letters_, _Proceedings of the Royal Society B: Biological Sciences_, _Biological Journal of the Linnean Society_, _Ethology_, _The American Naturalist_, _Ecology Letters_, _Journal of Evolutionary Biology_, _American Journal of Botany_, _Evology and Evolution_, _Oikos_, _Ecology_, _Journal of Ecology_, _Behavioral Ecology and Sociobiology_, _Current Zoology_, _Naturwissenschaften_, _New Phytologist_, _PLoS ONE_, _Evolution_, _Functional Ecology_, _Ethology_).
+To determine the current state of methods reporting in the field we searched for papers from 2013 in 22 top colour science journals (_Animal Behaviour_, _Behavioral Ecology_, _Journal of Experimental Biology_, _Biology Letters_, _Proceedings of the Royal Society B: Biological Sciences_, _Biological Journal of the Linnean Society_, _Ethology_, _The American Naturalist_, _Ecology Letters_, _Journal of Evolutionary Biology_, _American Journal of Botany_, _Ecology and Evolution_, _Oikos_, _Ecology_, _Journal of Ecology_, _Behavioral Ecology and Sociobiology_, _Current Zoology_, _Naturwissenschaften_, _New Phytologist_, _PLoS ONE_, _Evolution_, _Functional Ecology_).
 For each of the journals, we used the Boolean search terms 'colour*', 'color*' or 'spectra*' and searched for these in the title or abstract from each of the journal home pages.
@@
 We then tabulated what details of the methods were reported against our criteria.
+The data along with analysis script can be found at: https://github.com/daniel1noble/colsci_rep (TODO: GIVE IT A DOI?).
+We have kept the papers used in our dataset annoymous to avoid accusing specific colleagues of inadvertant reporting.
+We feel that the issue of reporting is something attributable to the ambiguity and lack of guidelines in our field and not something specific to individual research groups.
 
 Our search returned a total of 216 papers, however, only `r nrow(coldat)` met our criteria and were included in the final analysis.
-Overall, `r specs+both` studies used a spectrophotometer to measure color, while `r cam+both` used a camera.
-Of these studies, `r both` quantified color using both a spectrophotometer and a camera.
-Reporting of the make and model of the spectrophotometer and camera was high (Figure 1 - Specs = `r round(specModel, digits = 2)`% and Cameras = `r round(camModel, digits = 2)`%), however, only `r round(CamLight, digits = 2)`%  and `r round(specLight, digits = 2)`% of camera and spec studies reported a light source.
-In addition, the number of reflectance spectra and pixels averaged were only reported in `r round(specAVG, digits = 2)`% and `r round(camAVG, digits = 2)`% of the studies, respectively.
-Surprisingly, of the spec studies, dark standards were only reported in `r round(drk_std, digits = 2)`% of the studies, while white standards were reported in `r round(white_stdspec, digits = 2)`%.
-Studies quantifying spectral reflectance also often did not report the integration time of the spectrophotmeter (`r round(intTime, digits = 2)`%) or the angle (`r round(angle, digits = 2)`%) or the distance (`r round(dist, digits = 2)`%) the probe was from the color patch.
+Overall, `r specs+both` studies used a spectrometer (henceforth "spec") to measure colour, while `r cam+both` used a camera.
+Of these studies, `r both` quantified colour using both a spec and a camera.
+Reporting of the make and model of the spec and camera was high (Figure 1, spectrometers = `r round(specModel, digits = 2)`% and cameras = `r round(camModel, digits = 2)`%), however, only `r round(CamLight, digits = 2)`%  and `r round(specLight, digits = 2)`% of camera and spec studies reported a light source.
+In addition, the number of reflectance spectra and pixels averaged were only reported in `r round(specAVG, digits = 2)`% and `r round(camAVG, digits = 2)`% of studies, respectively.
+Surprisingly, dark standards were only reported in `r round(drk_std, digits = 2)`% of the spec studies, while white standards were reported in `r round(white_stdspec, digits = 2)`%.
+Studies quantifying spectral reflectance often did not report the integration time of the spectrophotmeter (`r round(intTime, digits = 2)`%), or the angle (`r round(angle, digits = 2)`%) or distance (`r round(dist, digits = 2)`%) the probe was from the colour patch.
 
-Analysis varied across studies but, overall `r analysisType[1]` analyzed their data by generating colormetric (brightness, hue, chroma) variables whereas `r analysisType[4]` used visual models, `r analysisType[2]` used multiple analysis types (e.g.
-visual model and colourmetric) and `r analysisType[3]` did other forms of analysis (e.g.
-RGB, PCA's, GLMMs etc.).
-Of the studies quantifying colour using spectral reflectance curves `r analysisTypeSpec[1]` used colormetric measures, `r analysisTypeSpec[2]` used multiple analyses, `r analysisTypeSpec[4]` used visual models and `r analysisTypeSpec[3]` used other forms of analysis.
-Studies using colormetric measurements defined their measure of brightness, hue and/or chroma in `r round(col_def, digits = 2)`% of the studies.
+Analysis varied across studies.
+Overall `r analysisType[1]` papers analyzed  data by generating colourimetric variables [i.e.
+'spectral' measures of hue, saturation, and/or brightness; @montgomerie_analyzing_2006] whereas `r analysisType[4]` used visual models, `r analysisType[2]` used multiple analysis types, and `r analysisType[3]` used other forms of analysis (e.g.
+purely statistical analyses, such as PCA's on reflectance spectra).
+Of the studies quantifying colour using spectral reflectance curves `r analysisTypeSpec[1]` used colourimetric measures, `r analysisTypeSpec[2]` used multiple analyses, `r analysisTypeSpec[4]` used visual models and `r analysisTypeSpec[3]` used other forms of analysis.
+Studies using colourimetric measurements defined their measure of brightness, hue and/or chroma in `r round(col_def, digits = 2)`% of the studies.
 
-Most studies using visual models analysed their data using receptor noise models (`r vismod_type[3]` of `r sum(vismod_type)` studies) while others used hexagon (`r vismod_type[1]` studies) or tcs (`r vismod_type[4]` studies).
-Case specific models were used in `r vismod_type[2]` cases.
-Overall, `r round(prop_irrad, digits =2)`% reported the type of irradiance, `r round(prop_spvismod, digits = 2)`% the species used in their visual model and `r round(propvis_mod_param, digits = 2)`% made explicit mention of the parameters they used (i.e.
-cone densities, neural vs quantum catchs etc.) in their methods.
+Most studies using visual models analysed their data using receptor-noise models (`r vismod_type[3]` of `r sum(vismod_type)` studies), while others used a hexagonal (`r vismod_type[1]` studies), or tetrahedral colour space (`r vismod_type[4]` studies) approach.
+Case specific models were used in `r vismod_type[2]` papers.
+Of visual modelling papers, `r round(prop_irrad, digits =2)`% reported the type of irradiance when necessary, `r round(prop_spvismod, digits = 2)`% the species modelled as a receiver, and `r round(propvis_mod_param, digits = 2)`% made explicit and complete mention of the parameters they used (e.g.
+cone densities, signal-to-noise ratios).
 The background used in visual models was reported `r round(prop_bkg, digits = 2)`% of the time.
-The type of visual noise authors assumed was reported in `r round(prop_noise, digits = 2)`% of the studies.
-In addition, explicit mention of the type of quantum catche (`r round(prop_qcatch, digits = 2)`% of studies) or whether photoreceptor adaptation (`r round(prop_adap, digits = 2)`%) was accounted for in visual models was not common.
+The type of receptor noise modelled was reported in only `r round(prop_noise, digits = 2)`% of the studies.
+Explicit mention of the type of quantum catch (`r round(prop_qcatch, digits = 2)`% of studies), or whether photoreceptor adaptation (`r round(prop_adap, digits = 2)`%) was modelled, was not common.
 
-While some of the values reported above seem alarming, it is important to note that papers often made reference to previous work for various methods (`r round(ref_stud, digits = 2)`%) and so these figures are slightly liberal given that referenced work may have comprehensively covered some of these criteria.
-Despite this, many of the studies that did not reference previous work were lacking in their reporting of many aspects of our criteria suggesting the need for more comprehensive reporting of data collection and analysis methods within studies.
+While some of the values reported above seem alarming, it is important to note that papers often made reference to previous work for various methods (`r round(ref_stud, digits = 2)`%), so these figures are slightly liberal given that referenced work may have comprehensively covered some of these criteria.
+Despite this, many of the studies that did not reference previous work were lacking in their reporting of many aspects of our criteria suggesting the need for more comprehensive reporting of data collection and analysis methods within such studies.
 
@@
 
-Our review of recent colour literarure highlights some major reporting gaps that will hinder the reproducibility of studies quantifying colour signals.
-While we recognize that journals have space limitations and that work may reference previous methological and analytical details, we strongly beleive that studies can do a much better job reporting details which are essential to the transparency and reproducibility of their work.
-When possible, we also recommend that authors justify data capture and analytical approaches - something few studies we investiagted expliclty stated.
-While all criteria are important some deserve special attanetion.
-In particular we found that colour scientists could do a much better job reporting light sources, number of pixel and spectra averaged used for analysis and the type of standards used.
-More improtantly, the distance and angle should be standard reporting for all studies using spectrophotomtry.
-These all have important consequences on downstream analysis and the reproducibility of colour signal quantification.
-As we all know, a picture can say 1000 words and we recommend the incoperation of spectral reflectance curves were possible as it provides insight on the colour signal, which is visually appealing and easy to understand by most readers.
-Surprisingly, we found that only `r round(fig, digits = 2)`% of studies using spectral reflectance data had an explicit figure in their paper showing these curves.
+Our review of recent colour literature highlights some important reporting gaps that will hinder the reproducibility of studies of biological colouration.
+While we recognize that journals have space limitations and that work may reference previous methological and analytical details, we strongly believe that researchers can do a better job of reporting details which are essential to the transparency and reproducibility of their work.
+While all criteria are important, some deserve special attention.
+In particular, we found that colour scientists could do a better job reporting light sources, the numbers of pixels or spectra averaged, and the type of standards used.
+More importantly, the probe distance and angle should be standard reporting for all studies using spectrometry.
 Visual models were regularly used to analyse spectral data, something we welcome.
-However, there were many important aspects of the analysis that were seldom reported, inclding the type of quantum catch used, whether photoreceptor adaptation was accounted for and the type of neural noise used in receptor noise models.
+However, there were many important aspects of the analysis that were seldom reported, including the type of quantum catch used, whether photoreceptor adaptation was modelled and the type of neural noise used in receptor-noise models.
 Photoreceptor densities should also be reported where possible even if using standard avian photorecptor densities.
+These all have important consequences on downstream analysis and the reproducibility of colour signal quantification [e.g.
+@kemp_ornamental_2008; @santos_comparison_2007].
+It is said that a picture is worth 1000 words, and we recommend the presentation of spectral reflectance curves where possible; these figures are visually appealing and allow for the rapid assesment of the nature and quality of the data, independent of downstream processing.
+Surprisingly, we found that only `r round(fig, digits = 2)`% of studies using spectral reflectance data had a figure showing these curves.
+When possible, we also recommend that authors justify data capture and analytical approaches - something few studies we investigated explicitly outlined.
 
-In conclusion, our objective in this paper was not to point to single studies and lay blame, but to highlight an important gap in reporting that can influence the reproducibility of colour research and cross taxa comparisons.
-Indeed, even in our own work [@umbers_bright_2013] we have failed to report important details, which in resptrospec can lead to confusion about how such data were collected and analysed.
-Nontheless, our critisim of reporting is meant to be constructive and is not to say that studies lacking improtant details are not excellent works, but it is only through critical appraisal and colleageal discussion that we can identify these gaps and constructively contribute to the growth of our field.
-Such assessments are particlularly important in an age where accountability, transparency and reproducibility of are utmost importance.
-We hope our critical evaluation of recent litearture has highlighted a need for more careful reporting of how colour signals have been measured and analysed.
-We beleive that part of the issue is simply that researchers seldom have clear guidelines to follow when undertaking research and miss what might be important aspects of their methods and analysis that need to be reported.
-We therefore hope that our paper will help estbalish more uniform reporting guidelines in the literature so that we can continue an exciting era of colour signal research in the life sciences.
+Our objective in this paper was not to point out single studies, but to highlight an important gap in reporting that can influence the reproducibility of colour research and cross-study comparisons.
+By no means do we wish to say that studies lacking important details are not of significant value.
+Indeed, even in our own work [e.g.
+@umbers_bright_2013] we have failed to report important details, which in retrospect can lead to confusion about how such data were collected and analysed.
+This review of reporting is aimed at providing constructive criticism; it is only through critical appraisal and collegial discussion that we can identify such gaps as areas for improvement, and constructively contribute to the growth of our field.
+Such assessments are particularly important in an age where accountability, transparency and reproducibility are of utmost importance.
+
+We believe that part of the issue at hand is simply that many researchers seldom have clear guidelines to follow when undertaking research, and may overlook outlining important aspects of methodology and analysis.We hope our critical evaluation of recent literature has highlighted a need for more careful description of how colour signals have been measured and analyzed.
+We are optimistic that this may help establish uniform reporting guidelines for publication, and to continue the advance of an exciting era in colour research.
 

### 2014-11-18 docs/ms.Rmd
Integrated Kate's edits into ms.Rmd

@@
 
+Article type: Commentary
+
+Word count:
+
+Number of figures:
+
+Number of tables: 2
+
+Data archiving:
+
+\newpage
+
 ## Abstract
@@
 
+\newpage
+
 ## Introduction
-Since Thayer [-@thayer_concealing_1909], Poulton [-@poulton_colours_1890] and Wallace [-@wallace_natural_1891] generations of biologists have endeavored to explain the mechanisms and functions of animal and plant colouration.
-For scientists of this discipline, identifying best practice in measuring colour and colour pattern and in determining how other species see colour has been a great challenge.
-Over the past 150 years biologists, both independently and in collaboration with physicists and chemists, have developed a range of techniques used to measure and analyse biological colouration including spectrometry (also called spectroradiometry, spectrophotometry etc.), photography, and visual modeling.
+Generations of biologists have endeavored to explain the mechanisms and functions of animal and plant colouration [@thayer_concealing_1909; @poulton_colours_1890; @wallace_natural_1891].
+For scientists of this discipline, identifying best practice in measuring colour and colour pattern has been a great challenge.
+For these biologists, identifying best practice in measuring color and color pattern has been a great challenge.
+Over the past 150 years, biologists have developed and adapted techniques that are now widespread, including spectrometry, photography, and visual modelling.
 
-The application of spectrometry to questions in ecology and evolution revolutionized the field [@dyck_determination_1966].
-Before the widespread use of spectrometry, researchers assigned categorical colour descriptions, arbitrary rankings or scores [e.g.
+The application of spectrometry to biological questions revolutionized the field [@dyck_determination_1966].
+Before its widespread use, researchers assigned categorical colour descriptions, arbitrary rankings or scores [e.g.
 @aldrich_ecogeographic_1991; @bailey_latitudinal_1978; @wheelwright_colors_1985] or matched paint swatches or munsell chip ratings based on their own visual assessment [e.g.
 @burtt_analysis_1986; @key_temperature_1954].
-Although no doubt applied carefully, these techniques limited progress in the field.
+Although no doubt applied carefully, these techniques limited progress.
 For example, many researchers designed colour assessment methods specifically for their projects [e.g.
-@aldrich_ecogeographic_1991; @bailey_latitudinal_1978] rendering data by different researchers incomparable.
+@aldrich_ecogeographic_1991; @bailey_latitudinal_1978] rendering data by incomparable across studies.
 This is especially problematic when different measurement techniques are used on the same species [e.g.
 the house finch _Carpodacus mexicanus_; @aldrich_origin_1978; @hill_female_1990].
-Also, human vision, insensitive to ultraviolet wavelengths, is inherently different to most other animals such as avian and reptilian tetrachromats, and insect trichromats [@bennett_sexual_1994; @douglas_spectral_2014].
-Human perception of colour is not only biologically irrelevant to most systems, but can lead to erroneous conclusions [@bennett_sexual_1994; @endler_measurement_1990].
+Likewise, because human vision is inherently different to most other animals [@bennett_sexual_1994; @douglas_spectral_2014], its use can lead to erroneous conclusions [@bennett_sexual_1994; @endler_measurement_1990].
 
-Measuring reflectance spectra is now a reasonably inexpensive and convenient technique.
+Measuring reflectance spectra via spectrometry is now a reasonably inexpensive and convenient technique.
 Spectrometry allows the capture of quantitative and arguably repeatable data and has been rapidly taken up by researchers the field.
 However one drawback to this methodology, is the ease with which erroneous data can be collected if attention is not paid to finer details of spectrum capture.
-The angle from which spectra are taken [@santos_strong_2007], consistent calibration of light and dark standards, the distance between light source and object, the expectations of the researcher and the size, shape, glossiness or iridescence of the specimen [@andersson_quantifying_2006; @kemp_ornamental_2008; @meadows_quantifying_2011; @schaefer_birds_2008], can all contribute to potentially skewed data or alter results.
-Experience, guidance and objectivity are necessary to ensure data are consistent and representative of the subject.
-To ensure repeatability, the conditions under which data were collected must be reported exactly.
+The angle from which spectra are taken [@santos_strong_2007], consistent calibration of light and dark standards, the distance between light source and surface, and the size, shape, glossiness or iridescence of the specimen can all potentially bias results [@andersson_quantifying_2006; @kemp_ornamental_2008; @meadows_quantifying_2011; @schaefer_birds_2008].
+Experience, guidance and objectivity (e.g.
+double-blind experimental design) are necessary to ensure data are consistent and representative of the subject.
+Crucially, the conditions under which data were collected must be reported exactly to ensure repeatability.
 
-In conjunction with spectrometry, digital photography is increasingly being implemented in colour quantification.
-High-resolution cameras are now inexpensive and portable, allow instantaneous sampling of multiple colour patches, and can produce highly repeatable images [@mckay_use_2013].
-The main challenge for gathering and using colour data from photographs is ensuring correct standardization methods, such as linearization, and consistent lighting [@stevens_using_2007].
-Crucially, most digital cameras used in colour science are consumer products designed for human visual colour space, and hardware modifications for ultraviolet and infrared imaging are generally necessary [e.g.
-@stevens_colour_2013], making photography complementary, rather than superior to spectrometry.
+Complementary to spectrometry, digital photography is increasingly being used to quantify biological colouration.
+High-resolution cameras are inexpensive,  and allow instantaneous, repeatable sampling of multiple colour patches [@mckay_use_2013].
+The main challenge for gathering colour data from photographs are in ensuring correct standardization - such as linearization - and consistent lighting [@stevens_using_2007].
+Most digital cameras used for the measurement of biological colouration are consumer products designed for the human visual system.
+Thus hardware modifications for ultraviolet and infrared imaging are generally necessary [e.g.
+@stevens_colour_2013].
 
@@
 @endler_comparing_2005; @vorobyev_receptor_1998].
-These models enable researchers to move beyond purely quantitative comparisons of reflectance spectra and adopt biologically relevant perspectives of signal receivers to test/formulate hypotheses.
-Although relatively easy to implement and potentially powerful, sensory models are built on multiple assumptions about the way in which stimuli are visually and cognitively processed that can dramatically shape the results of a given analysis [@lind_avian_2009; @pike_preserving_2012].
+These models enable researchers to move beyond purely quantitative comparisons of reflectance spectra and adopt biologically relevant perspectives of signal receivers to formulate and test hypotheses.
+
+Although relatively easy to implement and potentially powerful, sensory models are built on multiple assumptions about the way in which stimuli are visually and perceptually processed, and these can dramatically shape the results of a given analysis [@lind_avian_2009; @pike_preserving_2012].
 For example, Vorobyev et al.’s [-@vorobyev_receptor_1998; @vorobyev_colour_2001] widely used receptor-noise model requires researchers to specify the form of receptor quantum catches (i.e.
@@
 neural and/or receptor noise).
-These data are only available for a small number of model species, such as honeybees, and so assumptions and extrapolations are necessary when models are applied to related species.
+These data are only available for a small number of species, such as humans and honeybees, and so assumptions and extrapolations are necessary when models are applied to other species.
 Unless the biological rationale underlying every parameter choice is clearly understood, justified, and reported, such analyses may become ineffective, especially when comparing results between studies [@pike_preserving_2012].
@@
 Since the repeatability and transparency of science is paramount, it is imperative that we regularly and constructively reflect on our own practices [@alsheikh_public_2011; @sandve_ten_2013].
-The potential for subjectivity in colour measurement and the mathematical opacity of visual models for the average colour scientist means that, as in any field, is it critical that methods are reported rigorously.
+The potential for subjectivity in colour measurement and the mathematical opacity of visual models for the average biologist means that, as in any field, is it critical that methods are reported rigorously.
 Methods reporting should ideally have well defined standards such that sufficient detail is provided to allow comprehensive peer review, precise replication, cross-study comparison and data sharing [@alsheikh_public_2011].

### 2014-11-18 docs/ms.Rmd
Slight spelling and grammer changes and slight rewording of various places. Nothing big

@@
 Generations of biologists have endeavored to explain the mechanisms and functions of animal and plant colouration [@thayer_concealing_1909; @poulton_colours_1890; @wallace_natural_1891].
-For scientists of this discipline, identifying best practice in measuring colour and colour pattern has been a great challenge.
-For these biologists, identifying best practice in measuring color and color pattern has been a great challenge.
+For these biologists, identifying best practices in measuring colour and colour pattern has been a great challenge.
 Over the past 150 years, biologists have developed and adapted techniques that are now widespread, including spectrometry, photography, and visual modelling.
@@
 For example, many researchers designed colour assessment methods specifically for their projects [e.g.
-@aldrich_ecogeographic_1991; @bailey_latitudinal_1978] rendering data by incomparable across studies.
+@aldrich_ecogeographic_1991; @bailey_latitudinal_1978] rendering data incomparable across studies.
 This is especially problematic when different measurement techniques are used on the same species [e.g.
@@
 
-Although relatively easy to implement and potentially powerful, sensory models are built on multiple assumptions about the way in which stimuli are visually and perceptually processed, and these can dramatically shape the results of a given analysis [@lind_avian_2009; @pike_preserving_2012].
+Although relatively easy to implement, and potentially powerful, sensory models are built on multiple assumptions about the way in which stimuli are visually and perceptually processed and these can dramatically shape the results of a given analysis [@lind_avian_2009; @pike_preserving_2012].
 For example, Vorobyev et al.’s [-@vorobyev_receptor_1998; @vorobyev_colour_2001] widely used receptor-noise model requires researchers to specify the form of receptor quantum catches (i.e.
@@
 In particular, we found that colour scientists could do a better job reporting light sources, the numbers of pixels or spectra averaged, and the type of standards used.
-More importantly, the probe distance and angle should be standard reporting for all studies using spectrometry.
+More importantly, the probe distance and angle should be standard reporting for all studies using spectrophometry.
 Visual models were regularly used to analyse spectral data, something we welcome.
@@
 
-Our objective in this paper was not to point out single studies, but to highlight an important gap in reporting that can influence the reproducibility of colour research and cross-study comparisons.
+Our objective in this paper was not to point out negative aspects of single studies, but to highlight an important gap in reporting that can influence the reproducibility of colour research and cross-study comparisons.
 By no means do we wish to say that studies lacking important details are not of significant value.
@@
 
-We believe that part of the issue at hand is simply that many researchers seldom have clear guidelines to follow when undertaking research, and may overlook outlining important aspects of methodology and analysis.We hope our critical evaluation of recent literature has highlighted a need for more careful description of how colour signals have been measured and analyzed.
+We believe that part of the issue at hand is simply that many researchers seldom have clear guidelines to follow when undertaking research, and may overlook outlining important aspects of methodology and analysis.
+We hope our critical evaluation of recent literature has highlighted a need for more careful description of how colour signals have been measured and analyzed.
 We are optimistic that this may help establish uniform reporting guidelines for publication, and to continue the advance of an exciting era in colour research.

### 2014-11-18 docs/ms.Rmd
Incoperated some of james comments. But ms has changed quite a bit from his version and so some were not relevant

@@
 However one drawback to this methodology, is the ease with which erroneous data can be collected if attention is not paid to finer details of spectrum capture.
-The angle from which spectra are taken [@santos_strong_2007], consistent calibration of light and dark standards, the distance between light source and surface, and the size, shape, glossiness or iridescence of the specimen can all potentially bias results [@andersson_quantifying_2006; @kemp_ornamental_2008; @meadows_quantifying_2011; @schaefer_birds_2008].
-Experience, guidance and objectivity (e.g.
-double-blind experimental design) are necessary to ensure data are consistent and representative of the subject.
-Crucially, the conditions under which data were collected must be reported exactly to ensure repeatability.
+The angle from which spectra are taken [@santos_strong_2007], consistent calibration of light and dark standards, the distance between light source and surface, and the size, shape, glossiness or iridescence of the specimen can all contribute to variation in data and potentially bias results [@andersson_quantifying_2006; @kemp_ornamental_2008; @meadows_quantifying_2011; @schaefer_birds_2008].
+Experience, guidance, consistency and objectivity (e.g.
+double-blind experimental design) are necessary to ensure data are repeatable and representative of the subject.
+Crucially, to ensure repeatability across studies, the conditions under which data were collected must be reported exactly.
 
@@
 High-resolution cameras are inexpensive,  and allow instantaneous, repeatable sampling of multiple colour patches [@mckay_use_2013].
-The main challenge for gathering colour data from photographs are in ensuring correct standardization - such as linearization - and consistent lighting [@stevens_using_2007].
+The main challenge for gathering colour data from photographs involve ensuring standardization, such as linearization and consistent lighting [@stevens_using_2007].
 Most digital cameras used for the measurement of biological colouration are consumer products designed for the human visual system.
@@
 The aims of this paper were to (a) review methods reporting in papers published on colour in 2013 from 22 top tier journals and (b) provide a set of recommendations for best-practice in reporting on the most common methods.
+Our review specifically focuses on papers that quantify the external coloration of a plant or animal.
+These may be purely descriptive quantifications of color, or they may be incorporated into models of animal vision.
 Our hope is that this paper will provide a useful set of guidelines to improve the reproducibility and transparency of studies on colour.
@@
 For visual modeling techniques we characterized a set of criteria required for reproducing the analysis (Table 2).
+Much of these criteria specifically targeted the necessary parameters required to be explicitly defined by the user during analysis.
+These are essential if such analyses are to be reproduced.
 
@@
 We read the titles and abstracts of each paper and included only those papers that explicitly used either a spectrophotometer or camera to quantify colour signals.
-We excluded papers that were: 1) review papers; 2) methodological papers; 3) papers quantifying colour pattern; and 4) papers taking microspectrophotometric measurments of retinal absorbance.
+We excluded papers that were: 1) review papers; 2) methodological papers; 3) papers quantifying colour pattern rather than characteristics of each individual colour patch; and 4) papers taking microspectrophotometric measurments of retinal absorbance.
 To review the large number of papers returned from the searches, journals were haphazardly divided up between the authors.
@@
 Surprisingly, dark standards were only reported in `r round(drk_std, digits = 2)`% of the spec studies, while white standards were reported in `r round(white_stdspec, digits = 2)`%.
-Studies quantifying spectral reflectance often did not report the integration time of the spectrophotmeter (`r round(intTime, digits = 2)`%), or the angle (`r round(angle, digits = 2)`%) or distance (`r round(dist, digits = 2)`%) the probe was from the colour patch.
+Studies quantifying spectral reflectance often did not report the integration time of the spectrophotmeter (`r round(intTime, digits = 2)`%), or the angle of the optical probe relative to the surface (`r round(angle, digits = 2)`%) or distance (`r round(dist, digits = 2)`%) the probe was held from the measurement surface.
 
@@
 
-Most studies using visual models analysed their data using receptor-noise models (`r vismod_type[3]` of `r sum(vismod_type)` studies), while others used a hexagonal (`r vismod_type[1]` studies), or tetrahedral colour space (`r vismod_type[4]` studies) approach.
+Most studies using visual models analysed their data using receptor-noise models (`r vismod_type[3]` of `r sum(vismod_type)` studies), while others used the colour hexagon (`r vismod_type[1]` studies), or tetrahedral colour space (`r vismod_type[4]` studies) approach.
 Case specific models were used in `r vismod_type[2]` papers.
@@
 Our review of recent colour literature highlights some important reporting gaps that will hinder the reproducibility of studies of biological colouration.
-While we recognize that journals have space limitations and that work may reference previous methological and analytical details, we strongly believe that researchers can do a better job of reporting details which are essential to the transparency and reproducibility of their work.
+While we recognize that journals have space limitations and that work may reference previous methological and analytical details, we strongly believe that researchers can do a better job of reporting details that are essential to both the outcome of their results and reproducibility of their work.
 While all criteria are important, some deserve special attention.
-In particular, we found that colour scientists could do a better job reporting light sources, the numbers of pixels or spectra averaged, and the type of standards used.
-More importantly, the probe distance and angle should be standard reporting for all studies using spectrophometry.
+In particular, we found that colour scientists could do a better job reporting light sources, the numbers of pixels or spectra averaged, and the type of standards used (i.e.
+white and black reference samples).
+More importantly, the distance and angle of the optical probe relative to the measurment surface should be standard reporting for all studies using spectrophometry.
 Visual models were regularly used to analyse spectral data, something we welcome.
@@
 @kemp_ornamental_2008; @santos_comparison_2007].
-It is said that a picture is worth 1000 words, and we recommend the presentation of spectral reflectance curves where possible; these figures are visually appealing and allow for the rapid assesment of the nature and quality of the data, independent of downstream processing.
+In addition, we recommend the presentation of spectral reflectance curves where possible; these figures are visually appealing and allow for the rapid assesment of the nature and quality of the data, independent of downstream processing.
 Surprisingly, we found that only `r round(fig, digits = 2)`% of studies using spectral reflectance data had a figure showing these curves.
@@
 
-Our objective in this paper was not to point out negative aspects of single studies, but to highlight an important gap in reporting that can influence the reproducibility of colour research and cross-study comparisons.
+Our objective in this paper was not to highlight shortcomings in single studies, but to bring attantion to common oversights in reporting that can influence the reproducibility of colour research and cross-study comparisons.
 By no means do we wish to say that studies lacking important details are not of significant value.

### 2014-11-20 R/analysis.r
Minor changes to reduce package installation

@@ -9,7 +9,7 @@
 rm(list=ls())
 
 ## Packages. Coded to simplify multiple package installation and loading.
-packages <- c("car", "ggplot2", "plyr")
+packages <- c("car", "ggplot2", "plyr", "knitr", "markdown")
 install.packages(packages)
 sapply(packages, function(x) library(x, character.only = TRUE, verbose = FALSE))
 

### 2014-11-20 R/knit_doc.r

@@ -1,11 +1,5 @@
-
-packages <- c("knitr", "markdown")
-install.packages(packages)
-sapply(packages, function(x) library(x, character.only = TRUE))
-
 #Read necessary functions and files. There are two functions knit2pdf and knit2html which will produce documents directly, but there are problems with pdf and also problems rendering the markdown file when complete. Best to just use knit first then take md file and convert in ternimal. 
 
-rm(list=ls())
 setwd("./docs")
 knit("ms.Rmd")
 setwd("..")

### 2014-11-22 docs/ms.Rmd
Integrated Zurek's edits, but it's changed a fair bit since then

@@
 We read the titles and abstracts of each paper and included only those papers that explicitly used either a spectrophotometer or camera to quantify colour signals.
-We excluded papers that were: 1) review papers; 2) methodological papers; 3) papers quantifying colour pattern rather than characteristics of each individual colour patch; and 4) papers taking microspectrophotometric measurments of retinal absorbance.
+We excluded papers that were: 1) review papers; 2) methodological papers; 3) papers quantifying colour patterns rather than the chromatic properties of a colour patch; and 4) papers taking microspectrophotometric measurments of retinal absorbance.
 To review the large number of papers returned from the searches, journals were haphazardly divided up between the authors.
@@
 
-While some of the values reported above seem alarming, it is important to note that papers often made reference to previous work for various methods (`r round(ref_stud, digits = 2)`%), so these figures are slightly liberal given that referenced work may have comprehensively covered some of these criteria.
+While some of the percentages reported above seem alarming, it is important to note that papers often made reference to previous work for various methods (`r round(ref_stud, digits = 2)`%), so these figures are slightly liberal given that referenced work may have comprehensively covered some of these criteria.
 Despite this, many of the studies that did not reference previous work were lacking in their reporting of many aspects of our criteria suggesting the need for more comprehensive reporting of data collection and analysis methods within such studies.

### 2014-11-22 docs/ms.Rmd
Integrate Kate's latest edits to ms.rmd

--- 2014-11-18 docs/ms.Rmd
@@
 ## Abstract
-The colours of plants and animals have long fascinated biological scientists, but it is only recently that quantitative colour measurement techniques have become readily available.
-Spectrometry and digital photography are now relatively accessible and common techniques for the appraisal of biological colour, however, their rapid uptake has not been followed up by clear guidelines on which methodological details are crucial to report.
-We identified a number of details regarding measurement and analysis of colouration which have the potential to affect the outcomes of research.
-We reviewed reporting of relevant methodology in the 2013 literature on biological colour, and have outlined explicit recommendations for future research.
-In providing a constructive criticism of current practices, and identifying areas for improvement, we aim to better the accountability, transparency and reproducibility of colour research in the life sciences.
+Over the last 50 years or so, our enhanced capacity for quantitative colour measurement has accelerated understanding of the evolution of biological colouration.
+However, while spectrometry and digital photography are inexpensive, accessible and efficient, details of exactly how data are collected and analysed are not reported consistently in the literature.
+Similarly, estimates of animals’ visual capacity are widely applied to colour data, but details on assumptions and settings in analyses are not reported completely.
+To quantify the variability of methods reporting in studies of colouration, we reviewed papers published in 2013 in 22 leading journals.
+Using the search terms colour* OR color* OR spectra* in the title or abstract, we identified 60 papers in which colour had been measured and analysed.
+We extracted the methods and noted the details included.
+We found a great deal of variability in the level of detail in which colour methods is reported.
+In response we have outlined explicit recommendations on the important methodological details required to ensure studies are transparent and repeatable.
+We hope that this will help to standardize methods reporting in the field of biological colouration paving the way for direct comparison across studies.
 
@@
 Generations of biologists have endeavored to explain the mechanisms and functions of animal and plant colouration [@thayer_concealing_1909; @poulton_colours_1890; @wallace_natural_1891].
-For these biologists, identifying best practices in measuring colour and colour pattern has been a great challenge.
-Over the past 150 years, biologists have developed and adapted techniques that are now widespread, including spectrometry, photography, and visual modelling.
+For these biologists, identifying best practices in measuring colour has been a great challenge.
+Over the past 50 years, biologists have developed and adapted techniques for the objective measurement of colouration that are now widespread.
+These include spectrometry, photography, and visual modelling.
 
-The application of spectrometry to biological questions revolutionized the field [@dyck_determination_1966].
+The application of spectrometry to questions of biological colouration revolutionized the field [@dyck_determination_1966].
 Before its widespread use, researchers assigned categorical colour descriptions, arbitrary rankings or scores [e.g.
@@
 Although no doubt applied carefully, these techniques limited progress.
-For example, many researchers designed colour assessment methods specifically for their projects [e.g.
-@aldrich_ecogeographic_1991; @bailey_latitudinal_1978] rendering data incomparable across studies.
+For example, many researchers designed colour assessment methods specifically for their projects, rendering data incomparable across studies [e.g.
+@aldrich_ecogeographic_1991; @bailey_latitudinal_1978].
 This is especially problematic when different measurement techniques are used on the same species [e.g.
-the house finch _Carpodacus mexicanus_; @aldrich_origin_1978; @hill_female_1990].
-Likewise, because human vision is inherently different to most other animals [@bennett_sexual_1994; @douglas_spectral_2014], its use can lead to erroneous conclusions [@bennett_sexual_1994; @endler_measurement_1990].
+house finch _Carpodacus mexicanus_; @aldrich_origin_1978; @hill_female_1990].
+Likewise, because human vision is inherently different to most other animals [@bennett_sexual_1994; @douglas_spectral_2014], its use has lead to erroneous conclusions [@bennett_sexual_1994; @endler_measurement_1990].
 
@@
 
-Complementary to spectrometry, digital photography is increasingly being used to quantify biological colouration.
+Complementary to spectrometry, digital photography is increasingly being used to quantify biological colouration [@stevens_using_2007].
 High-resolution cameras are inexpensive,  and allow instantaneous, repeatable sampling of multiple colour patches [@mckay_use_2013].
 The main challenge for gathering colour data from photographs involve ensuring standardization, such as linearization and consistent lighting [@stevens_using_2007].
-Most digital cameras used for the measurement of biological colouration are consumer products designed for the human visual system.
+Also, most digital cameras used for the measurement of biological colouration are consumer products designed for the human visual system.
 Thus hardware modifications for ultraviolet and infrared imaging are generally necessary [e.g.
@@
 Expansion in the availability of objective methods for the measurement of colour has been matched by advances in theory and analysis.
-In particular, the development of sensory models has produced significant insights into the way in which colourful stimuli are perceived [@chittka_colour_1992; @endler_comparing_2005; @vorobyev_receptor_1998].
-Where the receptor sensitivities of an animal’s visual system are known, the physiological response of that animal’s visual system when viewing a given colour can be predicted.
-
+In particular, the development of sensory models has produced significant insights into how colourful stimuli are perceived [@chittka_colour_1992; @endler_comparing_2005; @vorobyev_receptor_1998].
 Sensory models typically attempt to describe the subjective perception of colour or brightness information as a function of an object’s reflectance, the ambient illumination, and a receiver’s sensory system [e.g.
@@
 Although relatively easy to implement, and potentially powerful, sensory models are built on multiple assumptions about the way in which stimuli are visually and perceptually processed and these can dramatically shape the results of a given analysis [@lind_avian_2009; @pike_preserving_2012].
-For example, Vorobyev et al.’s [-@vorobyev_receptor_1998; @vorobyev_colour_2001] widely used receptor-noise model requires researchers to specify the form of receptor quantum catches (i.e.
+For example, Vorobyev et al.’s [-@vorobyev_receptor_1998; -@vorobyev_colour_2001] widely used receptor-noise model requires researchers to specify the form of receptor quantum catches (i.e.
 raw versus log-transformed), the use of any colour constancy mechanism (e.g.
@@
 
-Since the repeatability and transparency of science is paramount, it is imperative that we regularly and constructively reflect on our own practices [@alsheikh_public_2011; @sandve_ten_2013].
  The potential for subjectivity in colour measurement and the mathematical opacity of visual models for the average biologist means that, as in any field, is it critical that methods are reported rigorously.
-Methods reporting should ideally have well defined standards such that sufficient detail is provided to allow comprehensive peer review, precise replication, cross-study comparison and data sharing [@alsheikh_public_2011].
+Methods reporting should ideally have well defined standards such that sufficient detail is provided to allow comprehensive peer review, precise replication, cross-study comparison and data sharing [@alsheikh_public_2011; @sandve_ten_2013].
 The aims of this paper were to (a) review methods reporting in papers published on colour in 2013 from 22 top tier journals and (b) provide a set of recommendations for best-practice in reporting on the most common methods.
@@
 
-We developed a list of metrics we deemed critical to be reported for reproducibility of colour methods (Table 1) and analysis (Table 2), loosely based on Andersson and Prager [-@andersson_quantifying_2006] and Montgomerie [@montgomerie_analyzing_2006].
-These included quantitative methods including spectrophotometric and photographic techniques [as in @kemp_ornamental_2008; @chiao_visualization_2009] but not qualitative methods such as the use of Munsell chips or paint swatches [as in @hogstad_gloger_2009].
+We developed a list of information we deemed critical to be reported for reproducibility of colour methods (Table 1) and analysis (Table 2), loosely based on Andersson and Prager [-@andersson_quantifying_2006] and Montgomerie [@montgomerie_analyzing_2006].
+These included quantitative methods including spectrometric and photographic techniques [as in @kemp_ornamental_2008; @chiao_visualization_2009] but not qualitative methods such as the use of Munsell chips or paint swatches [as in @hogstad_gloger_2009].
 For visual modeling techniques we characterized a set of criteria required for reproducing the analysis (Table 2).
@@
 These are essential if such analyses are to be reproduced.
+These criteria provided that baseline against which to hold our results.
 
@@
 
-To determine the current state of methods reporting in the field we searched for papers from 2013 in 22 top colour science journals (_Animal Behaviour_, _Behavioral Ecology_, _Journal of Experimental Biology_, _Biology Letters_, _Proceedings of the Royal Society B: Biological Sciences_, _Biological Journal of the Linnean Society_, _Ethology_, _The American Naturalist_, _Ecology Letters_, _Journal of Evolutionary Biology_, _American Journal of Botany_, _Ecology and Evolution_, _Oikos_, _Ecology_, _Journal of Ecology_, _Behavioral Ecology and Sociobiology_, _Current Zoology_, _Naturwissenschaften_, _New Phytologist_, _PLoS ONE_, _Evolution_, _Functional Ecology_).
-For each of the journals, we used the Boolean search terms 'colour*', 'color*' or 'spectra*' and searched for these in the title or abstract from each of the journal home pages.
-We read the titles and abstracts of each paper and included only those papers that explicitly used either a spectrophotometer or camera to quantify colour signals.
-We excluded papers that were: 1) review papers; 2) methodological papers; 3) papers quantifying colour pattern rather than characteristics of each individual colour patch; and 4) papers taking microspectrophotometric measurments of retinal absorbance.
-To review the large number of papers returned from the searches, journals were haphazardly divided up between the authors.
-In order to reduce the risk of observer bias in our scoring, papers were read and reassessed by a total of three authors.
-Any discrepancies between the assessor scores were discussed and agreed upon prior to analysis.
+To determine the current state of methods reporting in the field we searched for papers from 2013 in 22 leading journals (_American Journal of Botany_, _The American Naturalist_, _Animal Behaviour_, _Behavioral Ecology_, _Behavioral Ecology and Sociobiology_, _Biological Journal of the Linnean Society_, _Biology Letters_, _Current Zoology_, _Ecology_, _Ecology and Evolution_, _Ecology Letters_, _Ethology_, _Evolution_, _Functional Ecology_, _Journal of Ecology_, _The Journal of Evolutionary Biology_, _The Journal of Experimental Biology_, _Naturwissenschaften_, _New Phytologist_, _Oikos_, _PLoS ONE_, _Proceedings of the Royal Society B: Biological Sciences_).
+On each of the journals' homepages, we used the Boolean terms 'colour*', 'color*' or 'spectra*' to search the title or abstract.
+We read the titles and abstracts of each paper and included only those papers that used either a spectrometer or camera to quantify colouration.
+We excluded papers review papers, methodological papers, papers quantifying colour patterns rather than the chromatic properties of a colour patch, and papers taking microspectrophotometric measurments of retinal absorbance.
+To review the large number of papers returned from our search, journals were haphazardly divided up between the authors.
+In order to reduce the risk of observer bias in our assesment, each paper was read and reassessed by three authors.
+Any discrepancies between assessor scores were discussed and resolved prior to analysis.
 We then tabulated what details of the methods were reported against our criteria.
 The data along with analysis script can be found at: https://github.com/daniel1noble/colsci_rep (TODO: GIVE IT A DOI?).
-We have kept the papers used in our dataset annoymous to avoid accusing specific colleagues of inadvertant reporting.
-We feel that the issue of reporting is something attributable to the ambiguity and lack of guidelines in our field and not something specific to individual research groups.
+We have kept the papers used in our dataset annoymous as our aim is to make a general statement oabout methods reporting in the field.
 
 Our search returned a total of 216 papers, however, only `r nrow(coldat)` met our criteria and were included in the final analysis.
-Overall, `r specs+both` studies used a spectrometer (henceforth "spec") to measure colour, while `r cam+both` used a camera.
-Of these studies, `r both` quantified colour using both a spec and a camera.
-Reporting of the make and model of the spec and camera was high (Figure 1, spectrometers = `r round(specModel, digits = 2)`% and cameras = `r round(camModel, digits = 2)`%), however, only `r round(CamLight, digits = 2)`%  and `r round(specLight, digits = 2)`% of camera and spec studies reported a light source.
+Overall, `r specs+both` studies used a spectrometer to measure colour, while `r cam+both` used a camera.
+Of these studies, `r both` quantified colour using both a spectrometer and a camera.
+Reporting of the make and model of the spectrometer and camera was high (Figure 1, spectrometers = `r round(specModel, digits = 2)`% and cameras = `r round(camModel, digits = 2)`%), however, only `r round(CamLight, digits = 2)`%  and `r round(specLight, digits = 2)`% of camera and spectrometer studies reported a light source.
 In addition, the number of reflectance spectra and pixels averaged were only reported in `r round(specAVG, digits = 2)`% and `r round(camAVG, digits = 2)`% of studies, respectively.
-Surprisingly, dark standards were only reported in `r round(drk_std, digits = 2)`% of the spec studies, while white standards were reported in `r round(white_stdspec, digits = 2)`%.
+Surprisingly, dark standards were only reported in `r round(drk_std, digits = 2)`% of the spectrometer studies, while white standards were reported in `r round(white_stdspec, digits = 2)`%.
 Studies quantifying spectral reflectance often did not report the integration time of the spectrophotmeter (`r round(intTime, digits = 2)`%), or the angle of the optical probe relative to the surface (`r round(angle, digits = 2)`%) or distance (`r round(dist, digits = 2)`%) the probe was held from the measurement surface.
 
-Analysis varied across studies.
-Overall `r analysisType[1]` papers analyzed  data by generating colourimetric variables [i.e.
-'spectral' measures of hue, saturation, and/or brightness; @montgomerie_analyzing_2006] whereas `r analysisType[4]` used visual models, `r analysisType[2]` used multiple analysis types, and `r analysisType[3]` used other forms of analysis (e.g.
-purely statistical analyses, such as PCA's on reflectance spectra).
+The types of analysis varied across studies.
+Overall `r analysisType[1]` papers generated colourimetric variables [i.e.
+'spectral' measures of hue, saturation, and/or brightness; @montgomerie_analyzing_2006] whereas `r analysisType[4]` used visual models, `r analysisType[2]` used both colourimetrics and visual models, and `r analysisType[3]` used other forms of analysis (e.g.
+purely statistical analyses, such as principle component analysis on reflectance spectra).
 Of the studies quantifying colour using spectral reflectance curves `r analysisTypeSpec[1]` used colourimetric measures, `r analysisTypeSpec[2]` used multiple analyses, `r analysisTypeSpec[4]` used visual models and `r analysisTypeSpec[3]` used other forms of analysis.
@@
 
-While some of the values reported above seem alarming, it is important to note that papers often made reference to previous work for various methods (`r round(ref_stud, digits = 2)`%), so these figures are slightly liberal given that referenced work may have comprehensively covered some of these criteria.
-Despite this, many of the studies that did not reference previous work were lacking in their reporting of many aspects of our criteria suggesting the need for more comprehensive reporting of data collection and analysis methods within such studies.
+While some of the percentages reported above seem alarming, it is important to note that around one-third of papers (`r round(ref_stud, digits = 2)`%) made reference to previous work for various methods.
+This means that our above figures are slightly liberal given that the referenced works may have comprehensively covered some of these criteria.
+Despite this, two-thirds of the studies did not reference previous work, and were missing important methodological details, suggesting the need for more comprehensive reporting of data collection and analysis methods within such studies.
 
@@
 
-Our review of recent colour literature highlights some important reporting gaps that will hinder the reproducibility of studies of biological colouration.
-While we recognize that journals have space limitations and that work may reference previous methological and analytical details, we strongly believe that researchers can do a better job of reporting details that are essential to both the outcome of their results and reproducibility of their work.
+Our review of recent colour literature highlights some important reporting gaps that will hinder the reproducibility and comparability of studies of biological colouration (Fig.
+1).
+While we recognize that journals have space limitations and that work may reference previous methological and analytical details, we strongly believe that it is to the field's advantage that researchers report all details essential to a study's transparency and reproducibility wherever possible.
+
 While all criteria are important, some deserve special attention.
-In particular, we found that colour scientists could do a better job reporting light sources, the numbers of pixels or spectra averaged, and the type of standards used (i.e.
-white and black reference samples).
-More importantly, the distance and angle of the optical probe relative to the measurment surface should be standard reporting for all studies using spectrophometry.
-Visual models were regularly used to analyse spectral data, something we welcome.
-However, there were many important aspects of the analysis that were seldom reported, including the type of quantum catch used, whether photoreceptor adaptation was modelled and the type of neural noise used in receptor-noise models.
-Photoreceptor densities should also be reported where possible even if using standard avian photorecptor densities.
+In particular, we found that the types of light sources, numbers of pixels or spectra averaged, and types of standards used (i.e.
+white and black reference samples) are under-reported, each of which is critical to repeatability and cross-comparison.
+Similarly, the distance and angle of the optical probe relative to the measurment surface and collector should be standard reporting for all studies using spectrophometry, as variation in these factors can lead to dramatically different results in some cases (REFS).
+Visual models were regularly used to analyse spectral data, however, there were many important aspects of the analysis that were seldom reported.
+These included the type of quantum catch used, whether photoreceptor adaptation was modelled and the type of noise used in receptor-noise models.
+This is problematic because...
+Also, photoreceptor densities should be reported where possible even if using standard avian photorecptor densities because....
 These all have important consequences on downstream analysis and the reproducibility of colour signal quantification [e.g.
 @kemp_ornamental_2008; @santos_comparison_2007].
-In addition, we recommend the presentation of spectral reflectance curves where possible; these figures are visually appealing and allow for the rapid assesment of the nature and quality of the data, independent of downstream processing.
-Surprisingly, we found that only `r round(fig, digits = 2)`% of studies using spectral reflectance data had a figure showing these curves.
+We also recommend the presentation of spectral reflectance curves where possible.
+These figures allow for the rapid assesment of the nature and quality of the data, independent of downstream processing (if the conditions under which they were captured is comprehensively reported.
+We found that only `r round(fig, digits = 2)`% of studies using spectral reflectance data had a figure showing these curves.
 When possible, we also recommend that authors justify data capture and analytical approaches - something few studies we investigated explicitly outlined.
 
-Our objective in this paper was not to highlight shortcomings in single studies, but to bring attantion to common oversights in reporting that can influence the reproducibility of colour research and cross-study comparisons.
-By no means do we wish to say that studies lacking important details are not of significant value.
-Indeed, even in our own work [e.g.
-@umbers_bright_2013] we have failed to report important details, which in retrospect can lead to confusion about how such data were collected and analysed.
-This review of reporting is aimed at providing constructive criticism; it is only through critical appraisal and collegial discussion that we can identify such gaps as areas for improvement, and constructively contribute to the growth of our field.
-Such assessments are particularly important in an age where accountability, transparency and reproducibility are of utmost importance.
+Our objective in this paper was to highlight common oversights in reporting that can influence the reproducibility of colour research and cross-study comparisons.
+By no means do we wish to say that studies lacking important details are not of significant value, but rather to provide recommendations for future work.
+It is only through critical appraisal and collegial discussion that we can identify areas for improvement, and constructively contribute to the growth of our field.
 
-We believe that part of the issue at hand is simply that many researchers seldom have clear guidelines to follow when undertaking research, and may overlook outlining important aspects of methodology and analysis.
+We believe that part of the issue at hand is simply that many researchers seldom have clear guidelines to follow when reporting methods, and may overlook outlining important aspects of methodology and analysis.
 We hope our critical evaluation of recent literature has highlighted a need for more careful description of how colour signals have been measured and analyzed.
-We are optimistic that this may help establish uniform reporting guidelines for publication, and to continue the advance of an exciting era in colour research.
+More importantly, we hope our guidelines for methods reporting helps establish a standard in this area, to continue the advance of an exciting era in colour research.
 

### 2014-12-01 R/analysis.r
Rediced redundancy in the analysis script. Need to now correct RmD File and based on new names etc. But should be relatively quick.

@@ -40,27 +40,38 @@ camModel  <- (sum(coldat$cam_model, na.rm = TRUE)/(cam+both))*100
 
 ## Of the studies using a spec, what proportion report the light source. Note that column needs to be extracted because it combines cameras and specs
 
-DatSpec   <- subset(coldat, stud_type == 1 |stud_type == 3)
+DatSpec         <- subset(coldat, stud_type == 1 |stud_type == 3)
 
-# Note that we need to change the rows where spec and cemera differ
-specLight <- (sum(DatSpec$light_source)/length(DatSpec$light_source))*100
+#Compute proportions of the various criteria
+propSpec        <- apply(DatSpec[sapply(DatSpec, is.numeric)], 2, function(x) sum(x[! is.na(x)])/length(x[! is.na(x)])*100)
+
+lenthSpec        <- apply(DatSpec[sapply(DatSpec, is.numeric)], 2, function(x) length(x[! is.na(x)]))
+
+#Extract only the rows we are interested in using
+
+criteriaSpec <- c("light_source","specpix_avg","dark_std", "white_stdspec", "int_time", "spec_angle", "spec_dist", "irrad_type", "vis_mod", "vis_mod_sp", "vis_mod_param", "vis_mod_adapt", "vis_mod_qcatch", "vis_mod_bkg", "vis_mod_noise_type", "refl_fig")
+
+propSpec_fin <- propSpec[criteriaSpec]
+
+lenthSpec_fin <- lenthSpec[criteriaSpec]
+
+names(propSpec_fin) <- paste("prop", names(propSpec_fin), sep = "_")
+names(lenthSpec_fin) <- paste("prop", names(lenthSpec_fin), sep = "_")
 
 ## Of the studies using a cam, what proportion report the light source. Note that column needs to be extracted because it combines cameras and specs
 
 DatCam    <- subset(coldat, stud_type == 2 |stud_type == 3)
-CamLight  <- (sum(DatCam$light_source, na.rm = TRUE)/length(DatCam$light_source))*100
 
-## Of studies spec'ing what proportion report a dark standard
-drk_std       <- (sum(DatSpec$dark_std)/(length(DatSpec$dark_std)))*100
-white_stdspec <- (sum(DatSpec$white_stdspec)/length(DatSpec$white_stdspec))*100
[…]

### 2014-12-01 R/analysis.r
Missed the last figure with analyses so quickly cleaned up that script too

@@ -114,9 +114,9 @@ text(paste(round(dat[1,], digits = 0), "%", sep = ""), x = bp.out, y = round(dat
 mtext("a)", adj = -0.13, padj = -1, cex = 2)
 
 
-props_analy   <- c(prop_irrad, prop_spvismod,  prop_adap, prop_qcatch, prop_bkg, prop_noise)
+props_analy   <- propSpec_fin[c("prop_irrad_type","prop_vis_mod_sp", "prop_vis_mod_adapt", "prop_vis_mod_qcatch", "prop_vis_mod_bkg", "prop_vis_mod_noise_type")]
 
-N_props_analy <- c(length(DatSpec$irrad_type[! is.na(DatSpec$irrad_type)]), length(DatSpec$vis_mod_sp[! is.na(DatSpec$vis_mod_sp)]), length(DatSpec$vis_mod_adapt[! is.na(DatSpec$vis_mod_adapt)]), length(DatSpec$vis_mod_qcatch[! is.na(DatSpec$vis_mod_qcatch)]), length(DatSpec$vis_mod_bkg[! is.na(DatSpec$vis_mod_bkg)]), length(DatSpec$vis_mod_noise_type[! is.na(DatSpec$vis_mod_noise_type)]))
+N_props_analy <- lenthSpec_fin[c("prop_irrad_type","prop_vis_mod_sp", "prop_vis_mod_adapt", "prop_vis_mod_qcatch", "prop_vis_mod_bkg", "prop_vis_mod_noise_type")]
 
 dat_analy     <- t(arrange(data.frame(props_analy, N_props_analy), props_analy, decreasing = TRUE))
 

### 2014-12-01 docs/ms.Rmd
Updated the ms with the more recent analysis code to call on various proportions.

@@
 Of these studies, `r both` quantified colour using both a spectrometer and a camera.
-Reporting of the make and model of the spectrometer and camera was high (Figure 1, spectrometers = `r round(specModel, digits = 2)`% and cameras = `r round(camModel, digits = 2)`%), however, only `r round(CamLight, digits = 2)`%  and `r round(specLight, digits = 2)`% of camera and spectrometer studies reported a light source.
-In addition, the number of reflectance spectra and pixels averaged were only reported in `r round(specAVG, digits = 2)`% and `r round(camAVG, digits = 2)`% of studies, respectively.
-Surprisingly, dark standards were only reported in `r round(drk_std, digits = 2)`% of the spectrometer studies, while white standards were reported in `r round(white_stdspec, digits = 2)`%.
-Studies quantifying spectral reflectance often did not report the integration time of the spectrophotmeter (`r round(intTime, digits = 2)`%), or the angle of the optical probe relative to the surface (`r round(angle, digits = 2)`%) or distance (`r round(dist, digits = 2)`%) the probe was held from the measurement surface.
+Reporting of the make and model of the spectrometer and camera was high (Figure 1, spectrometers = `r round(specModel, digits = 2)`% and cameras = `r round(camModel, digits = 2)`%), however, only `r round(propCam_fin["prop_light_source"], digits = 2)`%  and `r round(propSpec_fin["prop_light_source"], digits = 2)`% of camera and spectrometer studies reported a light source.
+In addition, the number of reflectance spectra and pixels averaged were only reported in `r round(propSpec_fin["prop_specpix_avg"], digits = 2)`% and `r round(propCam_fin["prop_specpix_avg"], digits = 2)`% of studies, respectively.
+Surprisingly, dark standards were only reported in `r round(propSpec_fin["prop_dark_std"], digits = 2)`% of the spectrometer studies, while white standards were reported in `r round(propSpec_fin["prop_white_stdspec"], digits = 2)`%.
+Studies quantifying spectral reflectance often did not report the integration time of the spectrophotmeter (`r round(propSpec_fin["prop_int_time"], digits = 2)`%), or the angle of the optical probe relative to the surface (`r round(propSpec_fin["prop_spec_angle"], digits = 2)`%) or distance (`r round(propSpec_fin["prop_spec_dist"], digits = 2)`%) the probe was held from the measurement surface.
 
@@
 Case specific models were used in `r vismod_type[2]` papers.
-Of visual modelling papers, `r round(prop_irrad, digits =2)`% reported the type of irradiance when necessary, `r round(prop_spvismod, digits = 2)`% the species modelled as a receiver, and `r round(propvis_mod_param, digits = 2)`% made explicit and complete mention of the parameters they used (e.g.
-cone densities, signal-to-noise ratios).
-The background used in visual models was reported `r round(prop_bkg, digits = 2)`% of the time.
-The type of receptor noise modelled was reported in only `r round(prop_noise, digits = 2)`% of the studies.
-Explicit mention of the type of quantum catch (`r round(prop_qcatch, digits = 2)`% of studies), or whether photoreceptor adaptation (`r round(prop_adap, digits = 2)`%) was modelled, was not common.
+Of visual modelling papers, `r round(propSpec_fin["prop_irrad_type"], digits =2)`% reported the type of irradiance when necessary, `r round(propSpec_fin["prop_vis_mod_sp"], digits = 2)`% the species modelled as a receiver.
+The background used in visual models was reported `r round(propSpec_fin["prop_vis_mod_bkg"], digits = 2)`% of the time.
+The type of receptor noise modelled was reported in only `r round(propSpec_fin["prop_vis_mod_noise_type"], digits = 2)`% of the studies.
+Explicit mention of the type of quantum catch (`r round(propSpec_fin["prop_vis_mod_qcatch"], digits = 2)`% of studies), or whether photoreceptor adaptation (`r round(propSpec_fin["prop_vis_mod_adapt"], digits = 2)`%) was modelled, was not common.
 
@@
 These figures allow for the rapid assesment of the nature and quality of the data, independent of downstream processing (if the conditions under which they were captured is comprehensively reported.
-We found that only `r round(fig, digits = 2)`% of studies using spectral reflectance data had a figure showing these curves.
+We found that only `r round(propSpec["refl_fig"], digits = 2)`% of studies using spectral reflectance data had a figure showing these curves.
 When possible, we also recommend that authors justify data capture and analytical approaches - something few studies we investigated explicitly outlined.

### 2014-12-01 R/analysis.r
Minor cleaning of R script to remove some more lines of code.

@@ -42,20 +42,14 @@ camModel  <- (sum(coldat$cam_model, na.rm = TRUE)/(cam+both))*100
 
 DatSpec         <- subset(coldat, stud_type == 1 |stud_type == 3)
 
-#Compute proportions of the various criteria
-propSpec        <- apply(DatSpec[sapply(DatSpec, is.numeric)], 2, function(x) sum(x[! is.na(x)])/length(x[! is.na(x)])*100)
-
-lenthSpec        <- apply(DatSpec[sapply(DatSpec, is.numeric)], 2, function(x) length(x[! is.na(x)]))
-
-#Extract only the rows we are interested in using
-
+#Compute proportions of the various criteria. Extract only the rows we are interested in using
 criteriaSpec <- c("light_source","specpix_avg","dark_std", "white_stdspec", "int_time", "spec_angle", "spec_dist", "irrad_type", "vis_mod", "vis_mod_sp", "vis_mod_param", "vis_mod_adapt", "vis_mod_qcatch", "vis_mod_bkg", "vis_mod_noise_type", "refl_fig")
 
-propSpec_fin <- propSpec[criteriaSpec]
+propSpec_fin   <- apply(DatSpec[sapply(DatSpec, is.numeric)], 2, function(x) sum(x[! is.na(x)])/length(x[! is.na(x)])*100)[criteriaSpec]
 
-lenthSpec_fin <- lenthSpec[criteriaSpec]
+lenthSpec_fin  <- apply(DatSpec[sapply(DatSpec, is.numeric)], 2, function(x) length(x[! is.na(x)]))[criteriaSpec]
 
-names(propSpec_fin) <- paste("prop", names(propSpec_fin), sep = "_")
+names(propSpec_fin)  <- paste("prop", names(propSpec_fin), sep = "_")
 names(lenthSpec_fin) <- paste("prop", names(lenthSpec_fin), sep = "_")
 
 ## Of the studies using a cam, what proportion report the light source. Note that column needs to be extracted because it combines cameras and specs
@@ -63,15 +57,12 @@ names(lenthSpec_fin) <- paste("prop", names(lenthSpec_fin), sep = "_")
 DatCam    <- subset(coldat, stud_type == 2 |stud_type == 3)
 
 #Compute various proportions
[…]

### 2014-12-03 docs/ms.Rmd
Re-draft intro, begin restructuring to focus more broadly on reproducibility

@@
-# The transparency of methods-reporting in studies of biological colouration
+# Reproducible research in biological colour science
 
@@
 
-## Abstract
-Over the last 50 years or so, our enhanced capacity for quantitative colour measurement has accelerated understanding of the evolution of biological colouration.
-However, while spectrometry and digital photography are inexpensive, accessible and efficient, details of exactly how data are collected and analysed are not reported consistently in the literature.
-Similarly, estimates of animals’ visual capacity are widely applied to colour data, but details on assumptions and settings in analyses are not reported completely.
-To quantify the variability of methods reporting in studies of colouration, we reviewed papers published in 2013 in 22 leading journals.
-Using the search terms colour* OR color* OR spectra* in the title or abstract, we identified 60 papers in which colour had been measured and analysed.
-We extracted the methods and noted the details included.
-We found a great deal of variability in the level of detail in which colour methods is reported.
-In response we have outlined explicit recommendations on the important methodological details required to ensure studies are transparent and repeatable.
-We hope that this will help to standardize methods reporting in the field of biological colouration paving the way for direct comparison across studies.
+## Introduction
 
-\newpage
+Studies of biological colouration have produced insights into fundamental evolutionary and ecological processes, and research into colour traits is a rapidly growing, interdisciplinary field [].
+The resurgent interest in biological colouration has in large part been driven by the increased availability of key technologies, including spectrometry and photography, and concurrent developments in methods for analysing colour data, such as visual models [].
+While such advances are extremely positive for the field, the rapid rate of progression means that a deep understanding of common methods is increasingly beyond the grasp of the average researcher.
+Indeed, the subtle complexity of these techniques [e.g.
+the 3-7 parameters that define a typical visual model; ] is often overlooked by empiricists, which leads to critical methodological information often not being reported (Fig.
+1); thus limiting a study's utility.
 
-## Introduction
-Generations of biologists have endeavored to explain the mechanisms and functions of animal and plant colouration [@thayer_concealing_1909; @poulton_colours_1890; @wallace_natural_1891].
-For these biologists, identifying best practices in measuring colour has been a great challenge.
-Over the past 50 years, biologists have developed and adapted techniques for the objective measurement of colouration that are now widespread.
-These include spectrometry, photography, and visual modelling.
+Replication and transparency lie at the heart of the scientific enterprise.
+True replication of individual studies based on independantly collected data is the gold standard in cumulative science.
+This may often be unfeasable however, espescially in field-based research, which includes a great deal of research on colour-traits [].
+In contrast, reproducability - the re-analysis of a study's data by independant researchers - has been presented as a readily attainable minimum standard [@peng_reproducible_2006; @peng_reproducible_2011].
+A study may be considered truly reproducible when it satifies three broad criteria: (i) methods are reported completely, (ii) data are publicly archived and available, and (iii) analysis code is made publicly available, such that the chain of modification of raw data is documented and preserved.
+While completely reproducible research [e.g.
+@fitzjohn_much_2014] is a laudable goal, the considerable demands on researchers means that it will often, in practice, be unattanable.
+Nevertheless, even partial reproducibility is of tremendous value as it allows greater comparability between studies, and may serve as a launching point for future research.
 
+Our aim here to explore current levels of reproducibility in studies of biological colouration, and to outline ways in which they may be improved in line with points (i) - (iii) above.
+We begin by briefly outlining some of the most common methods for studying biological colouration.
+We then present a set of guidelines for the reporting of methods for the capture and analysis of colour data, and we assess a subset of the literature against these criteria to assess the current state of methods reporting in colour science.
+We also quantify the availability of publicly archived data and code, and suggest some tools...
+
+## Measuring and analysing colour
+
+todo: hack this back
+
+Generations of biologists have endeavored to explain the mechanisms and functions of animal and plant colouration [@thayer_concealing_1909; @poulton_colours_1890; @wallace_natural_1891], and identifying best practices in measuring colour has been a great challenge.
 The application of spectrometry to questions of biological colouration revolutionized the field [@dyck_determination_1966].
@@
 
-## Ideal reporting in colour studies
+## Assessing reproducibility
 
-We developed a list of information we deemed critical to be reported for reproducibility of colour methods (Table 1) and analysis (Table 2), loosely based on Andersson and Prager [-@andersson_quantifying_2006] and Montgomerie [@montgomerie_analyzing_2006].
-These included quantitative methods including spectrometric and photographic techniques [as in @kemp_ornamental_2008; @chiao_visualization_2009] but not qualitative methods such as the use of Munsell chips or paint swatches [as in @hogstad_gloger_2009].
-For visual modeling techniques we characterized a set of criteria required for reproducing the analysis (Table 2).
-Much of these criteria specifically targeted the necessary parameters required to be explicitly defined by the user during analysis.
-These are essential if such analyses are to be reproduced.
-These criteria provided that baseline against which to hold our results.
-
-## What is reporting like in the literature?
-
-To determine the current state of methods reporting in the field we searched for papers from 2013 in 22 leading journals (_American Journal of Botany_, _The American Naturalist_, _Animal Behaviour_, _Behavioral Ecology_, _Behavioral Ecology and Sociobiology_, _Biological Journal of the Linnean Society_, _Biology Letters_, _Current Zoology_, _Ecology_, _Ecology and Evolution_, _Ecology Letters_, _Ethology_, _Evolution_, _Functional Ecology_, _Journal of Ecology_, _The Journal of Evolutionary Biology_, _The Journal of Experimental Biology_, _Naturwissenschaften_, _New Phytologist_, _Oikos_, _PLoS ONE_, _Proceedings of the Royal Society B: Biological Sciences_).
+To determine the current state of reproducibility in the field we searched for papers from 2013 in 22 leading journals (_American Journal of Botany_, _The American Naturalist_, _Animal Behaviour_, _Behavioral Ecology_, _Behavioral Ecology and Sociobiology_, _Biological Journal of the Linnean Society_, _Biology Letters_, _Current Zoology_, _Ecology_, _Ecology and Evolution_, _Ecology Letters_, _Ethology_, _Evolution_, _Functional Ecology_, _Journal of Ecology_, _The Journal of Evolutionary Biology_, _The Journal of Experimental Biology_, _Naturwissenschaften_, _New Phytologist_, _Oikos_, _PLoS ONE_, _Proceedings of the Royal Society B: Biological Sciences_).
 On each of the journals' homepages, we used the Boolean terms 'colour*', 'color*' or 'spectra*' to search the title or abstract.
@@
 The data along with analysis script can be found at: https://github.com/daniel1noble/colsci_rep (TODO: GIVE IT A DOI?).
-We have kept the papers used in our dataset annoymous as our aim is to make a general statement oabout methods reporting in the field.
+We have kept the papers used in our dataset annoymous as our aim is to make a general statement about methods reporting in the field.
 
+## Methods reporting in colour studies
+
+_Guidelines for methods reporting_
+
+todo: expand table discussion
+
+We developed a list of information we deemed critical to be reported for reproducibility of colour methods (Table 1) and analysis (Table 2), loosely based on Andersson and Prager [-@andersson_quantifying_2006] and Montgomerie [@montgomerie_analyzing_2006].
+These included quantitative methods including spectrometric and photographic techniques [as in @kemp_ornamental_2008; @chiao_visualization_2009] but not qualitative methods such as the use of Munsell chips or paint swatches [as in @hogstad_gloger_2009].
+For analytical techniques we characterized a set of criteria required for reproducing the analysis (Table 2).
+Much of these criteria specifically targeted the necessary parameters required to be explicitly defined by the user during analysis.
+These are essential if such analyses are to be reproduced.
+These criteria provided that baseline against which to hold our results.
 Our search returned a total of 216 papers, however, only `r nrow(coldat)` met our criteria and were included in the final analysis.
+
+_The curent state of methods reporting in the literature_
+
 Overall, `r specs+both` studies used a spectrometer to measure colour, while `r cam+both` used a camera.
@@
 
-##Conclusions
+## The public availability of data and code
+
+todo: results here
+
+_Tools for reproducible research_
+
+Data - archiving (figshare, dryad), raw data preferable to processed data, concerns over data sharing, embargos are fine just get it archived.
+Code - benefits of code over point-and-click, open source shit (R, pavo, things), github, sourcefourge etc.
+
+## Conclusions
 
@@
 These figures allow for the rapid assesment of the nature and quality of the data, independent of downstream processing (if the conditions under which they were captured is comprehensively reported.
-We found that only `r round(propSpec["refl_fig"], digits = 2)`% of studies using spectral reflectance data had a figure showing these curves.
+We found that only `r round(propSpec_fin["prop_refl_fig"], digits = 2)`% of studies using spectral reflectance data had a figure showing these curves.
 When possible, we also recommend that authors justify data capture and analytical approaches - something few studies we investigated explicitly outlined.

### 2014-12-03 docs/ms.Rmd
Tighten intro some more, hack back methods bit

@@
-# Reproducible research in biological colour science
+# Reproducible research in the study of biological colouration
 
@@
 
-Keywords: #colsci, spectrometry, photography, visual ecology, pigment, structural colour, visual modeling, methodology
+Keywords: #colsci, spectrometry, photography, visual ecology, pigment, structural colour, visual modelling, methodology
 
@@
 
-Studies of biological colouration have produced insights into fundamental evolutionary and ecological processes, and research into colour traits is a rapidly growing, interdisciplinary field [].
-The resurgent interest in biological colouration has in large part been driven by the increased availability of key technologies, including spectrometry and photography, and concurrent developments in methods for analysing colour data, such as visual models [].
-While such advances are extremely positive for the field, the rapid rate of progression means that a deep understanding of common methods is increasingly beyond the grasp of the average researcher.
-Indeed, the subtle complexity of these techniques [e.g.
-the 3-7 parameters that define a typical visual model; ] is often overlooked by empiricists, which leads to critical methodological information often not being reported (Fig.
-1); thus limiting a study's utility.
+Studies of biological colouration have generated insights into fundamental evolutionary and ecological processes, and research into colour traits is a rapidly growing field [].
+The resurgent interest in biological colouration has in large part been driven by the increased availability of key technologies, including spectrometry and photography, and concurrent developments in methods for analysing colour data, such as visual models [@endler_comparing_2005; @stevens_using_2007; @kelber_animal_2003].
+While such advances are positive for the field, the rapid rate of progression means that a deep understanding of common methods is increasingly beyond the grasp of the average researcher.
+As a consequence, the subtle complexity of many techniques [e.g.
+the 3-7 parameters that define a typical visual model; @endler_comparing_2005; @vorobyev_receptor_1998] is often overlooked by empiricists, which leads to critical methodological information often not being reported (Fig.
+1).
 
 Replication and transparency lie at the heart of the scientific enterprise.
-True replication of individual studies based on independantly collected data is the gold standard in cumulative science.
-This may often be unfeasable however, espescially in field-based research, which includes a great deal of research on colour-traits [].
-In contrast, reproducability - the re-analysis of a study's data by independant researchers - has been presented as a readily attainable minimum standard [@peng_reproducible_2006; @peng_reproducible_2011].
-A study may be considered truly reproducible when it satifies three broad criteria: (i) methods are reported completely, (ii) data are publicly archived and available, and (iii) analysis code is made publicly available, such that the chain of modification of raw data is documented and preserved.
+The full replication of individual studies based on independently collected data is the gold standard in cumulative science.
+This may often be unfeasible however, especially in field-based research, which includes a great deal of research on colour-traits [].
+In contrast, reproducibility - the re-analysis of a study's data by independent researchers - has been presented as a readily attainable alternative [@peng_reproducible_2006; @peng_reproducible_2011].
+A study may be considered truly reproducible when it satisfies three broad criteria: (i) methods are reported completely, (ii) data are publicly archived and available, and (iii) analysis code is made publicly available, such that the chain of modification of raw data is documented and preserved.
 While completely reproducible research [e.g.
-@fitzjohn_much_2014] is a laudable goal, the considerable demands on researchers means that it will often, in practice, be unattanable.
-Nevertheless, even partial reproducibility is of tremendous value as it allows greater comparability between studies, and may serve as a launching point for future research.
+@fitzjohn_much_2014] is a laudable goal, the considerable demands it imposes on researchers means that it will often, in practice, be unattainable.
+Nevertheless, even partial reproducibility is of tremendous value both to individual researchers and the broader community as it ensures greater comparability between studies, and may provide a launching point for future research.
 
-Our aim here to explore current levels of reproducibility in studies of biological colouration, and to outline ways in which they may be improved in line with points (i) - (iii) above.
+Our aim here to explore current levels of reproducibility in studies of biological colouration, and to outline simple ways in which they may be improved in line with points (i) - (iii) above.
 We begin by briefly outlining some of the most common methods for studying biological colouration.
-We then present a set of guidelines for the reporting of methods for the capture and analysis of colour data, and we assess a subset of the literature against these criteria to assess the current state of methods reporting in colour science.
-We also quantify the availability of publicly archived data and code, and suggest some tools...
+We then present a set of guidelines for the reporting of methods, and we assess a subset of the literature against these criteria to assess the current state of methods reporting in colour science.
+We also analyse the availability of publicly archived data and code, and suggest some tools that may be useful for increasing reproducibility.
 
@@
 
-todo: hack this back
+todo: hack this back, make it more relevant to reproducibility
 
-Generations of biologists have endeavored to explain the mechanisms and functions of animal and plant colouration [@thayer_concealing_1909; @poulton_colours_1890; @wallace_natural_1891], and identifying best practices in measuring colour has been a great challenge.
+Generations of biologists have endeavoured to explain the mechanisms and functions of animal and plant colouration [@thayer_concealing_1909; @poulton_colours_1890; @wallace_natural_1891], and identifying best practices in measuring colour has been a great challenge.
 The application of spectrometry to questions of biological colouration revolutionized the field [@dyck_determination_1966].
-Before its widespread use, researchers assigned categorical colour descriptions, arbitrary rankings or scores [e.g.
-@aldrich_ecogeographic_1991; @bailey_latitudinal_1978; @wheelwright_colors_1985] or matched paint swatches or munsell chip ratings based on their own visual assessment [e.g.
-@burtt_analysis_1986; @key_temperature_1954].
-Although no doubt applied carefully, these techniques limited progress.
-For example, many researchers designed colour assessment methods specifically for their projects, rendering data incomparable across studies [e.g.
-@aldrich_ecogeographic_1991; @bailey_latitudinal_1978].
-This is especially problematic when different measurement techniques are used on the same species [e.g.
-house finch _Carpodacus mexicanus_; @aldrich_origin_1978; @hill_female_1990].
-Likewise, because human vision is inherently different to most other animals [@bennett_sexual_1994; @douglas_spectral_2014], its use has lead to erroneous conclusions [@bennett_sexual_1994; @endler_measurement_1990].
-
-Measuring reflectance spectra via spectrometry is now a reasonably inexpensive and convenient technique.
-Spectrometry allows the capture of quantitative and arguably repeatable data and has been rapidly taken up by researchers the field.
+Now relatively inexpensive and convenient, spectrometry allows the capture of quantitative and arguably repeatable data and has been rapidly taken up by researchers the field.
 However one drawback to this methodology, is the ease with which erroneous data can be collected if attention is not paid to finer details of spectrum capture.
@@
 High-resolution cameras are inexpensive,  and allow instantaneous, repeatable sampling of multiple colour patches [@mckay_use_2013].
-The main challenge for gathering colour data from photographs involve ensuring standardization, such as linearization and consistent lighting [@stevens_using_2007].
+The main challenge for gathering colour data from photographs involve ensuring standardization, such as linearisation and consistent lighting [@stevens_using_2007].
 Also, most digital cameras used for the measurement of biological colouration are consumer products designed for the human visual system.
-Thus hardware modifications for ultraviolet and infrared imaging are generally necessary [e.g.
+Thus hardware modifications for ultraviolet and infra-red imaging are generally necessary [e.g.
 @stevens_colour_2013].
@@
 Expansion in the availability of objective methods for the measurement of colour has been matched by advances in theory and analysis.
-In particular, the development of sensory models has produced significant insights into how colourful stimuli are perceived [@chittka_colour_1992; @endler_comparing_2005; @vorobyev_receptor_1998].
-Sensory models typically attempt to describe the subjective perception of colour or brightness information as a function of an object’s reflectance, the ambient illumination, and a receiver’s sensory system [e.g.
+In particular, the development of visual models has produced significant insights into the early visual processing of colourful stimuli [@chittka_colour_1992; @endler_comparing_2005; @vorobyev_receptor_1998] and, in a few cases, how they may be perceived [@kelber_spectral_2010].
+Visual models typically attempt to describe the subjective perception of colour information as a function of an object’s reflectance, the ambient illumination, and a receiver’s sensory system [e.g.
 @endler_comparing_2005; @vorobyev_receptor_1998].
-These models enable researchers to move beyond purely quantitative comparisons of reflectance spectra and adopt biologically relevant perspectives of signal receivers to formulate and test hypotheses.
+These models enable researchers to move beyond purely quantitative comparisons of reflectance spectra and adopt more biologically relevant perspectives to refine and test hypotheses.
 
-Although relatively easy to implement, and potentially powerful, sensory models are built on multiple assumptions about the way in which stimuli are visually and perceptually processed and these can dramatically shape the results of a given analysis [@lind_avian_2009; @pike_preserving_2012].
+Although relatively easy to implement, and potentially powerful, sensory models are built on multiple assumptions about the way in which stimuli are visually processed, that can dramatically shape the results of a given analysis [@lind_avian_2009; @pike_preserving_2012].
 For example, Vorobyev et al.’s [-@vorobyev_receptor_1998; -@vorobyev_colour_2001] widely used receptor-noise model requires researchers to specify the form of receptor quantum catches (i.e.
@@
 The potential for subjectivity in colour measurement and the mathematical opacity of visual models for the average biologist means that, as in any field, is it critical that methods are reported rigorously.
-Methods reporting should ideally have well defined standards such that sufficient detail is provided to allow comprehensive peer review, precise replication, cross-study comparison and data sharing [@alsheikh_public_2011; @sandve_ten_2013].
-The aims of this paper were to (a) review methods reporting in papers published on colour in 2013 from 22 top tier journals and (b) provide a set of recommendations for best-practice in reporting on the most common methods.
-Our review specifically focuses on papers that quantify the external coloration of a plant or animal.
-These may be purely descriptive quantifications of color, or they may be incorporated into models of animal vision.
-Our hope is that this paper will provide a useful set of guidelines to improve the reproducibility and transparency of studies on colour.
+Methods reporting should ideally have well defined standards such that sufficient detail is provided to allow comprehensive peer review, precise replication, cross-study comparison and data sharing [@alsheikh_public_2011; @sandve_ten_2013]....
 
@@
 
-To determine the current state of reproducibility in the field we searched for papers from 2013 in 22 leading journals (_American Journal of Botany_, _The American Naturalist_, _Animal Behaviour_, _Behavioral Ecology_, _Behavioral Ecology and Sociobiology_, _Biological Journal of the Linnean Society_, _Biology Letters_, _Current Zoology_, _Ecology_, _Ecology and Evolution_, _Ecology Letters_, _Ethology_, _Evolution_, _Functional Ecology_, _Journal of Ecology_, _The Journal of Evolutionary Biology_, _The Journal of Experimental Biology_, _Naturwissenschaften_, _New Phytologist_, _Oikos_, _PLoS ONE_, _Proceedings of the Royal Society B: Biological Sciences_).
+To assess the current state of reproducibility in the field we searched papers from 2013 in 22 leading journals (_American Journal of Botany_, _The American Naturalist_, _Animal Behaviour_, _Behavioral Ecology_, _Behavioral Ecology and Sociobiology_, _Biological Journal of the Linnean Society_, _Biology Letters_, _Current Zoology_, _Ecology_, _Ecology and Evolution_, _Ecology Letters_, _Ethology_, _Evolution_, _Functional Ecology_, _Journal of Ecology_, _The Journal of Evolutionary Biology_, _The Journal of Experimental Biology_, _Naturwissenschaften_, _New Phytologist_, _Oikos_, _PLoS ONE_, _Proceedings of the Royal Society B: Biological Sciences_).
 On each of the journals' homepages, we used the Boolean terms 'colour*', 'color*' or 'spectra*' to search the title or abstract.
 We read the titles and abstracts of each paper and included only those papers that used either a spectrometer or camera to quantify colouration.
-We excluded papers review papers, methodological papers, papers quantifying colour patterns rather than the chromatic properties of a colour patch, and papers taking microspectrophotometric measurments of retinal absorbance.
+We excluded papers review papers, methodological papers, papers quantifying colour patterns rather than the chromatic properties of a colour patch, and papers taking micro-spectrometric measurements of retinal absorbance.
 To review the large number of papers returned from our search, journals were haphazardly divided up between the authors.
@@
 The data along with analysis script can be found at: https://github.com/daniel1noble/colsci_rep (TODO: GIVE IT A DOI?).
-We have kept the papers used in our dataset annoymous as our aim is to make a general statement about methods reporting in the field.
+We have kept the papers used in our dataset anonymous as our aim is to make a general statement about methods reporting in the field.
 
@@
 white and black reference samples) are under-reported, each of which is critical to repeatability and cross-comparison.
-Similarly, the distance and angle of the optical probe relative to the measurment surface and collector should be standard reporting for all studies using spectrophometry, as variation in these factors can lead to dramatically different results in some cases (REFS).
+Similarly, the distance and angle of the optical probe relative to the measurement surface and collector should be standard reporting for all studies using spectrometry, as variation in these factors can lead to dramatically different results in some cases (REFS).
 Visual models were regularly used to analyse spectral data, however, there were many important aspects of the analysis that were seldom reported.
@@
 This is problematic because...
-Also, photoreceptor densities should be reported where possible even if using standard avian photorecptor densities because....
+Also, photoreceptor densities should be reported where possible even if using standard avian photoreceptor densities because....
 These all have important consequences on downstream analysis and the reproducibility of colour signal quantification [e.g.
@@
 We also recommend the presentation of spectral reflectance curves where possible.
-These figures allow for the rapid assesment of the nature and quality of the data, independent of downstream processing (if the conditions under which they were captured is comprehensively reported.
+These figures allow for the rapid assessment of the nature and quality of the data, independent of downstream processing (if the conditions under which they were captured is comprehensively reported.
 We found that only `r round(propSpec_fin["prop_refl_fig"], digits = 2)`% of studies using spectral reflectance data had a figure showing these curves.
@@
 We believe that part of the issue at hand is simply that many researchers seldom have clear guidelines to follow when reporting methods, and may overlook outlining important aspects of methodology and analysis.
-We hope our critical evaluation of recent literature has highlighted a need for more careful description of how colour signals have been measured and analyzed.
+We hope our critical evaluation of recent literature has highlighted a need for more careful description of how colour signals have been measured and analysed.
 More importantly, we hope our guidelines for methods reporting helps establish a standard in this area, to continue the advance of an exciting era in colour research.

### 2014-12-04 docs/ms.Rmd
Continued shaping intro, drafted half of data/code results section

@@
-# Reproducible research in the study of biological colouration
+# Reproducible research in the study of colour in nature
 
@@
 
-Studies of biological colouration have generated insights into fundamental evolutionary and ecological processes, and research into colour traits is a rapidly growing field [].
-The resurgent interest in biological colouration has in large part been driven by the increased availability of key technologies, including spectrometry and photography, and concurrent developments in methods for analysing colour data, such as visual models [@endler_comparing_2005; @stevens_using_2007; @kelber_animal_2003].
-While such advances are positive for the field, the rapid rate of progression means that a deep understanding of common methods is increasingly beyond the grasp of the average researcher.
-As a consequence, the subtle complexity of many techniques [e.g.
-the 3-7 parameters that define a typical visual model; @endler_comparing_2005; @vorobyev_receptor_1998] is often overlooked by empiricists, which leads to critical methodological information often not being reported (Fig.
-1).
+Studies of biological colouration have generated insights into fundamental evolutionary and ecological processes, and research into colour traits is a rapidly growing field [@kelber_spectral_2010].
+The resurgent interest in biological colouration has in large part been driven by the increased availability of key technologies, including spectrometry and photography, and concurrent advances in methods for analysing colour data, such as visual models [@endler_comparing_2005; @stevens_using_2007; @kelber_animal_2003].
+These developments are positive for the field, and have allowed increasing numbers of researchers to rapidly collect and analyse ever greater amounts of data.
+The rapid rate of progression, however, means that a deep understanding of common methods is increasingly beyond the grasp of the average researcher.
+As a consequence, the subtle complexity of many analytical techniques [e.g.
+the 3-7 parameters that define a typical visual model; @endler_comparing_2005; @vorobyev_receptor_1998] is often overlooked by empiricists, which leads to critical methodological information often not being reported.
 
 Replication and transparency lie at the heart of the scientific enterprise.
-The full replication of individual studies based on independently collected data is the gold standard in cumulative science.
-This may often be unfeasible however, especially in field-based research, which includes a great deal of research on colour-traits [].
-In contrast, reproducibility - the re-analysis of a study's data by independent researchers - has been presented as a readily attainable alternative [@peng_reproducible_2006; @peng_reproducible_2011].
+The full replication of individual studies may often be infeasible @peng_reproducible_2011, but reproducibility - the re-analysis of a study's data by independent researchers - has been presented as a readily attainable alternative [@peng_reproducible_2006; @peng_reproducible_2011].
 A study may be considered truly reproducible when it satisfies three broad criteria: (i) methods are reported completely, (ii) data are publicly archived and available, and (iii) analysis code is made publicly available, such that the chain of modification of raw data is documented and preserved.
@@
 @fitzjohn_much_2014] is a laudable goal, the considerable demands it imposes on researchers means that it will often, in practice, be unattainable.
-Nevertheless, even partial reproducibility is of tremendous value both to individual researchers and the broader community as it ensures greater comparability between studies, and may provide a launching point for future research.
+Nevertheless, even partial reproducibility through the simple practice of complete methods reporting and public data archiving is of tremendous value.
+It ensures greater comparability between studies, allows for the the independent verification of results, and provides a foundation for the testing of new ideas and methods [@whitlock_data_2011; @van_trouble_2011; @piwowar_sharing_2007].
 
-Our aim here to explore current levels of reproducibility in studies of biological colouration, and to outline simple ways in which they may be improved in line with points (i) - (iii) above.
+Our aim here to explore current levels of reproducibility in studies of biological colouration, and to outline ways in which they may be improved in line with points (i) - (iii) above.
 We begin by briefly outlining some of the most common methods for studying biological colouration.
-We then present a set of guidelines for the reporting of methods, and we assess a subset of the literature against these criteria to assess the current state of methods reporting in colour science.
-We also analyse the availability of publicly archived data and code, and suggest some tools that may be useful for increasing reproducibility.
+We then present a set of guidelines for the comprehensive reporting of methods, and we analyse a subset of the literature against these criteria to assess the current state of methods reporting in colour science.
+We also assess the availability of publicly archived data and code, and suggest some tools that may be useful for increasing the reproducibility of colour trait research.
 
@@
 
-_The curent state of methods reporting in the literature_
+_The current state of methods reporting in the literature_
 
@@
 This means that our above figures are slightly liberal given that the referenced works may have comprehensively covered some of these criteria.
-Despite this, two-thirds of the studies did not reference previous work, and were missing important methodological details, suggesting the need for more comprehensive reporting of data collection and analysis methods within such studies.
+Despite this, two-thirds of the studies did not reference previous work, and were missing important methodological details, suggesting the need for more comprehensive reporting of methods.
 
@@
 
-todo: results here
+Of the `r nrow(coldat)` studies analysed, (PERCENTAGE)% publicly provided the raw underlying data, (PERCENTAGE)% provided data in a pre-processed form, and (PERCENTAGE)% of studies did not provide any publicly accessible data.
+The paucity of data being made available post-publication in studies of biological colouration is in line with other fields [@vines_mandated_2013; @wolkovich_advances_2012], though this trend is gradually shifting as funders and journals increasingly mandate the release of data [@whitlock_data_2010].
+The clearest benefit of open access to data is that it allows researchers to build upon previous work for the testing and refinement of new ideas and methods.
+This is particularly relevant to the field of colour-science, where new methods for analysing colour data are regularly developed [e.g.
+@allen_analyzing_2013; @endler_framework_2012; @stoddard_pattern_2014].
+The provision of open data may also foster collaborations as researchers draw upon existing work, and increased data sharing has been shown to positively correlate with citations [@piwowar_sharing_2007].
+Given the increasing rate of retractions [@steen_retractions_2010; @van_trouble_2011], there is also impetus for researchers to maintain credibility through transparency.
+We thus encourage researchers to publicly archive data in as raw a form as possible (e.g.
+individual reflectance spectra).
+Several excellent data repositories now exist, such as figshare and dryad, and the preparation of data is relatively straightforward [@whitlock_data_2011].
+The use of modest data embargoes and appropriate licenses can help to ensure that original authors are able to make the best use of data, and are subsequently credited [@neylon_science_2012].
 
-_Tools for reproducible research_
-
-Data - archiving (figshare, dryad), raw data preferable to processed data, concerns over data sharing, embargos are fine just get it archived.
-Code - benefits of code over point-and-click, open source shit (R, pavo, things), github, sourcefourge etc.
+None of the included studies linked to any form of code...benefits of code over point-and-click, open source shit (R, pavo, things), github, sourcefourge etc.
 
@@
 While we recognize that journals have space limitations and that work may reference previous methological and analytical details, we strongly believe that it is to the field's advantage that researchers report all details essential to a study's transparency and reproducibility wherever possible.
+
+The permanent private storage of data [or, more often, the complete loss of data; @alsheikh_public_2011] is increasingly difficult to justify, and the benefits of open data are considerable.
 

### 2014-12-05 docs/ms.Rmd
Editing the intro & methods outline. Begin tightening the table discussion

@@
-# Reproducible research in the study of colour in nature
+# Reproducible research in biological colour-science
 
@@
 These developments are positive for the field, and have allowed increasing numbers of researchers to rapidly collect and analyse ever greater amounts of data.
-The rapid rate of progression, however, means that a deep understanding of common methods is increasingly beyond the grasp of the average researcher.
+This rate of progression, however, means that a deep understanding of common methods is increasingly beyond the grasp of the average researcher.
 As a consequence, the subtle complexity of many analytical techniques [e.g.
@@
 Replication and transparency lie at the heart of the scientific enterprise.
-The full replication of individual studies may often be infeasible @peng_reproducible_2011, but reproducibility - the re-analysis of a study's data by independent researchers - has been presented as a readily attainable alternative [@peng_reproducible_2006; @peng_reproducible_2011].
+While the full replication of individual studies may often be infeasible [@peng_reproducible_2011], the reproductioin of a study's findings by independent researchers is a readily attainable compromise [@peng_reproducible_2006; @peng_reproducible_2011].
 A study may be considered truly reproducible when it satisfies three broad criteria: (i) methods are reported completely, (ii) data are publicly archived and available, and (iii) analysis code is made publicly available, such that the chain of modification of raw data is documented and preserved.
@@
 
-Our aim here to explore current levels of reproducibility in studies of biological colouration, and to outline ways in which they may be improved in line with points (i) - (iii) above.
-We begin by briefly outlining some of the most common methods for studying biological colouration.
+Our aim here to explore the state of reproducibility in studies of biological colouration, and to outline ways in which it may be improved in line with points (i) - (iii) above.
+We begin by briefly outlining some of the most common methods for studying biological colouration, and the way in which....
 We then present a set of guidelines for the comprehensive reporting of methods, and we analyse a subset of the literature against these criteria to assess the current state of methods reporting in colour science.
-We also assess the availability of publicly archived data and code, and suggest some tools that may be useful for increasing the reproducibility of colour trait research.
+We also assess the availability of publicly archived data and code, and suggest some useful tools for increasing the reproducibility of colour trait research.
 
-## Measuring and analysing colour
+## Researcher degrees of freedom in colour analysis
 
@@
 The application of spectrometry to questions of biological colouration revolutionized the field [@dyck_determination_1966].
-Now relatively inexpensive and convenient, spectrometry allows the capture of quantitative and arguably repeatable data and has been rapidly taken up by researchers the field.
+Now relatively inexpensive and convenient, spectrometry allows for the rapid capture of quantitative data, and has been widely adopted as a standard in the field [@andersson_quantifying_2006].
+
 However one drawback to this methodology, is the ease with which erroneous data can be collected if attention is not paid to finer details of spectrum capture.
@@
 We have kept the papers used in our dataset anonymous as our aim is to make a general statement about methods reporting in the field.
+Our search returned a total of 216 papers, however, only `r nrow(coldat)` met our criteria and were included in the final analysis.
 
@@
 
-todo: expand table discussion
-
-We developed a list of information we deemed critical to be reported for reproducibility of colour methods (Table 1) and analysis (Table 2), loosely based on Andersson and Prager [-@andersson_quantifying_2006] and Montgomerie [@montgomerie_analyzing_2006].
-These included quantitative methods including spectrometric and photographic techniques [as in @kemp_ornamental_2008; @chiao_visualization_2009] but not qualitative methods such as the use of Munsell chips or paint swatches [as in @hogstad_gloger_2009].
-For analytical techniques we characterized a set of criteria required for reproducing the analysis (Table 2).
-Much of these criteria specifically targeted the necessary parameters required to be explicitly defined by the user during analysis.
+The simplest and most fundamental step in ensuring research is reproducible is through the comprehensive reporting of methods.
+The myriad researcher degrees of freedom in the capture and analyses of colour data (as outlined above) heighten this need.
+Accordingly, we developed a list of essential information pertaining to the capture (Table 1) and analysis (Table 2) of data that should be reported to ensure reproducibility.
+Given the current state of the field, we focus on spectrometry and photography for the capture of data [as in @kemp_ornamental_2008; @chiao_visualization_2009], and do not consider qualitative methods such as the use of Munsell chips or paint swatches [as in @hogstad_gloger_2009].
+Analytical techniques are diverse [kelber_spectral_2010], so we focus on two broad, common categories of analysis: colourimetric (or 'spectral') analyses, and visual modelling (Table 2).
+Extra little bit about vis models in particular.
 These are essential if such analyses are to be reproduced.
 These criteria provided that baseline against which to hold our results.
-Our search returned a total of 216 papers, however, only `r nrow(coldat)` met our criteria and were included in the final analysis.
 
@@
 Of the `r nrow(coldat)` studies analysed, (PERCENTAGE)% publicly provided the raw underlying data, (PERCENTAGE)% provided data in a pre-processed form, and (PERCENTAGE)% of studies did not provide any publicly accessible data.
-The paucity of data being made available post-publication in studies of biological colouration is in line with other fields [@vines_mandated_2013; @wolkovich_advances_2012], though this trend is gradually shifting as funders and journals increasingly mandate the release of data [@whitlock_data_2010].
+The paucity of data being made available post-publication in studies of biological colouration is in line with other fields [@vines_mandated_2013; @wolkovich_advances_2012], including the broader field of animal behaviour [@caetano_forgotten_2014].
+There is evidence that this trend is gradually shifting as funders and journals increasingly mandate the release of data [@whitlock_data_2010].
 The clearest benefit of open access to data is that it allows researchers to build upon previous work for the testing and refinement of new ideas and methods.
-This is particularly relevant to the field of colour-science, where new methods for analysing colour data are regularly developed [e.g.
+This is particularly relevant to the field of colour-science, where new methods for analysing colour data are frequently developed [e.g.
 @allen_analyzing_2013; @endler_framework_2012; @stoddard_pattern_2014].
@@
 individual reflectance spectra).
-Several excellent data repositories now exist, such as figshare and dryad, and the preparation of data is relatively straightforward [@whitlock_data_2011].
+Several excellent data repositories now exist, such as figshare and dryad, and the preparation of data for publication is relatively straightforward [@whitlock_data_2011].
 The use of modest data embargoes and appropriate licenses can help to ensure that original authors are able to make the best use of data, and are subsequently credited [@neylon_science_2012].
 
-None of the included studies linked to any form of code...benefits of code over point-and-click, open source shit (R, pavo, things), github, sourcefourge etc.
+(PARA) None of the included studies linked to any form of code...benefits of code over point-and-click, open source shit (R, pavo, things), github, sourcefourge etc.
 
@@
 
-The permanent private storage of data [or, more often, the complete loss of data; @alsheikh_public_2011] is increasingly difficult to justify, and the benefits of open data are considerable.
+The permanent private storage of data [or, more often, the complete loss of data; @alsheikh_public_2011] is increasingly difficult to justify, and the benefits of open data are considerable....
 

### 2014-12-06 docs/ms.Rmd
Mostly tightening the intro. I almost like it.

@@
 The resurgent interest in biological colouration has in large part been driven by the increased availability of key technologies, including spectrometry and photography, and concurrent advances in methods for analysing colour data, such as visual models [@endler_comparing_2005; @stevens_using_2007; @kelber_animal_2003].
-These developments are positive for the field, and have allowed increasing numbers of researchers to rapidly collect and analyse ever greater amounts of data.
-This rate of progression, however, means that a deep understanding of common methods is increasingly beyond the grasp of the average researcher.
-As a consequence, the subtle complexity of many analytical techniques [e.g.
-the 3-7 parameters that define a typical visual model; @endler_comparing_2005; @vorobyev_receptor_1998] is often overlooked by empiricists, which leads to critical methodological information often not being reported.
+While these developments are positive for the field, the increasingly complex analyses being run on ever greater amounts of data heighten the need for comprehensive methods reporting and proper data management [@alsheikh_public_2011; @nekrutenko_next_2012].
 
 Replication and transparency lie at the heart of the scientific enterprise.
-While the full replication of individual studies may often be infeasible [@peng_reproducible_2011], the reproductioin of a study's findings by independent researchers is a readily attainable compromise [@peng_reproducible_2006; @peng_reproducible_2011].
+Beyond simply allowing the independent verification of results, reproducible research engenders greater comparability between studies and provides a foundation for the testing of new ideas and methods [@whitlock_data_2011; @van_trouble_2011; @piwowar_sharing_2007].
 A study may be considered truly reproducible when it satisfies three broad criteria: (i) methods are reported completely, (ii) data are publicly archived and available, and (iii) analysis code is made publicly available, such that the chain of modification of raw data is documented and preserved.
@@
 @fitzjohn_much_2014] is a laudable goal, the considerable demands it imposes on researchers means that it will often, in practice, be unattainable.
-Nevertheless, even partial reproducibility through the simple practice of complete methods reporting and public data archiving is of tremendous value.
-It ensures greater comparability between studies, allows for the the independent verification of results, and provides a foundation for the testing of new ideas and methods [@whitlock_data_2011; @van_trouble_2011; @piwowar_sharing_2007].
+Nevertheless, even partial reproducibility through the relatively simple practices of complete methods reporting and public data archiving is of tremendous value.
 
 Our aim here to explore the state of reproducibility in studies of biological colouration, and to outline ways in which it may be improved in line with points (i) - (iii) above.
-We begin by briefly outlining some of the most common methods for studying biological colouration, and the way in which....
+We begin by briefly outlining some of the most common methods for studying biological colouration, with an emphasis on the way in which their subtle complexity can dramatically shape results.
 We then present a set of guidelines for the comprehensive reporting of methods, and we analyse a subset of the literature against these criteria to assess the current state of methods reporting in colour science.
@@
 
-Although relatively easy to implement, and potentially powerful, sensory models are built on multiple assumptions about the way in which stimuli are visually processed, that can dramatically shape the results of a given analysis [@lind_avian_2009; @pike_preserving_2012].
+Although relatively easy to implement, and potentially powerful, visual models are built on multiple assumptions about the way in which stimuli are visually processed that can dramatically shape the results of a given analysis [@lind_avian_2009; @pike_preserving_2012].
 For example, Vorobyev et al.’s [-@vorobyev_receptor_1998; -@vorobyev_colour_2001] widely used receptor-noise model requires researchers to specify the form of receptor quantum catches (i.e.
-raw versus log-transformed), the use of any colour constancy mechanism (e.g.
+raw versus transformed), the use of any colour constancy mechanism (e.g.
 von Kries), as well as the photoreceptor signal-to-noise ratios, their relative densities, and the type of noise predominating (i.e.
@@
 We then tabulated what details of the methods were reported against our criteria.
-The data along with analysis script can be found at: https://github.com/daniel1noble/colsci_rep (TODO: GIVE IT A DOI?).
-We have kept the papers used in our dataset anonymous as our aim is to make a general statement about methods reporting in the field.
+The data along with analysis script can be found on github (WHAT'S THE DOI?).
+We have kept the papers used in our dataset anonymous as our aim is to explore the general question of repeatability in the field.
 Our search returned a total of 216 papers, however, only `r nrow(coldat)` met our criteria and were included in the final analysis.
@@
 
-_Guidelines for methods reporting_
+_Guidelines for reporting_
 
-The simplest and most fundamental step in ensuring research is reproducible is through the comprehensive reporting of methods.
+The simplest step in ensuring research is reproducible is the comprehensive reporting of methods.
 The myriad researcher degrees of freedom in the capture and analyses of colour data (as outlined above) heighten this need.
-Accordingly, we developed a list of essential information pertaining to the capture (Table 1) and analysis (Table 2) of data that should be reported to ensure reproducibility.
-Given the current state of the field, we focus on spectrometry and photography for the capture of data [as in @kemp_ornamental_2008; @chiao_visualization_2009], and do not consider qualitative methods such as the use of Munsell chips or paint swatches [as in @hogstad_gloger_2009].
-Analytical techniques are diverse [kelber_spectral_2010], so we focus on two broad, common categories of analysis: colourimetric (or 'spectral') analyses, and visual modelling (Table 2).
-Extra little bit about vis models in particular.
+Accordingly, we developed a list of essential information about the capture (Table 1) and analysis (Table 2) of data that should be reported to ensure reproducibility.
+Given their overwhelming popularity, we focus on spectrometry and photography for data collection [e.g.
+@kemp_ornamental_2008; @chiao_visualization_2009], and do not consider qualitative methods such as the use of Munsell chips or paint swatches [e.g.
+@hogstad_gloger_2009].
+
+Analytical techniques are diverse and are being developed rapidly [kelber_spectral_2010].
+This rate of progression, however, means that a deep understanding of common methods is increasingly beyond the grasp of the average researcher.
+As a consequence, the subtle complexity of many analytical techniques [e.g.
+the 3-7 parameters that define a typical visual model; @endler_comparing_2005; @vorobyev_receptor_1998] is often overlooked by empiricists, which leads to critical methodological information often not being reported.
+Our guidelines for reporting analytical details (Table 2) cover two broad, common categories of analysis: colourimetric (or 'spectral') analyses, and visual modelling.
 These are essential if such analyses are to be reproduced.
@@
 
-Most studies using visual models analysed their data using receptor-noise models (`r vismod_type[3]` of `r sum(vismod_type)` studies), while others used the colour hexagon (`r vismod_type[1]` studies), or tetrahedral colour space (`r vismod_type[4]` studies) approach.
+Most studies using visual models analysed their data using receptor-noise models (`r vismod_type[3]` of `r sum(vismod_type)` studies), while others used the colour hexagon (`r vismod_type[1]` studies), or a tetrahedral colour space (`r vismod_type[4]` studies).
 Case specific models were used in `r vismod_type[2]` papers.
@@
 
-(PARA) None of the included studies linked to any form of code...benefits of code over point-and-click, open source shit (R, pavo, things), github, sourcefourge etc.
+None of the included studies linked to any form of code, which is not surprising given the ongoing popularity of graphical-user-interface based statistical [e.g.
+SPSS], and colour-analytical [e.g.
+avicol;] software.
+Benefits of code over point-and-click, maintain chain of mod for yourself and others, open source shit (R, pavo, things), github, sourcefourge etc.
+This paper was written openly on github,
 

### 2014-12-09 docs/ms.Rmd
Rewrote a bunch of shit

@@
-# Reproducible research in biological colour-science
+# Reproducible research in the study of biological colouration
 
@@
 
-^1^ Evolution & Ecology Research Centre, School of Biological, Earth and Environmental Sciences, University of New South Wales, Kensington, NSW, Australia, 2022 ^2^ Department of Biological Sciences, Macquarie University, North Ryde, NSW, Australia, 2109 ^3^ School of Biological Sciences, University of Wollongong, Wollongong, NSW, Australia 2252 ^4^ Centre for Evolutionary Biology, University of Western Australia, Crawley, WA, Australia 6008 ^5^ Department of Biological Sciences, University of Pittsburgh, Pittsburgh, PA, USA, 15260
+^1^ Evolution & Ecology Research Centre, School of Biological, Earth and Environmental Sciences, University of New South Wales, Kensington, NSW, Australia, 2022 ^2^ Department of Biological Sciences, Macquarie University, North Ryde, NSW, Australia, 2109 ^3^ School of Biological Sciences, University of Wollongong, Wollongong, NSW, Austrlaia 2252 ^4^ Centre for Evolutionary Biology, University of Western Australia, Crawley, WA, Australia 6008 ^5^ Department of Biological Sciences, University of Pittsburgh, Pittsburgh, PA, USA, 15260
 
@@
 
-Keywords: #colsci, spectrometry, photography, visual ecology, pigment, structural colour, visual modelling, methodology
+Keywords: #colsci, spectrometry, photography, visual ecology, pigment, structural colour, visual modelling, methods
 
@@
 
-Studies of biological colouration have generated insights into fundamental evolutionary and ecological processes, and research into colour traits is a rapidly growing field [@kelber_spectral_2010].
+The study of colour in nature has generated insights into fundamental evolutionary and ecological processes, and research into colour traits is a rapidly growing field [@kelber_spectral_2010].
 The resurgent interest in biological colouration has in large part been driven by the increased availability of key technologies, including spectrometry and photography, and concurrent advances in methods for analysing colour data, such as visual models [@endler_comparing_2005; @stevens_using_2007; @kelber_animal_2003].
@@
 Beyond simply allowing the independent verification of results, reproducible research engenders greater comparability between studies and provides a foundation for the testing of new ideas and methods [@whitlock_data_2011; @van_trouble_2011; @piwowar_sharing_2007].
-A study may be considered truly reproducible when it satisfies three broad criteria: (i) methods are reported completely, (ii) data are publicly archived and available, and (iii) analysis code is made publicly available, such that the chain of modification of raw data is documented and preserved.
+A study may be considered truly reproducible when it satisfies three broad criteria: (i) methods are reported completely, (ii) data are publicly available and archived, and (iii) analysis code is publicly available, such that the chain of modification of raw data is documented and preserved.
 While completely reproducible research [e.g.
@@
 Our aim here to explore the state of reproducibility in studies of biological colouration, and to outline ways in which it may be improved in line with points (i) - (iii) above.
-We begin by briefly outlining some of the most common methods for studying biological colouration, with an emphasis on the way in which their subtle complexity can dramatically shape results.
-We then present a set of guidelines for the comprehensive reporting of methods, and we analyse a subset of the literature against these criteria to assess the current state of methods reporting in colour science.
-We also assess the availability of publicly archived data and code, and suggest some useful tools for increasing the reproducibility of colour trait research.
+We begin by briefly outlining some of the most common methods for studying biological colouration, with an emphasis on the way in which their subtle complexity can shape results.
+We then present a set of guidelines for the comprehensive reporting of methods, and we analyse a subset of the literature against these criteria to assess the state of methods reporting in biological colour-science.
+We also assess the availability of publicly archived data and code, and suggest some useful tools for increasing the reproducibility of colour trait research more broadly.
 
-## Researcher degrees of freedom in colour analysis
+## Measuring colour
 
-todo: hack this back, make it more relevant to reproducibility
-
-Generations of biologists have endeavoured to explain the mechanisms and functions of animal and plant colouration [@thayer_concealing_1909; @poulton_colours_1890; @wallace_natural_1891], and identifying best practices in measuring colour has been a great challenge.
+Generations of biologists have endeavoured to explain the mechanisms and functions of animal and plant colouration [@thayer_concealing_1909; @poulton_colours_1890; @wallace_natural_1891], and uncovering best practices in measuring colour has been a great challenge.
 The application of spectrometry to questions of biological colouration revolutionized the field [@dyck_determination_1966].
 Now relatively inexpensive and convenient, spectrometry allows for the rapid capture of quantitative data, and has been widely adopted as a standard in the field [@andersson_quantifying_2006].
+Digital photography is also increasingly being used to quantify biological colouration [@stevens_using_2007], as high-resolution cameras are inexpensive and allow for the rapid, repeatable sampling of multiple colour patches [@mckay_use_2013].
 
-However one drawback to this methodology, is the ease with which erroneous data can be collected if attention is not paid to finer details of spectrum capture.
-The angle from which spectra are taken [@santos_strong_2007], consistent calibration of light and dark standards, the distance between light source and surface, and the size, shape, glossiness or iridescence of the specimen can all contribute to variation in data and potentially bias results [@andersson_quantifying_2006; @kemp_ornamental_2008; @meadows_quantifying_2011; @schaefer_birds_2008].
-Experience, guidance, consistency and objectivity (e.g.
-double-blind experimental design) are necessary to ensure data are repeatable and representative of the subject.
-Crucially, to ensure repeatability across studies, the conditions under which data were collected must be reported exactly.
-
-Complementary to spectrometry, digital photography is increasingly being used to quantify biological colouration [@stevens_using_2007].
-High-resolution cameras are inexpensive,  and allow instantaneous, repeatable sampling of multiple colour patches [@mckay_use_2013].
-The main challenge for gathering colour data from photographs involve ensuring standardization, such as linearisation and consistent lighting [@stevens_using_2007].
+Subtle variation in the application of these common methods can significantly shape the resulting data.
+When using spectrometry, the angle from which spectra are collected [@santos_strong_2007], the calibration of light and dark standards, the distance between light source and surface, and the size, shape, glossiness or iridescence of the specimen will all directly affect the recorded spectra [@andersson_quantifying_2006; @kemp_ornamental_2008; @meadows_quantifying_2011; @schaefer_birds_2008].
+The main challenge for gathering colour data from photographs is ensuring standardization against nonlinear responses to light intensity, biases towards particular wavebands, and inconsistent lighting [@stevens_using_2007].
 Also, most digital cameras used for the measurement of biological colouration are consumer products designed for the human visual system.
-Thus hardware modifications for ultraviolet and infra-red imaging are generally necessary [e.g.
+Thus hardware modifications for ultraviolet and infra-red imaging are often necessary [e.g.
 @stevens_colour_2013].
@@
 Expansion in the availability of objective methods for the measurement of colour has been matched by advances in theory and analysis.
-In particular, the development of visual models has produced significant insights into the early visual processing of colourful stimuli [@chittka_colour_1992; @endler_comparing_2005; @vorobyev_receptor_1998] and, in a few cases, how they may be perceived [@kelber_spectral_2010].
-Visual models typically attempt to describe the subjective perception of colour information as a function of an object’s reflectance, the ambient illumination, and a receiver’s sensory system [e.g.
+In particular, the development of visual models has enabled researchers to move beyond purely quantitative comparisons of reflectance spectra and adopt more biologically relevant perspectives when defining and testing hypotheses [@chittka_colour_1992; @endler_comparing_2005; @vorobyev_receptor_1998].
+Visual models typically attempt to describe the reception and early-stage processing of colour information as a function of an object’s reflectance, the ambient illumination, and a receiver’s sensory system [e.g.
 @endler_comparing_2005; @vorobyev_receptor_1998].
-These models enable researchers to move beyond purely quantitative comparisons of reflectance spectra and adopt more biologically relevant perspectives to refine and test hypotheses.
-
-Although relatively easy to implement, and potentially powerful, visual models are built on multiple assumptions about the way in which stimuli are visually processed that can dramatically shape the results of a given analysis [@lind_avian_2009; @pike_preserving_2012].
+Although relatively easy to implement, visual models are built on multiple assumptions about the way in which stimuli are processed that can dramatically shape the results of a given analysis [@lind_avian_2009; @pike_preserving_2012].
 For example, Vorobyev et al.’s [-@vorobyev_receptor_1998; -@vorobyev_colour_2001] widely used receptor-noise model requires researchers to specify the form of receptor quantum catches (i.e.
@@
 
-The potential for subjectivity in colour measurement and the mathematical opacity of visual models for the average biologist means that, as in any field, is it critical that methods are reported rigorously.
+Given the myriad degrees of freedom that exist when measuring and analysing colour, and the significant consequences of any variation, the conditions under which data were collected must be reported exactly....still working on this...
 Methods reporting should ideally have well defined standards such that sufficient detail is provided to allow comprehensive peer review, precise replication, cross-study comparison and data sharing [@alsheikh_public_2011; @sandve_ten_2013]....
@@
 
-To assess the current state of reproducibility in the field we searched papers from 2013 in 22 leading journals (_American Journal of Botany_, _The American Naturalist_, _Animal Behaviour_, _Behavioral Ecology_, _Behavioral Ecology and Sociobiology_, _Biological Journal of the Linnean Society_, _Biology Letters_, _Current Zoology_, _Ecology_, _Ecology and Evolution_, _Ecology Letters_, _Ethology_, _Evolution_, _Functional Ecology_, _Journal of Ecology_, _The Journal of Evolutionary Biology_, _The Journal of Experimental Biology_, _Naturwissenschaften_, _New Phytologist_, _Oikos_, _PLoS ONE_, _Proceedings of the Royal Society B: Biological Sciences_).
-On each of the journals' homepages, we used the Boolean terms 'colour*', 'color*' or 'spectra*' to search the title or abstract.
-We read the titles and abstracts of each paper and included only those papers that used either a spectrometer or camera to quantify colouration.
-We excluded papers review papers, methodological papers, papers quantifying colour patterns rather than the chromatic properties of a colour patch, and papers taking micro-spectrometric measurements of retinal absorbance.
+To assess the current state of reproducibility in the field we searched papers from 2013 in 22 leading journals: _American Journal of Botany_, _The American Naturalist_, _Animal Behaviour_, _Behavioral Ecology_, _Behavioral Ecology and Sociobiology_, _Biological Journal of the Linnean Society_, _Biology Letters_, _Current Zoology_, _Ecology_, _Ecology and Evolution_, _Ecology Letters_, _Ethology_, _Evolution_, _Functional Ecology_, _Journal of Ecology_, _The Journal of Evolutionary Biology_, _The Journal of Experimental Biology_, _Naturwissenschaften_, _New Phytologist_, _Oikos_, _PLoS ONE_, _Proceedings of the Royal Society B: Biological Sciences_.
+On each of the journals' homepages, we used the Boolean terms 'colour\*', 'color\*' or 'spectra\*' to search the title or abstract.
+We included only those papers that used either a spectrometer or camera to quantify colouration.
+We excluded review papers, methodological papers, papers quantifying colour patterns rather than the chromatic properties of a colour patch, and papers taking micro-spectrometric measurements of retinal absorbance.
 To review the large number of papers returned from our search, journals were haphazardly divided up between the authors.
-In order to reduce the risk of observer bias in our assesment, each paper was read and reassessed by three authors.
+In order to reduce the risk of observer bias in our assessment, each paper was read and reassessed by three authors.
 Any discrepancies between assessor scores were discussed and resolved prior to analysis.
@@
 
-_Guidelines for reporting_
+_Guidelines_
 
-The simplest step in ensuring research is reproducible is the comprehensive reporting of methods.
-The myriad researcher degrees of freedom in the capture and analyses of colour data (as outlined above) heighten this need.
-Accordingly, we developed a list of essential information about the capture (Table 1) and analysis (Table 2) of data that should be reported to ensure reproducibility.
-Given their overwhelming popularity, we focus on spectrometry and photography for data collection [e.g.
-@kemp_ornamental_2008; @chiao_visualization_2009], and do not consider qualitative methods such as the use of Munsell chips or paint swatches [e.g.
-@hogstad_gloger_2009].
-
-Analytical techniques are diverse and are being developed rapidly [kelber_spectral_2010].
-This rate of progression, however, means that a deep understanding of common methods is increasingly beyond the grasp of the average researcher.
-As a consequence, the subtle complexity of many analytical techniques [e.g.
-the 3-7 parameters that define a typical visual model; @endler_comparing_2005; @vorobyev_receptor_1998] is often overlooked by empiricists, which leads to critical methodological information often not being reported.
-Our guidelines for reporting analytical details (Table 2) cover two broad, common categories of analysis: colourimetric (or 'spectral') analyses, and visual modelling.
-These are essential if such analyses are to be reproduced.
-These criteria provided that baseline against which to hold our results.
+The comprehensive reporting of methods is the simplest step in ensuring research is reproducible.
+Accordingly, we developed a list of essential information about the capture (Table 1) and analysis (Table 2) of colour data that should ideally be reported.
+Given their overwhelming popularity, our data collection reporting guidelines are focused on spectrometry and photography for data collection (Table 1), and do not consider qualitative methods such as the use of Munsell chips or paint swatches.
+Analytical techniques are diverse and are being developed rapidly [@kelber_spectral_2010], and this rate of progress means that a deep understanding of common methods is increasingly beyond the grasp of the average researcher.
+As a consequence, the subtle complexity of many analytical techniques is often overlooked by empiricists, which leads to critical methodological information often not being reported.
+Our guidelines for reporting the details of colour analyses (Table 2) thus cover two broad, common categories of analysis: colourimetric (or 'spectral') analyses, and visual modelling.
+These tables present the framework for our subsequent analysis of methods reporting in the literature.
 
@@
 This means that our above figures are slightly liberal given that the referenced works may have comprehensively covered some of these criteria.
-Despite this, two-thirds of the studies did not reference previous work, and were missing important methodological details, suggesting the need for more comprehensive reporting of methods.
+Despite this, two-thirds of the studies did not reference previous work, and were missing important methodological details, suggesting the need for more comprehensive reporting of methods...touch more discussion here, instead of in conclusions(?)...e.g...When possible, we also recommend that authors also justify data capture and analytical approaches - something few studies we investigated explicitly outlined.
 
@@
 Of the `r nrow(coldat)` studies analysed, (PERCENTAGE)% publicly provided the raw underlying data, (PERCENTAGE)% provided data in a pre-processed form, and (PERCENTAGE)% of studies did not provide any publicly accessible data.
-The paucity of data being made available post-publication in studies of biological colouration is in line with other fields [@vines_mandated_2013; @wolkovich_advances_2012], including the broader field of animal behaviour [@caetano_forgotten_2014].
-There is evidence that this trend is gradually shifting as funders and journals increasingly mandate the release of data [@whitlock_data_2010].
+The paucity of data being made available post-publication in studies of biological colouration is in line with other fields [@drew_lost_2013; @vines_mandated_2013; @wolkovich_advances_2012], including the broader field of animal behaviour [@caetano_forgotten_2014].
+There is evidence, however, that this trend is shifting as funders and journals increasingly mandate the release of data [@whitlock_data_2010].
 The clearest benefit of open access to data is that it allows researchers to build upon previous work for the testing and refinement of new ideas and methods.
-This is particularly relevant to the field of colour-science, where new methods for analysing colour data are frequently developed [e.g.
+This is particularly relevant in the study of colour traits, where new methods for analysing colour data are regularly developed [e.g.
 @allen_analyzing_2013; @endler_framework_2012; @stoddard_pattern_2014].
-The provision of open data may also foster collaborations as researchers draw upon existing work, and increased data sharing has been shown to positively correlate with citations [@piwowar_sharing_2007].
-Given the increasing rate of retractions [@steen_retractions_2010; @van_trouble_2011], there is also impetus for researchers to maintain credibility through transparency.
+The provision of open data may also foster collaborations as researchers draw upon existing work [@whitlock_data_2011], and increased data sharing has been shown to positively correlate with citations [@piwowar_sharing_2007].
+Given the increasing rate of retractions, there is also impetus for researchers to maintain credibility through transparency [@steen_retractions_2010; @van_trouble_2011; @wicherts_willingness_2011].
 We thus encourage researchers to publicly archive data in as raw a form as possible (e.g.
-individual reflectance spectra).
-Several excellent data repositories now exist, such as figshare and dryad, and the preparation of data for publication is relatively straightforward [@whitlock_data_2011].
-The use of modest data embargoes and appropriate licenses can help to ensure that original authors are able to make the best use of data, and are subsequently credited [@neylon_science_2012].
+individual reflectance spectra), to maximise its utility.
+Public data repositories such as figshare (REF?) and dryad (REF?) facilitate the permanent archival of data as citable research objects, and the preparation of data for publication is often straightforward [@whitlock_data_2011].
+The use of modest data embargoes and appropriate licenses can help to ensure that original authors are able to make the best use of data, and are subsequently credited [though we acknowledge that tensions over data-sharing exist; discussed in @roche_troubleshooting_2014].
 
-None of the included studies linked to any form of code, which is not surprising given the ongoing popularity of graphical-user-interface based statistical [e.g.
-SPSS], and colour-analytical [e.g.
-avicol;] software.
-Benefits of code over point-and-click, maintain chain of mod for yourself and others, open source shit (R, pavo, things), github, sourcefourge etc.
-This paper was written openly on github,
+None of the included studies linked to any form of code, which is not surprising given the popularity of graphical-user-interface (GUI) based statistical, and colour-analytical software [e.g.
+AVICOL; @gomez_avicol_2006].
+Here we simply note that there are several advantages to moving from a GUI-based to a programming workflow, in spite of the initial time investment required.
+The open-source programming language R [@R_program], for example, is free and is host to a sizeable community of developers that build and implement analytical tools (via 'packages').
+Indeed, packages specifically for the analysis of colour data already exist [e.g.
+pavo; @maia_pavo_2013].
+Perhaps the most significant advantage is that analysis scripts produced in such a way represent a complete documented history of a study's analytical methods.
+For authors this means that the details of a study's methods may be precisely preserved, and it allows other researchers to directly build upon.....
+
+........greater speed and better control...documenting of methods & maintaining chain of modification for reproducibility...something about git facilitating collaboration & project management [@ram_git_2013].
+Maybe rmarkdown [@rmarkdown]
 
@@
 
-Our review of recent colour literature highlights some important reporting gaps that will hinder the reproducibility and comparability of studies of biological colouration (Fig.
+Our review of the recent colour literature highlights some significant impediments to reproducibility.
+The simplest step towards reproducible research is through the comprehensive reporting of methods, yet crucial methodological information is often unreported (Fig.
 1).
-While we recognize that journals have space limitations and that work may reference previous methological and analytical details, we strongly believe that it is to the field's advantage that researchers report all details essential to a study's transparency and reproducibility wherever possible.
+In particular, we found that the types of light sources, numbers of pixels or spectra averaged, the distance and angle of the spectrometers optical probe relative to the measurement surface and collector, and types of standards used (i.e.
+white and black reference samples) are under-reported (Fig.
+1).
+Visual models were regularly used to analyse spectral data, yet key details such as the form of quantum catch used, whether photoreceptor adaptation was modelled, and the type of noise used in receptor-noise models were often not reported (Fig.
+1).
 
-The permanent private storage of data [or, more often, the complete loss of data; @alsheikh_public_2011] is increasingly difficult to justify, and the benefits of open data are considerable....
+The permanent, private storage of data [or, more often, the complete loss of data; @alsheikh_public_2011] is increasingly difficult to justify, and the benefits of publicly archived data are considerable [@piwowar_sharing_2007; @vision_open_2010; @whitlock_data_2010].
+Most of the data underlying studies of biological colouration are unpublished, and are thus likely to be lost...
 
-While all criteria are important, some deserve special attention.
-In particular, we found that the types of light sources, numbers of pixels or spectra averaged, and types of standards used (i.e.
-white and black reference samples) are under-reported, each of which is critical to repeatability and cross-comparison.
-Similarly, the distance and angle of the optical probe relative to the measurement surface and collector should be standard reporting for all studies using spectrometry, as variation in these factors can lead to dramatically different results in some cases (REFS).
-Visual models were regularly used to analyse spectral data, however, there were many important aspects of the analysis that were seldom reported.
-These included the type of quantum catch used, whether photoreceptor adaptation was modelled and the type of noise used in receptor-noise models.
-This is problematic because...
-Also, photoreceptor densities should be reported where possible even if using standard avian photoreceptor densities because....
-These all have important consequences on downstream analysis and the reproducibility of colour signal quantification [e.g.
-@kemp_ornamental_2008; @santos_comparison_2007].
-We also recommend the presentation of spectral reflectance curves where possible.
-These figures allow for the rapid assessment of the nature and quality of the data, independent of downstream processing (if the conditions under which they were captured is comprehensively reported.
-We found that only `r round(propSpec_fin["prop_refl_fig"], digits = 2)`% of studies using spectral reflectance data had a figure showing these curves.
-When possible, we also recommend that authors justify data capture and analytical approaches - something few studies we investigated explicitly outlined.
-
-Our objective in this paper was to highlight common oversights in reporting that can influence the reproducibility of colour research and cross-study comparisons.
-By no means do we wish to say that studies lacking important details are not of significant value, but rather to provide recommendations for future work.
-It is only through critical appraisal and collegial discussion that we can identify areas for improvement, and constructively contribute to the growth of our field.
-
-We believe that part of the issue at hand is simply that many researchers seldom have clear guidelines to follow when reporting methods, and may overlook outlining important aspects of methodology and analysis.
-We hope our critical evaluation of recent literature has highlighted a need for more careful description of how colour signals have been measured and analysed.
-More importantly, we hope our guidelines for methods reporting helps establish a standard in this area, to continue the advance of an exciting era in colour research.
+The move towards increasingly reproducible research has many benefits.
+Our objective in this paper was to explore the issue of reproducibility in studies of biological colouration, and to offer guidelines on simple ways in which studies may be leveraged to their full potential.
+We hope our that our discussion of methods reporting and data archiving will contribute to the growing culture of reproducibility in science, and that researchers will consider these issues when deciding on the best practices in their work.
+The future of colour-trait research is bright (obligatory pun?!)...something something maybe...continuing the advance of an exciting field.
 

### 2014-12-10 docs/ms.Rmd
Updated some mnor bugs in the analysis script which prevented the proper caluclation of some percentages. Also added into analysis the new code and data deposition data collected by Tom.

@@
 
-Of the `r nrow(coldat)` studies analysed, (PERCENTAGE)% publicly provided the raw underlying data, (PERCENTAGE)% provided data in a pre-processed form, and (PERCENTAGE)% of studies did not provide any publicly accessible data.
+Of the `r nrow(coldat)` studies analysed, `r round((data["raw"]/sum(data))*100, digits = 2)`% publicly provided the raw underlying data, `round((data["proc"]/sum(data))*100, digits = 2)`% provided data in a pre-processed form, and `r round((data["none"]/sum(data))*100, digits = 2)`% of studies did not provide any publicly accessible data.
 The paucity of data being made available post-publication in studies of biological colouration is in line with other fields [@drew_lost_2013; @vines_mandated_2013; @wolkovich_advances_2012], including the broader field of animal behaviour [@caetano_forgotten_2014].

### 2014-12-10 R/analysis.r

@@ -64,6 +64,8 @@ propCam_fin  <- apply(DatCam[sapply(DatCam, is.numeric)], 2, function(x) sum(x[!
 
 lengthCam_fin <- apply(DatCam[sapply(DatCam, is.numeric)], 2, function(x) length(x[! is.na(x)]))[criteriaCam]
 
+names(propCam_fin) <- paste("prop_", names(propCam_fin), sep ="")
+
 ##--------------------------------------
 ## How do studies analyse colour?
 analysisType <- table(coldat$analysis_type)
@@ -83,6 +85,10 @@ vismod_type <- table(DatSpec$vis_mod_type)
 
 ## Number of studies referencing other work.
 ref_stud <- (sum(coldat$prev_pub)/length(coldat$prev_pub))*100
+
+##Code and data deposition
+
+data <- table(coldat$data)
 
 ##------------------------------------ Figure 1--------------------------------------##
 # Figure of proportions on the hardware/software reported

### 2014-12-15 docs/ms.Rmd
Mostly integrated Noble Dan & Kate's changes, modify a few refs

@@
 
-Thomas E. White^2,6^, Rhiannon L. Dalrymple^1,6^, Daniel W. A. Noble^2,6^, James C. O’Hanlon ^2,6^, Daniel B. Zurek^5,6^, Kate D. L. Umbers^3,4,6^
+Thomas E. White^2,7^, Rhiannon L. Dalrymple^1,7^, Daniel W. A. Noble^2,7^, James C. O’Hanlon ^2,7^, Daniel B. Zurek^5,7^, Kate D. L. Umbers^3,4,6,7^
 
-^1^ Evolution & Ecology Research Centre, School of Biological, Earth and Environmental Sciences, University of New South Wales, Kensington, NSW, Australia, 2022 ^2^ Department of Biological Sciences, Macquarie University, North Ryde, NSW, Australia, 2109 ^3^ School of Biological Sciences, University of Wollongong, Wollongong, NSW, Austrlaia 2252 ^4^ Centre for Evolutionary Biology, University of Western Australia, Crawley, WA, Australia 6008 ^5^ Department of Biological Sciences, University of Pittsburgh, Pittsburgh, PA, USA, 15260
+^1^ Evolution & Ecology Research Centre, School of Biological, Earth and Environmental Sciences, University of New South Wales, Kensington, NSW, Australia, 2022 ^2^ Department of Biological Sciences, Macquarie University, North Ryde, NSW, Australia, 2109 ^3^ School of Biological Sciences, University of Wollongong, Wollongong, NSW, Austrlaia 2252 ^4^ Centre for Evolutionary Biology, University of Western Australia, Crawley, WA, Australia 6008 ^5^ Department of Biological Sciences, University of Pittsburgh, Pittsburgh, PA, USA, 15260 ^6^ School of Science and Health, University of Western Sydney Hawkesbury, NSW, Australia 2751
 
-^6^ Authors contributed equally and are presented in random order
+^7^ Authors contributed equally and are presented in random order
 
@@
 
-Keywords: #colsci, spectrometry, photography, visual ecology, pigment, structural colour, visual modelling, methods
+Thomas E. White, Department of Biological Sciences, Macquarie University, North Ryde, NSW, Australia, 2109
+
+Phone: +61 2 9850 6279
+
+E-mail: thomas.white@mq.edu.au
+
+Keywords: spectrometry, photography, visual ecology, pigment, structural colour, visual modelling, methods
 
@@
 The study of colour in nature has generated insights into fundamental evolutionary and ecological processes, and research into colour traits is a rapidly growing field [@kelber_spectral_2010].
-The resurgent interest in biological colouration has in large part been driven by the increased availability of key technologies, including spectrometry and photography, and concurrent advances in methods for analysing colour data, such as visual models [@endler_comparing_2005; @stevens_using_2007; @kelber_animal_2003].
+Resurgent interest in biological colouration has in large part been driven by the increased availability of key technologies, including spectrometry and photography, and concurrent advances in methods for analysing colour data, such as visual models [@endler_comparing_2005; @stevens_using_2007; @kelber_animal_2003].
 While these developments are positive for the field, the increasingly complex analyses being run on ever greater amounts of data heighten the need for comprehensive methods reporting and proper data management [@alsheikh_public_2011; @nekrutenko_next_2012].
@@
 Replication and transparency lie at the heart of the scientific enterprise.
-Beyond simply allowing the independent verification of results, reproducible research engenders greater comparability between studies and provides a foundation for the testing of new ideas and methods [@whitlock_data_2011; @van_trouble_2011; @piwowar_sharing_2007].
+Beyond simply allowing independent verification of results, reproducible research ensures greater comparability between studies and provides a foundation for testing new ideas and methods [@whitlock_data_2011; @van_trouble_2011; @piwowar_sharing_2007].
 A study may be considered truly reproducible when it satisfies three broad criteria: (i) methods are reported completely, (ii) data are publicly available and archived, and (iii) analysis code is publicly available, such that the chain of modification of raw data is documented and preserved.
@@
 
-Our aim here to explore the state of reproducibility in studies of biological colouration, and to outline ways in which it may be improved in line with points (i) - (iii) above.
+Our aim was to explore the state of reproducibility in studies of biological colouration, and to outline ways in which it may be optimised.
 We begin by briefly outlining some of the most common methods for studying biological colouration, with an emphasis on the way in which their subtle complexity can shape results.
-We then present a set of guidelines for the comprehensive reporting of methods, and we analyse a subset of the literature against these criteria to assess the state of methods reporting in biological colour-science.
+We then present a set of guidelines for the comprehensive reporting of methods, and we analyse a subset of the literature against these criteria to assess the state of methods reporting in studies of biological colouration.
 We also assess the availability of publicly archived data and code, and suggest some useful tools for increasing the reproducibility of colour trait research more broadly.
@@
 The application of spectrometry to questions of biological colouration revolutionized the field [@dyck_determination_1966].
-Now relatively inexpensive and convenient, spectrometry allows for the rapid capture of quantitative data, and has been widely adopted as a standard in the field [@andersson_quantifying_2006].
+Now relatively inexpensive and convenient, spectrometry allows for the rapid capture of quantitative data, and has been widely adopted as standard across the field [@andersson_quantifying_2006].
 Digital photography is also increasingly being used to quantify biological colouration [@stevens_using_2007], as high-resolution cameras are inexpensive and allow for the rapid, repeatable sampling of multiple colour patches [@mckay_use_2013].
 
-Subtle variation in the application of these common methods can significantly shape the resulting data.
+Subtle variation in the application of common methods can significantly shape the resulting data.
 When using spectrometry, the angle from which spectra are collected [@santos_strong_2007], the calibration of light and dark standards, the distance between light source and surface, and the size, shape, glossiness or iridescence of the specimen will all directly affect the recorded spectra [@andersson_quantifying_2006; @kemp_ornamental_2008; @meadows_quantifying_2011; @schaefer_birds_2008].
-The main challenge for gathering colour data from photographs is ensuring standardization against nonlinear responses to light intensity, biases towards particular wavebands, and inconsistent lighting [@stevens_using_2007].
+Equally, the main challenge for gathering colour data from photographs is ensuring standardization against nonlinear responses to light intensity, biases towards particular wavebands, and inconsistent lighting [@stevens_using_2007].
 Also, most digital cameras used for the measurement of biological colouration are consumer products designed for the human visual system.
@@
 
-Given the myriad degrees of freedom that exist when measuring and analysing colour, and the significant consequences of any variation, the conditions under which data were collected must be reported exactly....still working on this...
-Methods reporting should ideally have well defined standards such that sufficient detail is provided to allow comprehensive peer review, precise replication, cross-study comparison and data sharing [@alsheikh_public_2011; @sandve_ten_2013]....
+The comprehensive reporting of methods a simple and crucial step in ensuring research is reproducible.
+Accordingly, we developed a list of essential information about the capture (Table 1) and analysis (Table 2) of colour data that should ideally be reported.
+Analytical techniques are diverse, mathematically complex, and are being developed rapidly [@kelber_spectral_2010].
+Speedy progress means that a deep understanding of common methods can quickly outstrip working knowledge for the average researcher.
+As a consequence, the subtle complexity of many analytical techniques is often overlooked by empiricists, which leads to critical methodological information often not being reported.
+Our guidelines for reporting the details of colour analyses (Table 2) thus cover two broad, common categories of analysis: colourimetric (or 'spectral') analyses, and visual modelling.
+These tables present the framework for our subsequent analysis of methods reporting in the literature.
 
-## Assessing reproducibility
+## Assessing reporting and reproducibility
 
 To assess the current state of reproducibility in the field we searched papers from 2013 in 22 leading journals: _American Journal of Botany_, _The American Naturalist_, _Animal Behaviour_, _Behavioral Ecology_, _Behavioral Ecology and Sociobiology_, _Biological Journal of the Linnean Society_, _Biology Letters_, _Current Zoology_, _Ecology_, _Ecology and Evolution_, _Ecology Letters_, _Ethology_, _Evolution_, _Functional Ecology_, _Journal of Ecology_, _The Journal of Evolutionary Biology_, _The Journal of Experimental Biology_, _Naturwissenschaften_, _New Phytologist_, _Oikos_, _PLoS ONE_, _Proceedings of the Royal Society B: Biological Sciences_.
-On each of the journals' homepages, we used the Boolean terms 'colour\*', 'color\*' or 'spectra\*' to search the title or abstract.
-We included only those papers that used either a spectrometer or camera to quantify colouration.
-We excluded review papers, methodological papers, papers quantifying colour patterns rather than the chromatic properties of a colour patch, and papers taking micro-spectrometric measurements of retinal absorbance.
-To review the large number of papers returned from our search, journals were haphazardly divided up between the authors.
-In order to reduce the risk of observer bias in our assessment, each paper was read and reassessed by three authors.
+On each of the journals' homepages, we used the Boolean phrase 'colour\*' or 'color\*' or 'spectra\*' to search the title or abstract.
+To review the 216 papers returned from our search, journals were haphazardly divided up between the authors.
+We excluded review papers, methodological papers, papers quantifying colour patterns rather than the chromatic properties of a colour patch, and studies taking micro-spectrometric measurements of retinal absorbance.
+The final set of `r nrow(coldat)` papers included only those that used either a spectrometer or camera to quantify colouration.
+In order to reduce the risk of observer bias in our assessment, each paper included in the final set was read and reassessed by three authors.
 Any discrepancies between assessor scores were discussed and resolved prior to analysis.
-We then tabulated what details of the methods were reported against our criteria.
-The data along with analysis script can be found on github (WHAT'S THE DOI?).
-We have kept the papers used in our dataset anonymous as our aim is to explore the general question of repeatability in the field.
+We then tabulated what details of the methods were reported against our suggested criteria (Tables 1 and 2), and recorded whether data (in either a 'raw' or 'processed' form) and/or any code were publicly available.
+The data along with analysis script can be found on github [![DOI](https://zenodo.org/badge/doi/10.5281/zenodo.13245.svg)](http://dx.doi.org/10.5281/zenodo.13245).
+We have kept the papers used in our dataset anonymous as our aim is to explore the general question of repeatability in the field and not single our any particular study.
 Our search returned a total of 216 papers, however, only `r nrow(coldat)` met our criteria and were included in the final analysis.
@@
 
-_Guidelines_
-
-The comprehensive reporting of methods is the simplest step in ensuring research is reproducible.
-Accordingly, we developed a list of essential information about the capture (Table 1) and analysis (Table 2) of colour data that should ideally be reported.
-Given their overwhelming popularity, our data collection reporting guidelines are focused on spectrometry and photography for data collection (Table 1), and do not consider qualitative methods such as the use of Munsell chips or paint swatches.
-Analytical techniques are diverse and are being developed rapidly [@kelber_spectral_2010], and this rate of progress means that a deep understanding of common methods is increasingly beyond the grasp of the average researcher.
-As a consequence, the subtle complexity of many analytical techniques is often overlooked by empiricists, which leads to critical methodological information often not being reported.
-Our guidelines for reporting the details of colour analyses (Table 2) thus cover two broad, common categories of analysis: colourimetric (or 'spectral') analyses, and visual modelling.
-These tables present the framework for our subsequent analysis of methods reporting in the literature.
-
-_The current state of methods reporting in the literature_
+_Reproducibility of data colection_
 
@@
 Reporting of the make and model of the spectrometer and camera was high (Figure 1, spectrometers = `r round(specModel, digits = 2)`% and cameras = `r round(camModel, digits = 2)`%), however, only `r round(propCam_fin["prop_light_source"], digits = 2)`%  and `r round(propSpec_fin["prop_light_source"], digits = 2)`% of camera and spectrometer studies reported a light source.
-In addition, the number of reflectance spectra and pixels averaged were only reported in `r round(propSpec_fin["prop_specpix_avg"], digits = 2)`% and `r round(propCam_fin["prop_specpix_avg"], digits = 2)`% of studies, respectively.
+The number of reflectance spectra and pixels averaged were only reported in `r round(propSpec_fin["prop_specpix_avg"], digits = 2)`% and `r round(propCam_fin["prop_specpix_avg"], digits = 2)`% of studies, respectively.
 Surprisingly, dark standards were only reported in `r round(propSpec_fin["prop_dark_std"], digits = 2)`% of the spectrometer studies, while white standards were reported in `r round(propSpec_fin["prop_white_stdspec"], digits = 2)`%.
 Studies quantifying spectral reflectance often did not report the integration time of the spectrophotmeter (`r round(propSpec_fin["prop_int_time"], digits = 2)`%), or the angle of the optical probe relative to the surface (`r round(propSpec_fin["prop_spec_angle"], digits = 2)`%) or distance (`r round(propSpec_fin["prop_spec_dist"], digits = 2)`%) the probe was held from the measurement surface.
+
+_Reproducibility of analyses_
 
@@
 
+_How to increase transparency_
+
 While some of the percentages reported above seem alarming, it is important to note that around one-third of papers (`r round(ref_stud, digits = 2)`%) made reference to previous work for various methods.
 This means that our above figures are slightly liberal given that the referenced works may have comprehensively covered some of these criteria.
-Despite this, two-thirds of the studies did not reference previous work, and were missing important methodological details, suggesting the need for more comprehensive reporting of methods...touch more discussion here, instead of in conclusions(?)...e.g...When possible, we also recommend that authors also justify data capture and analytical approaches - something few studies we investigated explicitly outlined.
+Despite this, two-thirds of the studies did not reference previous work, and were missing important methodological details.
+The probe distance and angle should be standard reporting for all studies using spectrophometry.
+Visual models were regularly used to analyse spectral data.
+However, there were many important aspects of the analysis that were seldom reported, including the type of quantum catch used, whether photoreceptor adaptation was modelled and the type of neural noise used in receptor-noise models.
+Photoreceptor densities should also be reported where possible even if using standard avian photorecptor densities.
+These all have important consequences for downstream analysis and the reproducibility of colour signal quantification [e.g.
+@kemp_ornamental_2008; @santos_comparison_2007].
+We also recommend the presentation of spectra where possible, as these figures allow for the rapid assessment of the nature and quality of the data, independent of downstream processing.
+We found that only `r round(propSpec["refl_fig"], digits = 2)`% of studies using spectral reflectance data had such a figure.
+Although not considered in our analysis, we also recommend that authors explicitly outline the biological justification underlying their choices for data capture and analysis.
 
@@
 The paucity of data being made available post-publication in studies of biological colouration is in line with other fields [@drew_lost_2013; @vines_mandated_2013; @wolkovich_advances_2012], including the broader field of animal behaviour [@caetano_forgotten_2014].
-There is evidence, however, that this trend is shifting as funders and journals increasingly mandate the release of data [@whitlock_data_2010].
+There is evidence, however, that this trend is shifting as funding agencies and publishers increasingly mandate the release of data [@whitlock_data_2010].
 The clearest benefit of open access to data is that it allows researchers to build upon previous work for the testing and refinement of new ideas and methods.
@@
 individual reflectance spectra), to maximise its utility.
-Public data repositories such as figshare (REF?) and dryad (REF?) facilitate the permanent archival of data as citable research objects, and the preparation of data for publication is often straightforward [@whitlock_data_2011].
+Public data repositories such as figshare and dryad allow the permanent archival of data as citable research objects, and the preparation of data for publication is often straightforward [@whitlock_data_2011].
 The use of modest data embargoes and appropriate licenses can help to ensure that original authors are able to make the best use of data, and are subsequently credited [though we acknowledge that tensions over data-sharing exist; discussed in @roche_troubleshooting_2014].
 
-None of the included studies linked to any form of code, which is not surprising given the popularity of graphical-user-interface (GUI) based statistical, and colour-analytical software [e.g.
+None of the included studies linked to any form of code, which is not surprising given the popularity of graphical-user-interface (GUI) based statistical and colour-analytical software [e.g.
 AVICOL; @gomez_avicol_2006].
-Here we simply note that there are several advantages to moving from a GUI-based to a programming workflow, in spite of the initial time investment required.
-The open-source programming language R [@R_program], for example, is free and is host to a sizeable community of developers that build and implement analytical tools (via 'packages').
-Indeed, packages specifically for the analysis of colour data already exist [e.g.
-pavo; @maia_pavo_2013].
-Perhaps the most significant advantage is that analysis scripts produced in such a way represent a complete documented history of a study's analytical methods.
-For authors this means that the details of a study's methods may be precisely preserved, and it allows other researchers to directly build upon.....
-
-........greater speed and better control...documenting of methods & maintaining chain of modification for reproducibility...something about git facilitating collaboration & project management [@ram_git_2013].
-Maybe rmarkdown [@rmarkdown]
+Here we simply note that there are several advantages to moving away from a GUI-based to a programming workflow, in spite of the initial time investment required.
+Programming languages such as Python [@van_python_1995]and R [@R_program], for example, are free, open-source, and are host to communities of developers that continuously build and implement analytical tools.
+Tools specifically for the analysis of colour data already exist [e.g.
+the package 'pavo' for R; @maia_pavo_2013].
+Programming languages are by nature flexible, which allows individual researchers to rapidly explore new methods as they are published, and methods developers to implement their analyses and potentially increase the reach and impact of their work.
+A significant advantage of programmatic analyses is that analysis scripts represent a complete documented history of a study's methods.
+The chain of modification of raw data may thus be preserved for researchers revisiting their own work in the future, as well as for other scientists looking to build on the results of previous studies.
 
@@
 1).
-In particular, we found that the types of light sources, numbers of pixels or spectra averaged, the distance and angle of the spectrometers optical probe relative to the measurement surface and collector, and types of standards used (i.e.
+We found that the types of light sources, numbers of pixels or spectra averaged, the distance and angle of the spectrometers optical probe relative to the measurement surface and collector, and types of standards used (i.e.
 white and black reference samples) are under-reported (Fig.
@@
 1).
+The permanent, public storage of data also plays a pivotal role in reproducibility and transparency.
+It is increasingly difficult to justify the complete loss of data [@alsheikh_public_2011], and the benefits of publicly archiving data are considerable [@piwowar_sharing_2007; @vision_open_2010; @whitlock_data_2010].
 
-The permanent, private storage of data [or, more often, the complete loss of data; @alsheikh_public_2011] is increasingly difficult to justify, and the benefits of publicly archived data are considerable [@piwowar_sharing_2007; @vision_open_2010; @whitlock_data_2010].
-Most of the data underlying studies of biological colouration are unpublished, and are thus likely to be lost...
-
-The move towards increasingly reproducible research has many benefits.
-Our objective in this paper was to explore the issue of reproducibility in studies of biological colouration, and to offer guidelines on simple ways in which studies may be leveraged to their full potential.
-We hope our that our discussion of methods reporting and data archiving will contribute to the growing culture of reproducibility in science, and that researchers will consider these issues when deciding on the best practices in their work.
-The future of colour-trait research is bright (obligatory pun?!)...something something maybe...continuing the advance of an exciting field.
+Our objective in this paper was to highlight some important gaps that can influence the reproducibility of colour research, and to offer guidelines on simple ways in which studies may be leveraged to their full potential.
+By no means do we wish to say that studies lacking important details are not of significant value.
+We simply hope that our discussion of methods reporting and data archiving will contribute to the growing culture of reproducibility in science, and that researchers will consider these issues when deciding on the best practices in their work.
+It is only through critical appraisal and collegial discussion that we can identify areas for improvement, and constructively contribute to the growth of our field.
+Such assessments are particularly important in an age where accountability, transparency and reproducibility are at the forefront of discussions of the state of scientific research [@drew_lost_2013; @peng_reproducible_2011; @wicherts_willingness_2011].
+We believe that part of the issue at hand is simply that researchers seldom have clear guidelines to follow when undertaking their research, and may overlook important aspects of methodology and analysis.
+We hope our critical evaluation of recent literature has highlighted a need for more careful descriptions of how colour traits are measured and analyzed.
+We are optimistic that this may help to establish uniform reporting guidelines for publication and increase transparency and reproducibility, which will continue the advance of an exciting era in colour research.
 

### 2014-12-15 docs/ms.Rmd
Squashed a few bugs in the code before sending out to the gang

@@
 
-Our aim was to explore the state of reproducibility in studies of biological colouration, and to outline ways in which it may be optimised.
-We begin by briefly outlining some of the most common methods for studying biological colouration, with an emphasis on the way in which their subtle complexity can shape results.
+Our aim was to explore the state of reproducibility in studies of biological colouration, and to outline ways in which it may be improved.
+We begin by briefly outlining some of the most common methods for studying biological colouration, with an emphasis on the way in which their subtle complexity can dramatically shape results.
 We then present a set of guidelines for the comprehensive reporting of methods, and we analyse a subset of the literature against these criteria to assess the state of methods reporting in studies of biological colouration.
@@
 Generations of biologists have endeavoured to explain the mechanisms and functions of animal and plant colouration [@thayer_concealing_1909; @poulton_colours_1890; @wallace_natural_1891], and uncovering best practices in measuring colour has been a great challenge.
-The application of spectrometry to questions of biological colouration revolutionized the field [@dyck_determination_1966].
+The application of spectrometry (also called spectroradiometry, or spectrophotometry) to questions of biological colouration revolutionized the field [@dyck_determination_1966].
 Now relatively inexpensive and convenient, spectrometry allows for the rapid capture of quantitative data, and has been widely adopted as standard across the field [@andersson_quantifying_2006].
@@
 
-Subtle variation in the application of common methods can significantly shape the resulting data.
-When using spectrometry, the angle from which spectra are collected [@santos_strong_2007], the calibration of light and dark standards, the distance between light source and surface, and the size, shape, glossiness or iridescence of the specimen will all directly affect the recorded spectra [@andersson_quantifying_2006; @kemp_ornamental_2008; @meadows_quantifying_2011; @schaefer_birds_2008].
+Subtle variation in the application of these common methods can significantly alter the resulting data.
+When using spectrometry, for example, the angle from which spectra are collected [@santos_strong_2007], the calibration of light and dark standards, the distance between light source and surface, and the size, shape, glossiness or iridescence of the specimen will all directly affect the recorded spectra [@andersson_quantifying_2006; @kemp_ornamental_2008; @meadows_quantifying_2011; @schaefer_birds_2008].
 Equally, the main challenge for gathering colour data from photographs is ensuring standardization against nonlinear responses to light intensity, biases towards particular wavebands, and inconsistent lighting [@stevens_using_2007].
@@
 
-The comprehensive reporting of methods a simple and crucial step in ensuring research is reproducible.
+The comprehensive reporting of methods is a simple and crucial step in ensuring research is reproducible.
 Accordingly, we developed a list of essential information about the capture (Table 1) and analysis (Table 2) of colour data that should ideally be reported.
@@
 Speedy progress means that a deep understanding of common methods can quickly outstrip working knowledge for the average researcher.
-As a consequence, the subtle complexity of many analytical techniques is often overlooked by empiricists, which leads to critical methodological information often not being reported.
-Our guidelines for reporting the details of colour analyses (Table 2) thus cover two broad, common categories of analysis: colourimetric (or 'spectral') analyses, and visual modelling.
+As a consequence, the subtle complexity of many analytical techniques is often overlooked by empiricists, which leads to critical methodological information not being reported.
+Our guidelines for reporting the details of colour analyses (Table 2) thus cover two broad, common methods: colourimetric (or 'spectral') analyses, and visual modelling.
 These tables present the framework for our subsequent analysis of methods reporting in the literature.
@@
 We then tabulated what details of the methods were reported against our suggested criteria (Tables 1 and 2), and recorded whether data (in either a 'raw' or 'processed' form) and/or any code were publicly available.
-The data along with analysis script can be found on github [![DOI](https://zenodo.org/badge/doi/10.5281/zenodo.13245.svg)](http://dx.doi.org/10.5281/zenodo.13245).
-We have kept the papers used in our dataset anonymous as our aim is to explore the general question of repeatability in the field and not single our any particular study.
-Our search returned a total of 216 papers, however, only `r nrow(coldat)` met our criteria and were included in the final analysis.
+The data along with analysis script can be found on github (http://dx.doi.org/10.5281/zenodo.13245).
+We have kept the papers used in our dataset anonymous as our aim is to explore the general question of repeatability in the field and not single out particular studies.
 
@@
 
-_Reproducibility of data colection_
+_Reproducibility of data collection_
 
@@
 purely statistical analyses, such as principle component analysis on reflectance spectra).
-Of the studies quantifying colour using spectral reflectance curves `r analysisTypeSpec[1]` used colourimetric measures, `r analysisTypeSpec[2]` used multiple analyses, `r analysisTypeSpec[4]` used visual models and `r analysisTypeSpec[3]` used other forms of analysis.
+Of the studies quantifying colour using spectral reflectance, `r analysisTypeSpec[1]` used colourimetric measures, `r analysisTypeSpec[2]` used multiple analyses, `r analysisTypeSpec[4]` used visual models and `r analysisTypeSpec[3]` used other forms of analysis.
 Studies using colourimetric measurements defined their measure of brightness, hue and/or chroma in `r round(col_def, digits = 2)`% of the studies.
@@
 
-While some of the percentages reported above seem alarming, it is important to note that around one-third of papers (`r round(ref_stud, digits = 2)`%) made reference to previous work for various methods.
+While some of the figures reported above seem alarming, it is important to note that around one-third of papers (`r round(ref_stud, digits = 2)`%) made reference to previous work for various methods.
 This means that our above figures are slightly liberal given that the referenced works may have comprehensively covered some of these criteria.
 Despite this, two-thirds of the studies did not reference previous work, and were missing important methodological details.
-The probe distance and angle should be standard reporting for all studies using spectrophometry.
+In particular, the light source - sample - probe geometry should be standard reporting for all studies using spectrometry.
 Visual models were regularly used to analyse spectral data.
 However, there were many important aspects of the analysis that were seldom reported, including the type of quantum catch used, whether photoreceptor adaptation was modelled and the type of neural noise used in receptor-noise models.
-Photoreceptor densities should also be reported where possible even if using standard avian photorecptor densities.
-These all have important consequences for downstream analysis and the reproducibility of colour signal quantification [e.g.
+Photoreceptor densities should also be reported where possible even if using standard avian photorecptor densities, as all of these parameters have important consequences for downstream analysis and, hence, the reproducibility of colour trait quantification [e.g.
 @kemp_ornamental_2008; @santos_comparison_2007].
-We also recommend the presentation of spectra where possible, as these figures allow for the rapid assessment of the nature and quality of the data, independent of downstream processing.
-We found that only `r round(propSpec["refl_fig"], digits = 2)`% of studies using spectral reflectance data had such a figure.
+We also recommend the presentation of spectra where possible as these figures allow for the rapid assessment of the nature and quality of the data, independent of downstream processing.
+We found that only `r round(propSpec_fin["prop_refl_fig"], digits = 2)`% of studies using spectral reflectance data had such a figure.
 Although not considered in our analysis, we also recommend that authors explicitly outline the biological justification underlying their choices for data capture and analysis.
@@
 
-Of the `r nrow(coldat)` studies analysed, `r round((data["raw"]/sum(data))*100, digits = 2)`% publicly provided the raw underlying data, `round((data["proc"]/sum(data))*100, digits = 2)`% provided data in a pre-processed form, and `r round((data["none"]/sum(data))*100, digits = 2)`% of studies did not provide any publicly accessible data.
+Of the `r nrow(coldat)` studies analysed, `r round((data["raw"]/sum(data))*100, digits = 2)`% publicly provided the raw underlying data, `r round((data["proc"]/sum(data))*100, digits = 2)`% provided data in a pre-processed form, and `r round((data["none"]/sum(data))*100, digits = 2)`% of studies did not provide any publicly accessible data.
 The paucity of data being made available post-publication in studies of biological colouration is in line with other fields [@drew_lost_2013; @vines_mandated_2013; @wolkovich_advances_2012], including the broader field of animal behaviour [@caetano_forgotten_2014].
@@
 @allen_analyzing_2013; @endler_framework_2012; @stoddard_pattern_2014].
-The provision of open data may also foster collaborations as researchers draw upon existing work [@whitlock_data_2011], and increased data sharing has been shown to positively correlate with citations [@piwowar_sharing_2007].
+More generally, the provision of open data may also foster collaborations as researchers draw upon existing work [@whitlock_data_2011], and increased data sharing has been shown to positively correlate with citations [@piwowar_sharing_2007].
 Given the increasing rate of retractions, there is also impetus for researchers to maintain credibility through transparency [@steen_retractions_2010; @van_trouble_2011; @wicherts_willingness_2011].
@@
 Public data repositories such as figshare and dryad allow the permanent archival of data as citable research objects, and the preparation of data for publication is often straightforward [@whitlock_data_2011].
-The use of modest data embargoes and appropriate licenses can help to ensure that original authors are able to make the best use of data, and are subsequently credited [though we acknowledge that tensions over data-sharing exist; discussed in @roche_troubleshooting_2014].
+The use of modest data embargoes and appropriate licenses can help to ensure that original authors are able to make the best use of data, and are subsequently credited, though we acknowledge that tensions over data-sharing exist [discussed in @roche_troubleshooting_2014].
 
@@
 Here we simply note that there are several advantages to moving away from a GUI-based to a programming workflow, in spite of the initial time investment required.
-Programming languages such as Python [@van_python_1995]and R [@R_program], for example, are free, open-source, and are host to communities of developers that continuously build and implement analytical tools.
-Tools specifically for the analysis of colour data already exist [e.g.
+Programming languages such as Python [@van_python_1995] and R [@R_program], for example, are free, open-source, and are host to communities of developers that continuously build and implement analytical tools.
+Indeed, tools specifically for the analysis of colour data already exist [e.g.
 the package 'pavo' for R; @maia_pavo_2013].
-Programming languages are by nature flexible, which allows individual researchers to rapidly explore new methods as they are published, and methods developers to implement their analyses and potentially increase the reach and impact of their work.
-A significant advantage of programmatic analyses is that analysis scripts represent a complete documented history of a study's methods.
-The chain of modification of raw data may thus be preserved for researchers revisiting their own work in the future, as well as for other scientists looking to build on the results of previous studies.
+Programming languages are flexible by nature, which allows individual researchers to rapidly explore new methods as they are published, and methods developers to implement their analyses to potentially increase the reach and impact of their work.
+A significant advantage of programmatic analyses is that analysis scripts, when properly curated, represent a complete documented history of a study's methods.
+The chain of modification of raw data may thus be preserved, both for researchers revisiting their own work in the future and for other scientists looking to build directly on the results of previous studies.
 
@@
 
-Our review of the recent colour literature highlights some significant impediments to reproducibility.
+Our review of the recent literature highlights some significant impediments to reproducibility in the study of biological colouration.
 The simplest step towards reproducible research is through the comprehensive reporting of methods, yet crucial methodological information is often unreported (Fig.
@@
 1).
-The permanent, public storage of data also plays a pivotal role in reproducibility and transparency.
+The permanent, public storage of data also plays a pivotal role in reproducibility and transparency, yet public data storage in studies of colour traits is not yet the norm.
 It is increasingly difficult to justify the complete loss of data [@alsheikh_public_2011], and the benefits of publicly archiving data are considerable [@piwowar_sharing_2007; @vision_open_2010; @whitlock_data_2010].

### 2014-12-16 docs/ms.Rmd
Changed rounding to 1 decimal place and fixed up a bit of analysis

@@
 Of these studies, `r both` quantified colour using both a spectrometer and a camera.
-Reporting of the make and model of the spectrometer and camera was high (Figure 1, spectrometers = `r round(specModel, digits = 2)`% and cameras = `r round(camModel, digits = 2)`%), however, only `r round(propCam_fin["prop_light_source"], digits = 2)`%  and `r round(propSpec_fin["prop_light_source"], digits = 2)`% of camera and spectrometer studies reported a light source.
-The number of reflectance spectra and pixels averaged were only reported in `r round(propSpec_fin["prop_specpix_avg"], digits = 2)`% and `r round(propCam_fin["prop_specpix_avg"], digits = 2)`% of studies, respectively.
-Surprisingly, dark standards were only reported in `r round(propSpec_fin["prop_dark_std"], digits = 2)`% of the spectrometer studies, while white standards were reported in `r round(propSpec_fin["prop_white_stdspec"], digits = 2)`%.
-Studies quantifying spectral reflectance often did not report the integration time of the spectrophotmeter (`r round(propSpec_fin["prop_int_time"], digits = 2)`%), or the angle of the optical probe relative to the surface (`r round(propSpec_fin["prop_spec_angle"], digits = 2)`%) or distance (`r round(propSpec_fin["prop_spec_dist"], digits = 2)`%) the probe was held from the measurement surface.
+Reporting of the make and model of the spectrometer and camera was high (Figure 1, spectrometers = `r round(specModel, digits = 1)`% and cameras = `r round(camModel, digits = 1)`%), however, only `r round(propCam_fin["prop_light_source"], digits = 1)`%  and `r round(propSpec_fin["prop_light_source"], digits = 1)`% of camera and spectrometer studies reported a light source.
+The number of reflectance spectra and pixels averaged were only reported in `r round(propSpec_fin["prop_specpix_avg"], digits = 1)`% and `r round(propCam_fin["prop_specpix_avg"], digits = 1)`% of studies, respectively.
+Surprisingly, dark standards were only reported in `r round(propSpec_fin["prop_dark_std"], digits = 1)`% of the spectrometer studies, while white standards were reported in `r round(propSpec_fin["prop_white_stdspec"], digits = 1)`%.
+Studies quantifying spectral reflectance often did not report the integration time of the spectrophotmeter (`r round(propSpec_fin["prop_int_time"], digits = 1)`%), or the angle of the optical probe relative to the surface (`r round(propSpec_fin["prop_spec_angle"], digits = 1)`%) or distance (`r round(propSpec_fin["prop_spec_dist"], digits = 1)`%) the probe was held from the measurement surface.
 
@@
 Of the studies quantifying colour using spectral reflectance, `r analysisTypeSpec[1]` used colourimetric measures, `r analysisTypeSpec[2]` used multiple analyses, `r analysisTypeSpec[4]` used visual models and `r analysisTypeSpec[3]` used other forms of analysis.
-Studies using colourimetric measurements defined their measure of brightness, hue and/or chroma in `r round(col_def, digits = 2)`% of the studies.
+Studies using colourimetric measurements defined their measure of brightness, hue and/or chroma in `r round(col_def, digits = 1)`% of the studies.
 
@@
 Case specific models were used in `r vismod_type[2]` papers.
-Of visual modelling papers, `r round(propSpec_fin["prop_irrad_type"], digits =2)`% reported the type of irradiance when necessary, `r round(propSpec_fin["prop_vis_mod_sp"], digits = 2)`% the species modelled as a receiver.
-The background used in visual models was reported `r round(propSpec_fin["prop_vis_mod_bkg"], digits = 2)`% of the time.
-The type of receptor noise modelled was reported in only `r round(propSpec_fin["prop_vis_mod_noise_type"], digits = 2)`% of the studies.
-Explicit mention of the type of quantum catch (`r round(propSpec_fin["prop_vis_mod_qcatch"], digits = 2)`% of studies), or whether photoreceptor adaptation (`r round(propSpec_fin["prop_vis_mod_adapt"], digits = 2)`%) was modelled, was not common.
+Of visual modelling papers, `r round(propSpec_fin["prop_irrad_type"], digits =2)`% reported the type of irradiance when necessary, `r round(propSpec_fin["prop_vis_mod_sp"], digits = 1)`% the species modelled as a receiver.
+The background used in visual models was reported `r round(propSpec_fin["prop_vis_mod_bkg"], digits = 1)`% of the time.
+The type of receptor noise modelled was reported in only `r round(propSpec_fin["prop_vis_mod_noise_type"], digits = 1)`% of the studies.
+Explicit mention of the type of quantum catch (`r round(propSpec_fin["prop_vis_mod_qcatch"], digits = 1)`% of studies), or whether photoreceptor adaptation (`r round(propSpec_fin["prop_vis_mod_adapt"], digits = 1)`%) was modelled, was not common.
 
@@
 
-While some of the figures reported above seem alarming, it is important to note that around one-third of papers (`r round(ref_stud, digits = 2)`%) made reference to previous work for various methods.
+While some of the figures reported above seem alarming, it is important to note that around one-third of papers (`r round(ref_stud, digits = 1)`%) made reference to previous work for various methods.
 This means that our above figures are slightly liberal given that the referenced works may have comprehensively covered some of these criteria.
@@
 We also recommend the presentation of spectra where possible as these figures allow for the rapid assessment of the nature and quality of the data, independent of downstream processing.
-We found that only `r round(propSpec_fin["prop_refl_fig"], digits = 2)`% of studies using spectral reflectance data had such a figure.
+We found that only `r round(propSpec_fin["prop_refl_fig"], digits = 1)`% of studies using spectral reflectance data had such a figure.
 Although not considered in our analysis, we also recommend that authors explicitly outline the biological justification underlying their choices for data capture and analysis.
@@
 
-Of the `r nrow(coldat)` studies analysed, `r round((data["raw"]/sum(data))*100, digits = 2)`% publicly provided the raw underlying data, `r round((data["proc"]/sum(data))*100, digits = 2)`% provided data in a pre-processed form, and `r round((data["none"]/sum(data))*100, digits = 2)`% of studies did not provide any publicly accessible data.
+Of the `r nrow(coldat)` studies analysed, `r round((data["raw"]/sum(data))*100, digits = 1)`% publicly provided the raw underlying data, `r round((data["proc"]/sum(data))*100, digits = 1)`% provided data in a pre-processed form, and `r round((data["none"]/sum(data))*100, digits = 1)`% of studies did not provide any publicly accessible data.
 The paucity of data being made available post-publication in studies of biological colouration is in line with other fields [@drew_lost_2013; @vines_mandated_2013; @wolkovich_advances_2012], including the broader field of animal behaviour [@caetano_forgotten_2014].

### 2014-12-16 R/analysis.r

@@ -17,13 +17,6 @@ sapply(packages, function(x) library(x, character.only = TRUE, verbose = FALSE))
 coldat <- read.csv("data/col_data.csv", header = TRUE)
 str(coldat)
 
-## Process data. Remove irrelevant variables for analaysis. 
-## Remove justification columns, which are just simply for our use and complicate analysis
-col_just   <- colnames(coldat)[c(grep("_just", colnames(coldat)))]
-coldat     <- coldat[,! colnames(coldat) %in% col_just][-c(2:13, 35:36)] # Additioally remove details on analysis and details on papers, but keep paper number.
-
-## Things we need to check. The studies categorized as "multiple" Need to make clear what the values (0 or 1) correspond to camera or specs. 
-
 #--------------------------------------------- ANALYSIS -----------------------------------------#
 ## How many studies utilized specs (1), cameras (2), both (3)?
 studytype <- table(coldat$stud_type)
@@ -40,7 +33,7 @@ camModel  <- (sum(coldat$cam_model, na.rm = TRUE)/(cam+both))*100
 
 ## Of the studies using a spec, what proportion report the light source. Note that column needs to be extracted because it combines cameras and specs
 
-DatSpec         <- subset(coldat, stud_type == 1 |stud_type == 3)
+DatSpec   <- subset(coldat, stud_type == 1 |stud_type == 3)
 
 #Compute proportions of the various criteria. Extract only the rows we are interested in using
 criteriaSpec <- c("light_source","specpix_avg","dark_std", "white_stdspec", "int_time", "spec_angle", "spec_dist", "irrad_type", "vis_mod", "vis_mod_sp", "vis_mod_param", "vis_mod_adapt", "vis_mod_qcatch", "vis_mod_bkg", "vis_mod_noise_type", "refl_fig")

### 2014-12-17 docs/ms.Rmd
Integrating latest round of changes to ms from everyone - almost ready

@@
 
-Word count:
+Word count: 2613
 
-Number of figures:
+Number of figures: 1
 
@@
 
-Data archiving:
+Data archiving: Github repo (http://dx.doi.org/10.5281/zenodo.13245)
 
@@
 The study of colour in nature has generated insights into fundamental evolutionary and ecological processes, and research into colour traits is a rapidly growing field [@kelber_spectral_2010].
-Resurgent interest in biological colouration has in large part been driven by the increased availability of key technologies, including spectrometry and photography, and concurrent advances in methods for analysing colour data, such as visual models [@endler_comparing_2005; @stevens_using_2007; @kelber_animal_2003].
-While these developments are positive for the field, the increasingly complex analyses being run on ever greater amounts of data heighten the need for comprehensive methods reporting and proper data management [@alsheikh_public_2011; @nekrutenko_next_2012].
+The ongoing interest in biological colouration has in some part been driven by the increased availability of key technologies, including spectrometry and photography, and concurrent advances in methods for analysing colour data, such as visual models [@endler_comparing_2005; @stevens_using_2007; @kelber_animal_2003].
+While these developments are positive for the field, the increasingly complex analyses being run on ever greater amounts of data heighten the need for comprehensive methods reporting and diligent data management [@alsheikh_public_2011; @nekrutenko_next_2012].
 
@@
 Our aim was to explore the state of reproducibility in studies of biological colouration, and to outline ways in which it may be improved.
-We begin by briefly outlining some of the most common methods for studying biological colouration, with an emphasis on the way in which their subtle complexity can dramatically shape results.
-We then present a set of guidelines for the comprehensive reporting of methods, and we analyse a subset of the literature against these criteria to assess the state of methods reporting in studies of biological colouration.
-We also assess the availability of publicly archived data and code, and suggest some useful tools for increasing the reproducibility of colour trait research more broadly.
+We begin by outlining common methods for studying biological colouration, with an emphasis on the way in which their subtle complexity can dramatically shape results.
+We then present a set of guidelines for the comprehensive reporting of methods, and analyse a subset of the literature against these criteria to assess the state of methods reporting.
+We also briefly assess the availability of publicly archived data and code, and suggest some useful tools for increasing the reproducibility of colour trait research more broadly.
 
@@
 Now relatively inexpensive and convenient, spectrometry allows for the rapid capture of quantitative data, and has been widely adopted as standard across the field [@andersson_quantifying_2006].
-Digital photography is also increasingly being used to quantify biological colouration [@stevens_using_2007], as high-resolution cameras are inexpensive and allow for the rapid, repeatable sampling of multiple colour patches [@mckay_use_2013].
+Digital photography is also increasingly being used to quantify biological colouration [@stevens_using_2007], as high-resolution cameras are inexpensive and allow for the simultaneous, repeatable sampling of multiple colour patches [@mckay_use_2013].
 
@@
 Analytical techniques are diverse, mathematically complex, and are being developed rapidly [@kelber_spectral_2010].
-Speedy progress means that a deep understanding of common methods can quickly outstrip working knowledge for the average researcher.
-As a consequence, the subtle complexity of many analytical techniques is often overlooked by empiricists, which leads to critical methodological information not being reported.
+Rapid progress means that the need for a deep understanding of common methods can quickly outstrip the working knowledge of the average researcher.
+As a consequence, the subtle complexity of many analytical techniques can be overlooked by empiricists, leading to critical methodological information not being reported.
 Our guidelines for reporting the details of colour analyses (Table 2) thus cover two broad, common methods: colourimetric (or 'spectral') analyses, and visual modelling.
@@
 To review the 216 papers returned from our search, journals were haphazardly divided up between the authors.
-We excluded review papers, methodological papers, papers quantifying colour patterns rather than the chromatic properties of a colour patch, and studies taking micro-spectrometric measurements of retinal absorbance.
+We excluded review papers, methodological papers, papers quantifying spatial rather than chromatic properties of a colour patch, and studies taking micro-spectrometric measurements of retinal absorbance.
 The final set of `r nrow(coldat)` papers included only those that used either a spectrometer or camera to quantify colouration.
@@
 We then tabulated what details of the methods were reported against our suggested criteria (Tables 1 and 2), and recorded whether data (in either a 'raw' or 'processed' form) and/or any code were publicly available.
-The data along with analysis script can be found on github (http://dx.doi.org/10.5281/zenodo.13245).
+The data along with analysis script has been stored as a github repo (http://dx.doi.org/10.5281/zenodo.13245).
 We have kept the papers used in our dataset anonymous as our aim is to explore the general question of repeatability in the field and not single out particular studies.
@@
 Surprisingly, dark standards were only reported in `r round(propSpec_fin["prop_dark_std"], digits = 1)`% of the spectrometer studies, while white standards were reported in `r round(propSpec_fin["prop_white_stdspec"], digits = 1)`%.
-Studies quantifying spectral reflectance often did not report the integration time of the spectrophotmeter (`r round(propSpec_fin["prop_int_time"], digits = 1)`%), or the angle of the optical probe relative to the surface (`r round(propSpec_fin["prop_spec_angle"], digits = 1)`%) or distance (`r round(propSpec_fin["prop_spec_dist"], digits = 1)`%) the probe was held from the measurement surface.
+Studies quantifying spectral reflectance often did not report the integration time of the spectrometer (`r round(propSpec_fin["prop_int_time"], digits = 1)`%), the angle of the optical probe relative to the surface (`r round(propSpec_fin["prop_spec_angle"], digits = 1)`%) or the distance (`r round(propSpec_fin["prop_spec_dist"], digits = 1)`%) the probe was held from the measurement surface.
 
@@
 The types of analysis varied across studies.
-Overall `r analysisType[1]` papers generated colourimetric variables [i.e.
+Overall, `r analysisType[1]` papers generated colourimetric variables [i.e.
 'spectral' measures of hue, saturation, and/or brightness; @montgomerie_analyzing_2006] whereas `r analysisType[4]` used visual models, `r analysisType[2]` used both colourimetrics and visual models, and `r analysisType[3]` used other forms of analysis (e.g.
 purely statistical analyses, such as principle component analysis on reflectance spectra).
-Of the studies quantifying colour using spectral reflectance, `r analysisTypeSpec[1]` used colourimetric measures, `r analysisTypeSpec[2]` used multiple analyses, `r analysisTypeSpec[4]` used visual models and `r analysisTypeSpec[3]` used other forms of analysis.
-Studies using colourimetric measurements defined their measure of brightness, hue and/or chroma in `r round(col_def, digits = 1)`% of the studies.
+Of the studies quantifying colour using spectral reflectance, `r analysisTypeSpec[1]` used colourimetric measures, `r analysisTypeSpec[2]` used multiple analyses, `r analysisTypeSpec[4]` used visual models and `r analysisTypeSpec[3]` used other forms of analysis (e.g.
+.
+Studies using colourimetric measurements specifically defined their measure of brightness, hue and/or chroma in `r round(col_def, digits = 1)`% of the studies.
 
@@
 
-_How to increase transparency_
-
 While some of the figures reported above seem alarming, it is important to note that around one-third of papers (`r round(ref_stud, digits = 1)`%) made reference to previous work for various methods.
@@
 Despite this, two-thirds of the studies did not reference previous work, and were missing important methodological details.
-In particular, the light source - sample - probe geometry should be standard reporting for all studies using spectrometry.
-Visual models were regularly used to analyse spectral data.
-However, there were many important aspects of the analysis that were seldom reported, including the type of quantum catch used, whether photoreceptor adaptation was modelled and the type of neural noise used in receptor-noise models.
-Photoreceptor densities should also be reported where possible even if using standard avian photorecptor densities, as all of these parameters have important consequences for downstream analysis and, hence, the reproducibility of colour trait quantification [e.g.
-@kemp_ornamental_2008; @santos_comparison_2007].
-We also recommend the presentation of spectra where possible as these figures allow for the rapid assessment of the nature and quality of the data, independent of downstream processing.
-We found that only `r round(propSpec_fin["prop_refl_fig"], digits = 1)`% of studies using spectral reflectance data had such a figure.
-Although not considered in our analysis, we also recommend that authors explicitly outline the biological justification underlying their choices for data capture and analysis.
 
-## The public availability of data and code
+## Public availability of data and code
 
@@
 1).
-We found that the types of light sources, numbers of pixels or spectra averaged, the distance and angle of the spectrometers optical probe relative to the measurement surface and collector, and types of standards used (i.e.
-white and black reference samples) are under-reported (Fig.
+In particular, the light source - sample - probe geometry should be standard reporting for all studies using spectrometry.
+Visual models were regularly used to analyse spectral data.
+However, there were many important aspects of the analysis that were seldom reported, including the type of quantum catch used, whether photoreceptor adaptation was modelled and the type of neural noise used in receptor-noise models (Fig.
 1).
-Visual models were regularly used to analyse spectral data, yet key details such as the form of quantum catch used, whether photoreceptor adaptation was modelled, and the type of noise used in receptor-noise models were often not reported (Fig.
-1).
+Photoreceptor densities should also be reported where possible even if using standard avian densities, as all of these parameters have important consequences for downstream analysis and, hence, the reproducibility of colour trait quantification [e.g.
+@kemp_ornamental_2008; @santos_comparison_2007].
+We also recommend the presentation of spectra where possible as these figures allow for the rapid assessment of the nature and quality of the data, independent of downstream processing.
+We found that only `r round(propSpec_fin["prop_refl_fig"], digits = 1)`% of studies using spectral reflectance data had such a figure.
+Although not considered in our analysis, we also recommend that authors explicitly outline the biological justification underlying their choices for data capture and analysis.
 The permanent, public storage of data also plays a pivotal role in reproducibility and transparency, yet public data storage in studies of colour traits is not yet the norm.
-It is increasingly difficult to justify the complete loss of data [@alsheikh_public_2011], and the benefits of publicly archiving data are considerable [@piwowar_sharing_2007; @vision_open_2010; @whitlock_data_2010].
+We encourage researchers to consider the substantial benefits of publicly archiving their data [@piwowar_sharing_2007; @vision_open_2010; @whitlock_data_2010].
 
 Our objective in this paper was to highlight some important gaps that can influence the reproducibility of colour research, and to offer guidelines on simple ways in which studies may be leveraged to their full potential.
-By no means do we wish to say that studies lacking important details are not of significant value.
-We simply hope that our discussion of methods reporting and data archiving will contribute to the growing culture of reproducibility in science, and that researchers will consider these issues when deciding on the best practices in their work.
+We do not wish to deprecate studies that happen to lack some methodological details.
+However, we hope that our discussion of methods reporting and data archiving will contribute to the growing culture of reproducibility in science, and that researchers will consider these issues when deciding on the best practices in their work.
 It is only through critical appraisal and collegial discussion that we can identify areas for improvement, and constructively contribute to the growth of our field.
 Such assessments are particularly important in an age where accountability, transparency and reproducibility are at the forefront of discussions of the state of scientific research [@drew_lost_2013; @peng_reproducible_2011; @wicherts_willingness_2011].
-We believe that part of the issue at hand is simply that researchers seldom have clear guidelines to follow when undertaking their research, and may overlook important aspects of methodology and analysis.
-We hope our critical evaluation of recent literature has highlighted a need for more careful descriptions of how colour traits are measured and analyzed.
+We believe that part of the issue at hand is simply that researchers seldom have clear guidelines to follow when undertaking their research, and may overlook important aspects of methods and analysis reporting.
+We hope our critical evaluation of recent literature has highlighted a need for more careful descriptions of how colour is measured and analysed.
 We are optimistic that this may help to establish uniform reporting guidelines for publication and increase transparency and reproducibility, which will continue the advance of an exciting era in colour research.

### 2014-12-19 docs/ms.Rmd
Quick tidy of ms, add in last of the recent changes

@@
 The study of colour in nature has generated insights into fundamental evolutionary and ecological processes, and research into colour traits is a rapidly growing field [@kelber_spectral_2010].
-The ongoing interest in biological colouration has in some part been driven by the increased availability of key technologies, including spectrometry and photography, and concurrent advances in methods for analysing colour data, such as visual models [@endler_comparing_2005; @stevens_using_2007; @kelber_animal_2003].
+The ongoing interest in biological colouration has in some part been driven by the increased availability of key technologies, including spectrometry and photography, and concurrent advances in methods for analysing colour data, such as visual models [e.g.
+@endler_comparing_2005; @stevens_using_2007; @kelber_animal_2003].
 While these developments are positive for the field, the increasingly complex analyses being run on ever greater amounts of data heighten the need for comprehensive methods reporting and diligent data management [@alsheikh_public_2011; @nekrutenko_next_2012].
@@
 
-Our aim was to explore the state of reproducibility in studies of biological colouration, and to outline ways in which it may be improved.
+Our aim was to explore the state of reproducibility in the study of biological colouration, and to outline simple ways in which it may be improved.
 We begin by outlining common methods for studying biological colouration, with an emphasis on the way in which their subtle complexity can dramatically shape results.
-We then present a set of guidelines for the comprehensive reporting of methods, and analyse a subset of the literature against these criteria to assess the state of methods reporting.
-We also briefly assess the availability of publicly archived data and code, and suggest some useful tools for increasing the reproducibility of colour trait research more broadly.
+We then present guidelines for the comprehensive reporting of methods, and analyse a subset of the literature against these criteria to assess the state of methods reporting.
+We also quantify the availability of publicly archived data and code, and suggest some useful tools for increasing the reproducibility of colour trait research more broadly.
 
@@
 Subtle variation in the application of these common methods can significantly alter the resulting data.
-When using spectrometry, for example, the angle from which spectra are collected [@santos_strong_2007], the calibration of light and dark standards, the distance between light source and surface, and the size, shape, glossiness or iridescence of the specimen will all directly affect the recorded spectra [@andersson_quantifying_2006; @kemp_ornamental_2008; @meadows_quantifying_2011; @schaefer_birds_2008].
+When using spectrometry, for example, the angle from which spectra are collected [@santos_strong_2007], the calibration of light and dark standards, the distance between light source and surface, and the size, shape, glossiness or iridescence of the specimen will all directly affect the recorded spectra [@andersson_quantifying_2006; @dalrymple_roses_2014; @meadows_quantifying_2011; @schaefer_birds_2008].
 Equally, the main challenge for gathering colour data from photographs is ensuring standardization against nonlinear responses to light intensity, biases towards particular wavebands, and inconsistent lighting [@stevens_using_2007].
@@
 The data along with analysis script has been stored as a github repo (http://dx.doi.org/10.5281/zenodo.13245).
-We have kept the papers used in our dataset anonymous as our aim is to explore the general question of repeatability in the field and not single out particular studies.
+We have kept the papers used in our dataset anonymous as our aim was to explore the general question of reproducibility in the field and not single out particular studies.
 
@@
 Of these studies, `r both` quantified colour using both a spectrometer and a camera.
-Reporting of the make and model of the spectrometer and camera was high (Figure 1, spectrometers = `r round(specModel, digits = 1)`% and cameras = `r round(camModel, digits = 1)`%), however, only `r round(propCam_fin["prop_light_source"], digits = 1)`%  and `r round(propSpec_fin["prop_light_source"], digits = 1)`% of camera and spectrometer studies reported a light source.
+Reporting of the make and model of the spectrometer and camera was high (Figure 1; spectrometers = `r round(specModel, digits = 1)`% and cameras = `r round(camModel, digits = 1)`%), however, only `r round(propCam_fin["prop_light_source"], digits = 1)`%  and `r round(propSpec_fin["prop_light_source"], digits = 1)`% of camera and spectrometer studies reported a light source.
 The number of reflectance spectra and pixels averaged were only reported in `r round(propSpec_fin["prop_specpix_avg"], digits = 1)`% and `r round(propCam_fin["prop_specpix_avg"], digits = 1)`% of studies, respectively.
@@
 purely statistical analyses, such as principle component analysis on reflectance spectra).
-Of the studies quantifying colour using spectral reflectance, `r analysisTypeSpec[1]` used colourimetric measures, `r analysisTypeSpec[2]` used multiple analyses, `r analysisTypeSpec[4]` used visual models and `r analysisTypeSpec[3]` used other forms of analysis (e.g.
-.
-Studies using colourimetric measurements specifically defined their measure of brightness, hue and/or chroma in `r round(col_def, digits = 1)`% of the studies.
+Of the studies quantifying colour using spectral reflectance, `r analysisTypeSpec[1]` used colourimetric measures, `r analysisTypeSpec[2]` used multiple analyses, `r analysisTypeSpec[4]` used visual models alone, and `r analysisTypeSpec[3]` used other analyses.
+Studies using colourimetric measures specifically defined their measure of brightness, hue and/or chroma in `r round(col_def, digits = 1)`% of the studies.
 
 Most studies using visual models analysed their data using receptor-noise models (`r vismod_type[3]` of `r sum(vismod_type)` studies), while others used the colour hexagon (`r vismod_type[1]` studies), or a tetrahedral colour space (`r vismod_type[4]` studies).
-Case specific models were used in `r vismod_type[2]` papers.
-Of visual modelling papers, `r round(propSpec_fin["prop_irrad_type"], digits =2)`% reported the type of irradiance when necessary, `r round(propSpec_fin["prop_vis_mod_sp"], digits = 1)`% the species modelled as a receiver.
+Unique, case-specific models were used in `r vismod_type[2]` papers.
+Of visual modelling papers, `r round(propSpec_fin["prop_irrad_type"], digits =2)`% reported the type of irradiance when necessary, and `r round(propSpec_fin["prop_vis_mod_sp"], digits = 1)`% mentioned the species being modelled as receiver.
 The background used in visual models was reported `r round(propSpec_fin["prop_vis_mod_bkg"], digits = 1)`% of the time.
 The type of receptor noise modelled was reported in only `r round(propSpec_fin["prop_vis_mod_noise_type"], digits = 1)`% of the studies.
-Explicit mention of the type of quantum catch (`r round(propSpec_fin["prop_vis_mod_qcatch"], digits = 1)`% of studies), or whether photoreceptor adaptation (`r round(propSpec_fin["prop_vis_mod_adapt"], digits = 1)`%) was modelled, was not common.
+Explicit mention of the type of quantum catch used (`r round(propSpec_fin["prop_vis_mod_qcatch"], digits = 1)`% of studies), or whether photoreceptor adaptation (`r round(propSpec_fin["prop_vis_mod_adapt"], digits = 1)`%) was modelled, was not common.
 
-While some of the figures reported above seem alarming, it is important to note that around one-third of papers (`r round(ref_stud, digits = 1)`%) made reference to previous work for various methods.
+While some of the figures reported above seem alarming, it is important to note that around one-third of papers (`r round(ref_stud, digits = 1)`%) made reference to previous work for details on various methods.
 This means that our above figures are slightly liberal given that the referenced works may have comprehensively covered some of these criteria.
@@
 The clearest benefit of open access to data is that it allows researchers to build upon previous work for the testing and refinement of new ideas and methods.
-This is particularly relevant in the study of colour traits, where new methods for analysing colour data are regularly developed [e.g.
+This is particularly relevant in the study of colour traits, where new methods for analysing colour data are frequently developed [e.g.
 @allen_analyzing_2013; @endler_framework_2012; @stoddard_pattern_2014].
-More generally, the provision of open data may also foster collaborations as researchers draw upon existing work [@whitlock_data_2011], and increased data sharing has been shown to positively correlate with citations [@piwowar_sharing_2007].
-Given the increasing rate of retractions, there is also impetus for researchers to maintain credibility through transparency [@steen_retractions_2010; @van_trouble_2011; @wicherts_willingness_2011].
+More generally, the provision of open data may also foster collaborations as researchers draw upon existing work [@whitlock_data_2011], and considerably reduce the difficulty of meta-analyses and large scale comparative work [e.g.
+@burd_distinctive_2014; @maia_key_2013].
+Increased data sharing has been shown to positively correlate with citations [@piwowar_sharing_2007], and given the increasing rate of retractions there is also impetus for researchers to maintain credibility through transparency [@steen_retractions_2010; @van_trouble_2011; @wicherts_willingness_2011].
 We thus encourage researchers to publicly archive data in as raw a form as possible (e.g.
@@
 Public data repositories such as figshare and dryad allow the permanent archival of data as citable research objects, and the preparation of data for publication is often straightforward [@whitlock_data_2011].
-The use of modest data embargoes and appropriate licenses can help to ensure that original authors are able to make the best use of data, and are subsequently credited, though we acknowledge that tensions over data-sharing exist [discussed in @roche_troubleshooting_2014].
+The use of modest data embargoes and appropriate licenses can help to ensure that original authors are able to make the best use of data and are subsequently credited, though we acknowledge that tensions over data-sharing exist [discussed in @roche_troubleshooting_2014].
 
@@
 the package 'pavo' for R; @maia_pavo_2013].
-Programming languages are flexible by nature, which allows individual researchers to rapidly explore new methods as they are published, and methods developers to implement their analyses to potentially increase the reach and impact of their work.
+Programming languages are flexible by nature, which allows individual researchers to rapidly explore new methods as they are published, and methods developers to implement their analyses to increase the reach and impact of their work.
 A significant advantage of programmatic analyses is that analysis scripts, when properly curated, represent a complete documented history of a study's methods.
@@
 1).
-In particular, the light source - sample - probe geometry should be standard reporting for all studies using spectrometry.
+In particular, the light source - sample - probe geometry was rarely reported, yet should be standard reporting for all studies using spectrometry.
 Visual models were regularly used to analyse spectral data.
-However, there were many important aspects of the analysis that were seldom reported, including the type of quantum catch used, whether photoreceptor adaptation was modelled and the type of neural noise used in receptor-noise models (Fig.
+However, there were many important aspects of the analysis that were seldom reported, including the type of quantum catch used, whether photoreceptor adaptation was modelled and the type of noise used in receptor-noise models (Fig.
 1).
@@
 @kemp_ornamental_2008; @santos_comparison_2007].
-We also recommend the presentation of spectra where possible as these figures allow for the rapid assessment of the nature and quality of the data, independent of downstream processing.
+We also recommend the presentation of spectra where possible, as these figures allow for the rapid assessment of the nature and quality of the data independent of downstream processing.
 We found that only `r round(propSpec_fin["prop_refl_fig"], digits = 1)`% of studies using spectral reflectance data had such a figure.
@@
 We do not wish to deprecate studies that happen to lack some methodological details.
-However, we hope that our discussion of methods reporting and data archiving will contribute to the growing culture of reproducibility in science, and that researchers will consider these issues when deciding on the best practices in their work.
-It is only through critical appraisal and collegial discussion that we can identify areas for improvement, and constructively contribute to the growth of our field.
+We hope that our discussion of methods reporting and data archiving will contribute to the growing culture of reproducibility in science, and that researchers will consider these issues when deciding on the best practices in their work.
+It is only through critical appraisal that we can identify areas for improvement, and constructively contribute to the growth of our field.
 Such assessments are particularly important in an age where accountability, transparency and reproducibility are at the forefront of discussions of the state of scientific research [@drew_lost_2013; @peng_reproducible_2011; @wicherts_willingness_2011].

### 2015-04-15 docs/ms.Rmd
Updated text to accepted version, excluding the results section.

@@
-# Reproducible research in the study of biological colouration
+# Reproducible research in the study of biological coloration
 
@@
 
-Data archiving: Github repo (http://dx.doi.org/10.5281/zenodo.13245)
+Data archiving: Github repo (?)
 
@@
 The study of colour in nature has generated insights into fundamental evolutionary and ecological processes, and research into colour traits is a rapidly growing field [@kelber_spectral_2010].
-The ongoing interest in biological colouration has in some part been driven by the increased availability of key technologies, including spectrometry and photography, and concurrent advances in methods for analysing colour data, such as visual models [e.g.
-@endler_comparing_2005; @stevens_using_2007; @kelber_animal_2003].
-While these developments are positive for the field, the increasingly complex analyses being run on ever greater amounts of data heighten the need for comprehensive methods reporting and diligent data management [@alsheikh_public_2011; @nekrutenko_next_2012].
+The ongoing interest in biological coloration has in part been driven by the increased availability of key technologies, including spectrometry and photography, and concurrent advances in methods for analysing colour data, such as visual models [e.g., @endler_comparing_2005; @stevens_using_2007; @kelber_animal_2003].
+While these developments are positive for the field, the increasingly complex analyses being run on ever-greater amounts of data heighten the need for comprehensive methods reporting and diligent data management [@alsheikh_public_2011; @nekrutenko_next_2012].
 
@@
 Beyond simply allowing independent verification of results, reproducible research ensures greater comparability between studies and provides a foundation for testing new ideas and methods [@whitlock_data_2011; @van_trouble_2011; @piwowar_sharing_2007].
-A study may be considered truly reproducible when it satisfies three broad criteria: (i) methods are reported completely, (ii) data are publicly available and archived, and (iii) analysis code is publicly available, such that the chain of modification of raw data is documented and preserved.
-While completely reproducible research [e.g.
-@fitzjohn_much_2014] is a laudable goal, the considerable demands it imposes on researchers means that it will often, in practice, be unattainable.
+A study may be considered truly reproducible when it satisfies three broad criteria: (i) methods are reported completely, (ii) data are publicly available and archived, and (iii) the chain of modification of raw data is documented and preserved.
+While completely reproducible research [e.g., @fitzjohn_much_2014] is a laudable goal, the considerable demands it imposes on researchers means that it will often, in practice, be unattainable.
 Nevertheless, even partial reproducibility through the relatively simple practices of complete methods reporting and public data archiving is of tremendous value.
 
-Our aim was to explore the state of reproducibility in the study of biological colouration, and to outline simple ways in which it may be improved.
-We begin by outlining common methods for studying biological colouration, with an emphasis on the way in which their subtle complexity can dramatically shape results.
-We then present guidelines for the comprehensive reporting of methods, and analyse a subset of the literature against these criteria to assess the state of methods reporting.
-We also quantify the availability of publicly archived data and code, and suggest some useful tools for increasing the reproducibility of colour trait research more broadly.
+Our aim was to explore the state of reproducibility in the study of biological coloration, and to outline simple ways in which it may be improved.
+We first outline common methods for studying biological coloration and present guidelines for comprehensive methods reporting.
+We then explore how well some of these important criteria are reported in the literature.
+We also quantify the availability of publicly archived data and code and suggest some useful tools for increasing the reproducibility of colour trait research more broadly.
 
@@
 
-Generations of biologists have endeavoured to explain the mechanisms and functions of animal and plant colouration [@thayer_concealing_1909; @poulton_colours_1890; @wallace_natural_1891], and uncovering best practices in measuring colour has been a great challenge.
-The application of spectrometry (also called spectroradiometry, or spectrophotometry) to questions of biological colouration revolutionized the field [@dyck_determination_1966].
-Now relatively inexpensive and convenient, spectrometry allows for the rapid capture of quantitative data, and has been widely adopted as standard across the field [@andersson_quantifying_2006].
-Digital photography is also increasingly being used to quantify biological colouration [@stevens_using_2007], as high-resolution cameras are inexpensive and allow for the simultaneous, repeatable sampling of multiple colour patches [@mckay_use_2013].
-
-Subtle variation in the application of these common methods can significantly alter the resulting data.
-When using spectrometry, for example, the angle from which spectra are collected [@santos_strong_2007], the calibration of light and dark standards, the distance between light source and surface, and the size, shape, glossiness or iridescence of the specimen will all directly affect the recorded spectra [@andersson_quantifying_2006; @dalrymple_roses_2014; @meadows_quantifying_2011; @schaefer_birds_2008].
-Equally, the main challenge for gathering colour data from photographs is ensuring standardization against nonlinear responses to light intensity, biases towards particular wavebands, and inconsistent lighting [@stevens_using_2007].
-Also, most digital cameras used for the measurement of biological colouration are consumer products designed for the human visual system.
-Thus hardware modifications for ultraviolet and infra-red imaging are often necessary [e.g.
-@stevens_colour_2013].
+Generations of biologists have endeavoured to explain the mechanisms and functions of animal and plant colouration [@endler_comparing_2005; @thayer_concealing_1909; @poulton_colours_1890; @wallace_natural_1891], and uncovering best practices in measuring colour has been a great challenge.
+The direct measurement of reflectance and/or transmittance through spectrometry revolutionized the study of biological coloration [@dyck_determination_1966], and has been widely adopted as the standard [@andersson_quantifying_2006].
+Digital photography is also increasingly being used to quantify colour [@stevens_using_2007], as high-resolution cameras are inexpensive and allow for the simultaneous, rapid sampling of multiple colour patches [@mckay_use_2013].
 
 Expansion in the availability of objective methods for the measurement of colour has been matched by advances in theory and analysis.
-In particular, the development of visual models has enabled researchers to move beyond purely quantitative comparisons of reflectance spectra and adopt more biologically relevant perspectives when defining and testing hypotheses [@chittka_colour_1992; @endler_comparing_2005; @vorobyev_receptor_1998].
-Visual models typically attempt to describe the reception and early-stage processing of colour information as a function of an object’s reflectance, the ambient illumination, and a receiver’s sensory system [e.g.
-@endler_comparing_2005; @vorobyev_receptor_1998].
-Although relatively easy to implement, visual models are built on multiple assumptions about the way in which stimuli are processed that can dramatically shape the results of a given analysis [@lind_avian_2009; @pike_preserving_2012].
-For example, Vorobyev et al.’s [-@vorobyev_receptor_1998; -@vorobyev_colour_2001] widely used receptor-noise model requires researchers to specify the form of receptor quantum catches (i.e.
-raw versus transformed), the use of any colour constancy mechanism (e.g.
-von Kries), as well as the photoreceptor signal-to-noise ratios, their relative densities, and the type of noise predominating (i.e.
-neural and/or receptor noise).
-These data are only available for a small number of species, such as humans and honeybees, and so assumptions and extrapolations are necessary when models are applied to other species.
-Unless the biological rationale underlying every parameter choice is clearly understood, justified, and reported, such analyses may become ineffective, especially when comparing results between studies [@pike_preserving_2012].
+In particular, the development of visual models has enabled researchers to move beyond quantitative comparisons of reflectance spectra and adopt potentially more biologically relevant perspectives when defining and testing hypotheses [@chittka_colour_1992; @endler_comparing_2005; @vorobyev_receptor_1998].
+Visual models typically attempt to describe the reception and early-stage processing of chromatic and achromatic information as a function of an object’s reflectance, the ambient illumination, and a receiver’s sensory system [e.g., @endler_comparing_2005; @vorobyev_receptor_1998].
+Although relatively easy to implement, visual models are built on multiple assumptions about the way in which stimuli are processed that can dramatically shape the results of a given analysis [e.g., @lind_avian_2009; @pike_preserving_2012].
 
-The comprehensive reporting of methods is a simple and crucial step in ensuring research is reproducible.
-Accordingly, we developed a list of essential information about the capture (Table 1) and analysis (Table 2) of colour data that should ideally be reported.
-Analytical techniques are diverse, mathematically complex, and are being developed rapidly [@kelber_spectral_2010].
-Rapid progress means that the need for a deep understanding of common methods can quickly outstrip the working knowledge of the average researcher.
+## Guidelines for methods reporting
+
+Given that methodological variation may shape results in significant and unpredictable ways (Tables 1 & 2, and references therein), the comprehensive reporting of methods is a simple and crucial step in ensuring research is reproducible.
+Accordingly, we developed a list of information about the capture (Table 1) and analysis (Table 2) of colour data that should ideally be reported.
+With regards to the measurement of colour, we focus on the two most frequently used methods; photography and spectrometry.
+Analytical techniques are diverse, mathematically complex, and are being developed rapidly [@thery_insect_2010; @kelber_spectral_2010].
+Such progress means that the need for a deep understanding of common methods can quickly outstrip the working knowledge of the average researcher.
 As a consequence, the subtle complexity of many analytical techniques can be overlooked by empiricists, leading to critical methodological information not being reported.
 Our guidelines for reporting the details of colour analyses (Table 2) thus cover two broad, common methods: colourimetric (or 'spectral') analyses, and visual modelling.
-These tables present the framework for our subsequent analysis of methods reporting in the literature.
+
+While we wish to emphasize that these details are essential to ensuring full reproducibility of data capture and analysis, we recognize that space restrictions in the main text of manuscripts may preclude the incorporation of all these details.
+In such cases, we suggest that details be included in meta-data or in a supplementary file so that the necessary information is available to researchers.
+It is also the case that there is variability in the degree to which each parameter may affect results, and so we have aimed to provide a brief, qualitative outline of the potential effects that variation in each parameter may have on the data (Table 1 & 2).
+These tables are not intended as a guide to the selection of methods, however, for which we refer readers to excellent recent reviews as well as the original publications [@stevens_using_2007; @kelber_animal_2003; @montgomerie_analyzing_2006; @kemp_integrative_2015, and references in Tables 1 & 2].
 
@@
 
-To assess the current state of reproducibility in the field we searched papers from 2013 in 22 leading journals: _American Journal of Botany_, _The American Naturalist_, _Animal Behaviour_, _Behavioral Ecology_, _Behavioral Ecology and Sociobiology_, _Biological Journal of the Linnean Society_, _Biology Letters_, _Current Zoology_, _Ecology_, _Ecology and Evolution_, _Ecology Letters_, _Ethology_, _Evolution_, _Functional Ecology_, _Journal of Ecology_, _The Journal of Evolutionary Biology_, _The Journal of Experimental Biology_, _Naturwissenschaften_, _New Phytologist_, _Oikos_, _PLoS ONE_, _Proceedings of the Royal Society B: Biological Sciences_.
-On each of the journals' homepages, we used the Boolean phrase 'colour\*' or 'color\*' or 'spectra\*' to search the title or abstract.
-To review the 216 papers returned from our search, journals were haphazardly divided up between the authors.
-We excluded review papers, methodological papers, papers quantifying spatial rather than chromatic properties of a colour patch, and studies taking micro-spectrometric measurements of retinal absorbance.
-The final set of `r nrow(coldat)` papers included only those that used either a spectrometer or camera to quantify colouration.
-In order to reduce the risk of observer bias in our assessment, each paper included in the final set was read and reassessed by three authors.
-Any discrepancies between assessor scores were discussed and resolved prior to analysis.
-We then tabulated what details of the methods were reported against our suggested criteria (Tables 1 and 2), and recorded whether data (in either a 'raw' or 'processed' form) and/or any code were publicly available.
-The data along with analysis script has been stored as a github repo (http://dx.doi.org/10.5281/zenodo.13245).
-We have kept the papers used in our dataset anonymous as our aim was to explore the general question of reproducibility in the field and not single out particular studies.
+To determine the current state of reproducibility in the field we assessed a sample of the literature against a set of our criteria (Table 1 and 2), which we expected should be commonly reported based on our background reading.
+We searched papers from 2013 in 22 leading journals: _American Journal of Botany_, _The American Naturalist_, _Animal Behaviour_, _Behavioral Ecology_, _Behavioral Ecology and Sociobiology_, _Biological Journal of the Linnean Society_, _Biology Letters_, _Current Zoology_, _Ecology_, _Ecology and Evolution_, _Ecology Letters_, _Ethology_, _Evolution_, _Functional Ecology_, _Journal of Ecology_, _The Journal of Evolutionary Biology_, _The Journal of Experimental Biology_, _Naturwissenschaften_, _New Phytologist_, _Oikos_, _PLoS ONE_, _Proceedings of the Royal Society B: Biological Sciences_.
+On each of the journals' homepages, we used the Boolean phrase 'colour*' or 'color*' or 'spectra*' to search the title and/or abstract.
+Journals were haphazardly divided up between the authors to review the 216 papers returned from our search.
+On first pass, we excluded review papers, methodological papers, papers quantifying spatial (i.e., pattern) rather than chromatic properties of a colour patch, and studies taking micro-spectrometric measurements of retinal absorbance.
+The final set of 60 papers included only those that used either a spectrometer or camera to quantify coloration.
+In order to reduce the risk of observer bias in our assessment, each paper included in the final set was read and reassessed by two further authors.
+Any discrepancies between assessment scores were discussed by the three authors that had read the article and resolved prior to analysis.
+We also recorded whether data (in either a 'raw' or 'processed' form) and/or any code were publicly available.
+The data along with our analysis script has been stored as a github repository.
+We have kept the papers used in our dataset anonymous as our aim was to explore the general question of reproducibility in the field.
 
@@
 Overall, `r analysisType[1]` papers generated colourimetric variables [i.e.
-'spectral' measures of hue, saturation, and/or brightness; @montgomerie_analyzing_2006] whereas `r analysisType[4]` used visual models, `r analysisType[2]` used both colourimetrics and visual models, and `r analysisType[3]` used other forms of analysis (e.g.
-purely statistical analyses, such as principle component analysis on reflectance spectra).
+'spectral' measures of hue, saturation, and/or brightness; @montgomerie_analyzing_2006] whereas `r analysisType[4]` used visual models, `r analysisType[2]` used both colourimetrics and visual models, and `r analysisType[3]` used other forms of analysis (e.g., purely statistical analyses, such as principle component analysis on reflectance spectra).
 Of the studies quantifying colour using spectral reflectance, `r analysisTypeSpec[1]` used colourimetric measures, `r analysisTypeSpec[2]` used multiple analyses, `r analysisTypeSpec[4]` used visual models alone, and `r analysisTypeSpec[3]` used other analyses.
@@
 Of the `r nrow(coldat)` studies analysed, `r round((data["raw"]/sum(data))*100, digits = 1)`% publicly provided the raw underlying data, `r round((data["proc"]/sum(data))*100, digits = 1)`% provided data in a pre-processed form, and `r round((data["none"]/sum(data))*100, digits = 1)`% of studies did not provide any publicly accessible data.
-The paucity of data being made available post-publication in studies of biological colouration is in line with other fields [@drew_lost_2013; @vines_mandated_2013; @wolkovich_advances_2012], including the broader field of animal behaviour [@caetano_forgotten_2014].
+The paucity of data being made available post-publication in studies of biological coloration is in line with other fields [@drew_lost_2013; @vines_mandated_2013; @wolkovich_advances_2012], including the broader field of animal behaviour [@caetano_forgotten_2014].
 There is evidence, however, that this trend is shifting as funding agencies and publishers increasingly mandate the release of data [@whitlock_data_2010].
+
 The clearest benefit of open access to data is that it allows researchers to build upon previous work for the testing and refinement of new ideas and methods.
-This is particularly relevant in the study of colour traits, where new methods for analysing colour data are frequently developed [e.g.
-@allen_analyzing_2013; @endler_framework_2012; @stoddard_pattern_2014].
-More generally, the provision of open data may also foster collaborations as researchers draw upon existing work [@whitlock_data_2011], and considerably reduce the difficulty of meta-analyses and large scale comparative work [e.g.
-@burd_distinctive_2014; @maia_key_2013].
-Increased data sharing has been shown to positively correlate with citations [@piwowar_sharing_2007], and given the increasing rate of retractions there is also impetus for researchers to maintain credibility through transparency [@steen_retractions_2010; @van_trouble_2011; @wicherts_willingness_2011].
-We thus encourage researchers to publicly archive data in as raw a form as possible (e.g.
-individual reflectance spectra), to maximise its utility.
-Public data repositories such as figshare and dryad allow the permanent archival of data as citable research objects, and the preparation of data for publication is often straightforward [@whitlock_data_2011].
+As discussed above, it is also essential that complete methodological details be provided, either in the manuscript or associated meta-data, to ensure data can be used to their full potential.
+This is particularly relevant in the study of colour, where new analysis methods are frequently developed [e.g., @allen_analyzing_2013; @endler_framework_2012; @stoddard_pattern_2014].
+More generally, the provision of open data is likely to foster collaborations as researchers draw upon existing work [@piwowar_sharing_2007; @whitlock_data_2011], and considerably reduce the difficulty of meta-analyses and large scale comparative work [e.g., @burd_distinctive_2014; @maia_key_2013].
+We thus encourage researchers in our field to publicly archive raw data (e.g., individual reflectance spectra), to maximise its utility.
 The use of modest data embargoes and appropriate licenses can help to ensure that original authors are able to make the best use of data and are subsequently credited, though we acknowledge that tensions over data-sharing exist [discussed in @roche_troubleshooting_2014].
 
-None of the included studies linked to any form of code, which is not surprising given the popularity of graphical-user-interface (GUI) based statistical and colour-analytical software [e.g.
-AVICOL; @gomez_avicol_2006].
-Here we simply note that there are several advantages to moving away from a GUI-based to a programming workflow, in spite of the initial time investment required.
+None of the included studies linked to any form of code.
+This is not surprising given the popularity of graphical-user-interface (GUI) based statistical and colour-analytical software [e.g., AVICOL; @gomez_avicol_2006].
+When such software is freely available, and is combined with the comprehensive reporting of methods, its use represents a valuable approach to conducting reproducible research.
+We also note that there are advantages in adopting a programming workflow, in spite of the initial time investment required.
 Programming languages such as Python [@van_python_1995] and R [@R_program], for example, are free, open-source, and are host to communities of developers that continuously build and implement analytical tools.
-Indeed, tools specifically for the analysis of colour data already exist [e.g.
-the package 'pavo' for R; @maia_pavo_2013].
-Programming languages are flexible by nature, which allows individual researchers to rapidly explore new methods as they are published, and methods developers to implement their analyses to increase the reach and impact of their work.
-A significant advantage of programmatic analyses is that analysis scripts, when properly curated, represent a complete documented history of a study's methods.
-The chain of modification of raw data may thus be preserved, both for researchers revisiting their own work in the future and for other scientists looking to build directly on the results of previous studies.
+Indeed, tools specifically for the analysis of colour data already exist [e.g., the package 'pavo' for R; @maia_pavo_2013].
+While no program can compensate for having a clear understanding of the analysis at hand, programming languages are flexible by nature.
+This allows individual researchers to rapidly explore new methods as they are published, and methods developers to implement their analyses to increase the reach and impact of their work.
+A further advantage of programmatic analyses is that analysis scripts, when properly curated, represent a complete documented history of a study's methods.
+The chain of modification of raw data may thus be preserved, both for researchers revisiting their own work in the future, and for other scientists looking to build directly on the results of previous studies.
 
@@
 
-Our review of the recent literature highlights some significant impediments to reproducibility in the study of biological colouration.
-The simplest step towards reproducible research is through the comprehensive reporting of methods, yet crucial methodological information is often unreported (Fig.
+Our review of the recent literature highlights some impediments to reproducibility in the study of biological colourationcoloration.
+The simplest step towards reproducible research is through the comprehensive reporting of methods, yet key information is often unreported (Fig.
 1).
-In particular, the light source - sample - probe geometry was rarely reported, yet should be standard reporting for all studies using spectrometry.
-Visual models were regularly used to analyse spectral data.
-However, there were many important aspects of the analysis that were seldom reported, including the type of quantum catch used, whether photoreceptor adaptation was modelled and the type of noise used in receptor-noise models (Fig.
-1).
-Photoreceptor densities should also be reported where possible even if using standard avian densities, as all of these parameters have important consequences for downstream analysis and, hence, the reproducibility of colour trait quantification [e.g.
-@kemp_ornamental_2008; @santos_comparison_2007].
-We also recommend the presentation of spectra where possible, as these figures allow for the rapid assessment of the nature and quality of the data independent of downstream processing.
-We found that only `r round(propSpec_fin["prop_refl_fig"], digits = 1)`% of studies using spectral reflectance data had such a figure.
-Although not considered in our analysis, we also recommend that authors explicitly outline the biological justification underlying their choices for data capture and analysis.
-The permanent, public storage of data also plays a pivotal role in reproducibility and transparency, yet public data storage in studies of colour traits is not yet the norm.
-We encourage researchers to consider the substantial benefits of publicly archiving their data [@piwowar_sharing_2007; @vision_open_2010; @whitlock_data_2010].
-
-Our objective in this paper was to highlight some important gaps that can influence the reproducibility of colour research, and to offer guidelines on simple ways in which studies may be leveraged to their full potential.
-We do not wish to deprecate studies that happen to lack some methodological details.
-We hope that our discussion of methods reporting and data archiving will contribute to the growing culture of reproducibility in science, and that researchers will consider these issues when deciding on the best practices in their work.
-It is only through critical appraisal that we can identify areas for improvement, and constructively contribute to the growth of our field.
-Such assessments are particularly important in an age where accountability, transparency and reproducibility are at the forefront of discussions of the state of scientific research [@drew_lost_2013; @peng_reproducible_2011; @wicherts_willingness_2011].
-We believe that part of the issue at hand is simply that researchers seldom have clear guidelines to follow when undertaking their research, and may overlook important aspects of methods and analysis reporting.
-We hope our critical evaluation of recent literature has highlighted a need for more careful descriptions of how colour is measured and analysed.
-We are optimistic that this may help to establish uniform reporting guidelines for publication and increase transparency and reproducibility, which will continue the advance of an exciting era in colour research.
+Here we provide a list of methodological information that we suggest be specified in studies that focus on the spectrometric- or camera-based analysis of colour (Tables 1 & 2).
+In addition to the suggested information, we also recommend that raw spectra be presented where possible, as these figures allow for the rapid assessment of the nature and quality of the data independent of downstream processing.
+We also emphasise the importance of explicitly outlining the biological justification underlying all choices for data capture and analysis (@kemp_integrative_2015).
+Finally, the public storage of data along with detailed meta-data plays a pivotal role in reproducibility, yet public data storage in studies of colour traits is not yet the norm.
+We therefore encourage researchers to consider the substantial benefits of publicly archiving their data [@whitlock_data_2011; @piwowar_sharing_2007].
+Overall, we hope that our guidelines will encourage researchers to think about the reproducibility of their work and the advantages of increased transparency, which will continue the advance of an exciting era in colour research.
 

### 2015-04-15 docs/ms.Rmd
Updated results section to accepted version without breaking the r code. I think.

@@
 Keywords: spectrometry, photography, visual ecology, pigment, structural colour, visual modelling, methods
-
-Article type: Commentary
-
-Word count: 2613
-
-Number of figures: 1
-
-Number of tables: 2
-
-Data archiving: Github repo (?)
 
@@
 In particular, the development of visual models has enabled researchers to move beyond quantitative comparisons of reflectance spectra and adopt potentially more biologically relevant perspectives when defining and testing hypotheses [@chittka_colour_1992; @endler_comparing_2005; @vorobyev_receptor_1998].
-Visual models typically attempt to describe the reception and early-stage processing of chromatic and achromatic information as a function of an object’s reflectance, the ambient illumination, and a receiver’s sensory system [e.g., @endler_comparing_2005; @vorobyev_receptor_1998].
+Visual models typically attempt to describe the reception and early-stage processing of chromatic and achromatic information as a function of an object’s reflectance, the ambient illumination, and a receiver’s sensory system [@endler_comparing_2005; @vorobyev_receptor_1998].
 Although relatively easy to implement, visual models are built on multiple assumptions about the way in which stimuli are processed that can dramatically shape the results of a given analysis [e.g., @lind_avian_2009; @pike_preserving_2012].
@@
 
-_Reproducibility of data collection_
+Our literature survey suggests there is surprising inconsistency and incompleteness in commonly reported methodological details (Fig.
+1).
+Most studies (`r specs+both`) used a spectrometer to measure colour, yet integration times (`r round(propSpec_fin["prop_int_time"], digits = 0)`% reported), and probe-sample geometry (`r round(propSpec_fin["prop_spec_angle"], digits = 1)`%) and distance (`r round(propSpec_fin["prop_spec_dist"], digits = 0)`%) were often not reported.
+Among studies that used photography (`r cam+both`), `r round(propCam_fin["prop_specpix_avg"], digits = 0)`% detailed the number of pixels averaged, though camera models were more frequently reported (`r round(camModel, digits = 0)`%).
+Light sources were detailed in `r round(propSpec_fin["prop_light_source"], digits = 0)`% and `r round(propCam_fin["prop_light_source"], digits = 0)`% of spectrometer- and camera-based studies, respectively.
+With regard to data analysis, of the `r analysisType[1]` studies that calculated colourimeteric variables, `r round(col_def, digits = 0)`% specifically defined their measure of brightness, hue and/or chroma.
+The receptor-noise limited model (Vorobyev et al.
+1998) was commonly used among studies with visual modelling (`r vismod_type[3]` of `r sum(vismod_type)`), though there was considerable variation in the detailing of the type of receptor noise used (`r round(propSpec_fin["prop_vis_mod_noise_type"], digits = 0)`% reported), the type of quantum catch used (`r round(propSpec_fin["prop_vis_mod_qcatch"], digits = 0)`%), or whether photoreceptor adaptation (`r round(propSpec_fin["prop_vis_mod_adapt"], digits = 0)`%) was modelled.
+In contrast, details of the species’ visual system being modelled (`r round(propSpec_fin["prop_vis_mod_sp"], digits = 0)`%), the background used (`r round(propSpec_fin["prop_vis_mod_bkg"], digits = 0)`%), and the modelled illuminant (`r round(propSpec_fin["prop_irrad_type"], digits = 0)`%) were more commonly reported.
 
-Overall, `r specs+both` studies used a spectrometer to measure colour, while `r cam+both` used a camera.
-Of these studies, `r both` quantified colour using both a spectrometer and a camera.
-Reporting of the make and model of the spectrometer and camera was high (Figure 1; spectrometers = `r round(specModel, digits = 1)`% and cameras = `r round(camModel, digits = 1)`%), however, only `r round(propCam_fin["prop_light_source"], digits = 1)`%  and `r round(propSpec_fin["prop_light_source"], digits = 1)`% of camera and spectrometer studies reported a light source.
-The number of reflectance spectra and pixels averaged were only reported in `r round(propSpec_fin["prop_specpix_avg"], digits = 1)`% and `r round(propCam_fin["prop_specpix_avg"], digits = 1)`% of studies, respectively.
-Surprisingly, dark standards were only reported in `r round(propSpec_fin["prop_dark_std"], digits = 1)`% of the spectrometer studies, while white standards were reported in `r round(propSpec_fin["prop_white_stdspec"], digits = 1)`%.
-Studies quantifying spectral reflectance often did not report the integration time of the spectrometer (`r round(propSpec_fin["prop_int_time"], digits = 1)`%), the angle of the optical probe relative to the surface (`r round(propSpec_fin["prop_spec_angle"], digits = 1)`%) or the distance (`r round(propSpec_fin["prop_spec_dist"], digits = 1)`%) the probe was held from the measurement surface.
-
-_Reproducibility of analyses_
-
-The types of analysis varied across studies.
-Overall, `r analysisType[1]` papers generated colourimetric variables [i.e.
-'spectral' measures of hue, saturation, and/or brightness; @montgomerie_analyzing_2006] whereas `r analysisType[4]` used visual models, `r analysisType[2]` used both colourimetrics and visual models, and `r analysisType[3]` used other forms of analysis (e.g., purely statistical analyses, such as principle component analysis on reflectance spectra).
-Of the studies quantifying colour using spectral reflectance, `r analysisTypeSpec[1]` used colourimetric measures, `r analysisTypeSpec[2]` used multiple analyses, `r analysisTypeSpec[4]` used visual models alone, and `r analysisTypeSpec[3]` used other analyses.
-Studies using colourimetric measures specifically defined their measure of brightness, hue and/or chroma in `r round(col_def, digits = 1)`% of the studies.
-
-Most studies using visual models analysed their data using receptor-noise models (`r vismod_type[3]` of `r sum(vismod_type)` studies), while others used the colour hexagon (`r vismod_type[1]` studies), or a tetrahedral colour space (`r vismod_type[4]` studies).
-Unique, case-specific models were used in `r vismod_type[2]` papers.
-Of visual modelling papers, `r round(propSpec_fin["prop_irrad_type"], digits =2)`% reported the type of irradiance when necessary, and `r round(propSpec_fin["prop_vis_mod_sp"], digits = 1)`% mentioned the species being modelled as receiver.
-The background used in visual models was reported `r round(propSpec_fin["prop_vis_mod_bkg"], digits = 1)`% of the time.
-The type of receptor noise modelled was reported in only `r round(propSpec_fin["prop_vis_mod_noise_type"], digits = 1)`% of the studies.
-Explicit mention of the type of quantum catch used (`r round(propSpec_fin["prop_vis_mod_qcatch"], digits = 1)`% of studies), or whether photoreceptor adaptation (`r round(propSpec_fin["prop_vis_mod_adapt"], digits = 1)`%) was modelled, was not common.
-
-While some of the figures reported above seem alarming, it is important to note that around one-third of papers (`r round(ref_stud, digits = 1)`%) made reference to previous work for details on various methods.
-This means that our above figures are slightly liberal given that the referenced works may have comprehensively covered some of these criteria.
-Despite this, two-thirds of the studies did not reference previous work, and were missing important methodological details.
+While some of the figures reported above seem troubling, it is important to note that `r round(ref_stud, digits = 0)`% of papers made reference to previous work for details on some or all methods.
+The referenced works may have comprehensively covered some of these criteria, but were often incomplete as well, or referenced yet another paper.
+To avoid "decay" of methodological detail reporting over successive papers, we suggest reporting all details along with the current manuscript.
+That aside, the remaining `r 100 - round(ref_stud, digits = 0)`% of studies did not reference previous work, and were missing potentially important methodological details.
 
@@
 In addition to the suggested information, we also recommend that raw spectra be presented where possible, as these figures allow for the rapid assessment of the nature and quality of the data independent of downstream processing.
-We also emphasise the importance of explicitly outlining the biological justification underlying all choices for data capture and analysis (@kemp_integrative_2015).
+We also emphasise the importance of explicitly outlining the biological justification underlying all choices for data capture and analysis [@kemp_integrative_2015].
 Finally, the public storage of data along with detailed meta-data plays a pivotal role in reproducibility, yet public data storage in studies of colour traits is not yet the norm.

### 2015-04-16 R/SpecComp.R
Removed the old figure file which we were told to remove based on reviewer comments.

@@ -1,207 +1,0 @@
-library(pavo)
-setwd("ColTest/")
-
-#####FIX GET SPEC####
-
-getspec <- function(where=getwd(), ext='txt', lim=c(300,700), decimal=".", 
-           subdir=FALSE, subdir.names=FALSE)
-{
-
-extension <- paste('.', ext, sep='')
-
-file_names <- list.files(where, pattern=extension, recursive=subdir, include.dirs=subdir)
-files <- paste(where,'/',file_names,sep='')
-
-cat(length(files),' files found; importing spectra\n')
-
-if(subdir.names){
-	file_names <- gsub(extension,'',file_names)}else{
-	file_names <- gsub(extension,'',basename(file_names))
-	 }
-
-if(length(file_names)==0){
-	stop('No files found. Try a different ext')
-	} 
-
-range <- lim[1]:lim[2]
-
-final <- data.frame(matrix(nrow=length(range), ncol=length(file_names)+1))
-final[,1] <- range
-
-# Setting a progress bar
-progbar <- txtProgressBar(min=0, max=length(files), style=2)
-
-for(i in 1:length(files))
-{
-
-raw <- scan(file=files[i], what='', quiet=T, dec=decimal, sep='\n')
-#ToDo we can actually use this raw string to import metadata if we want
-
-# find last line with text
-# correct for spectrasuite files, which have a "End Processed Spectral Data" at the end
-
-start <- grep('[A-Da-dF-Zf-z]',raw)
-isendline <- length(grep('End.*Spectral Data', raw)) > 0
-
-if(isendline)
-  start <- start[-length(start)]
-
-
-start <- max(start)
-
-end <- length(raw) - start
-
-if( isendline > 0 )
-  end <- end - 1
-
-# Avantes has an extra skipped line between header and data. Bad Avantes.
-newavaheader <- length(grep("Wave.*;Sample.*;Dark.*;Reference;Reflectance", raw)) > 0
-
-if(newavaheader)
-  start <- start + 1
-
-# find if columns are separated by semicolon or tab
-issem <- length(grep(';',raw)) > 0
-istab <- length(grep('\t',raw)) > 0
-
-if(issem & istab)
[…]

### 2015-04-16 R/analysis.r
Cleaning and slight reorganization of the analysis file to make indentation more clear, removed unnecessary annotation

@@ -1,126 +1,118 @@
 #-------------------------------------------------------------------------------------------------#
-# Title: Quality of colour reporting 
+# Title: Reproducible research in the study of biological coloration 
 # Authors: Daniel Noble, Tom White, Kate Umbers, James O'Hanlon, Rhiannon Dalrymple, Daniel Zurek
 # Description: Analysis of reporting throughout a series of "colour" journals. Mainly summary 
 #			   stats and graphs of the data. 
 #-------------------------------------------------------------------------------------------------#
 
 # Clear workspace
-rm(list=ls())
+	rm(list=ls())
 
 ## Packages. Coded to simplify multiple package installation and loading.
-packages <- c("car", "ggplot2", "plyr", "knitr", "markdown")
-install.packages(packages)
-sapply(packages, function(x) library(x, character.only = TRUE, verbose = FALSE))
+	packages <- c("car", "ggplot2", "plyr", "knitr", "markdown")
+	install.packages(packages)
+	sapply(packages, function(x) library(x, character.only = TRUE, verbose = FALSE))
 
 ## Load data 
-coldat <- read.csv("data/col_data.csv", header = TRUE)
-str(coldat)
+	coldat <- read.csv("data/col_data.csv", header = TRUE)
+	str(coldat)
 
 #--------------------------------------------- ANALYSIS -----------------------------------------#
 ## How many studies utilized specs (1), cameras (2), both (3)?
-studytype <- table(coldat$stud_type)
+	studytype <- table(coldat$stud_type)
 
-specs <- as.numeric(studytype[1])
-cam   <- as.numeric(studytype[2])
-both  <- as.numeric(studytype[3])
-
-## Of the studies using a spec, what proportion report the make and model?
-specModel <- (sum(coldat$spec_rep, na.rm = TRUE)/(specs+both))*100
+	specs <- as.numeric(studytype[1])
+	cam   <- as.numeric(studytype[2])
+	both  <- as.numeric(studytype[3])
 
[…]

### 2015-04-16 R/analysis.r
removed some more unnecessary annotation

@@ -69,8 +69,6 @@
 ## How did studies fair in their reporting of visual models?
 	vismod_type <- table(DatSpec$vis_mod_type)
 
-# Note that there are a few studies under "multiple" that need to be checked because irrad_type is NA and yet the visual model they report is not given. We should check these.
-
 ## Number of studies referencing other work.
 	ref_stud <- (sum(coldat$prev_pub)/length(coldat$prev_pub))*100
 

### 2015-04-16 docs/ms.Rmd
Fix author affiliations, add highlights & acknowledgements

@@
 
-Thomas E. White^2,7^, Rhiannon L. Dalrymple^1,7^, Daniel W. A. Noble^2,7^, James C. O’Hanlon ^2,7^, Daniel B. Zurek^5,7^, Kate D. L. Umbers^3,4,6,7^
+Thomas E. White^1,7^, Rhiannon L. Dalrymple^2,7^, Daniel W. A. Noble^1,2,7^, James C. O’Hanlon ^1,7^, Daniel B. Zurek^3,7^, Kate D. L. Umbers^4,5,6,7^
 
-^1^ Evolution & Ecology Research Centre, School of Biological, Earth and Environmental Sciences, University of New South Wales, Kensington, NSW, Australia, 2022 ^2^ Department of Biological Sciences, Macquarie University, North Ryde, NSW, Australia, 2109 ^3^ School of Biological Sciences, University of Wollongong, Wollongong, NSW, Austrlaia 2252 ^4^ Centre for Evolutionary Biology, University of Western Australia, Crawley, WA, Australia 6008 ^5^ Department of Biological Sciences, University of Pittsburgh, Pittsburgh, PA, USA, 15260 ^6^ School of Science and Health, University of Western Sydney Hawkesbury, NSW, Australia 2751
+^1^ Department of Biological Sciences, Macquarie University, North Ryde, NSW, Australia, 2109 ^2^ Evolution & Ecology Research Centre, School of Biological, Earth and Environmental Sciences, University of New South Wales, Kensington, NSW, Australia, 2022 ^3^ Department of Biological Sciences, University of Pittsburgh, Pittsburgh, PA, USA, 15260 ^4^ School of Biological Sciences, University of Wollongong, Wollongong, NSW, Austrlaia 2252 ^5^ Centre for Evolutionary Biology, University of Western Australia, Crawley, WA, Australia 6008 ^6^ School of Science and Health, University of Western Sydney Hawkesbury, NSW, Australia 2751
 
@@
 
-Corresponding Author:
+**Corresponding Author:**
 
@@
 
-Phone: +61 2 9850 6279
+**Phone:** +61 2 9850 6279
 
-E-mail: thomas.white@mq.edu.au
+**E-mail:** thomas.white@mq.edu.au
 
-Keywords: spectrometry, photography, visual ecology, pigment, structural colour, visual modelling, methods
+**Keywords:** methods, photography, pigment, spectrometry, structural colour, visual ecology, visual modelling
+
+**Highlights:**
+
+- We review the reproducibility of contemporary studies of biological coloration.
+- Methodological details are often missing, and data are seldom publicly available.
+- We present guidelines for methods reporting, and for improving reproducibility.
 
@@
 
+## Acknowledgements
+We thank Rafael Maia, the Morehouse lab, and three anonymous reviewers for valuable discussions and suggestions.
+TEW and RLD were supported by Australian Postgraduate Awards, and DWAN was supported by an NSERC PGSD and MQRES scholarship.
+
 ## References

### 2015-04-16 README.Rmd
Updated README files and also created .Rmd to intergaret code for better presentation

# Reproducible research in the study of biological coloration

Thomas E. White^1,7^, Rhiannon L. Dalrymple^2,7^, Daniel W. A. Noble^1,2,7^, James C. O’Hanlon ^1,7^, Daniel B. Zurek^3,7^, Kate D. L. Umbers^4,5,6,7^

^1^ Department of Biological Sciences, Macquarie University, North Ryde, NSW, Australia, 2109 ^2^ Evolution & Ecology Research Centre, School of Biological, Earth and Environmental Sciences, University of New South Wales, Kensington, NSW, Australia, 2022 ^3^ Department of Biological Sciences, University of Pittsburgh, Pittsburgh, PA, USA, 15260 ^4^ School of Biological Sciences, University of Wollongong, Wollongong, NSW, Austrlaia 2252 ^5^ Centre for Evolutionary Biology, University of Western Australia, Crawley, WA, Australia 6008 ^6^ School of Science and Health, University of Western Sydney Hawkesbury, NSW, Australia 2751

### Introduction
This repository contains data collected from the literature to assess to quality of reporting in signal-based colour studies of animal and plant taxa. Our literature search includes studies collected in 2013 from research journals likley to do colour signal research. We also detail important methodological deatils that ideally should be reported in studies that quantify colour using both spectrophotometry and camera equipment along with important analytical parameters used for visual modelling.

### Executing code and compiling the manuscript

To compile ms, knit to ms.md file. First you need to set working directory to the respository

```{r, eval = FALSE, echo = TRUE}
# Set wd to col_sci repo
setwd("~/col_sci/")

# Source the knit_doc.R
source("R/knit_doc.R")

```

This command will generate a markdown file (.md) which can then be used to compile the manuscript in .pdf or .dox using pandoc. In terminal navigate to the "docs/" folder and run:

```{r, eval = FALSE, echo = TRUE}
pandoc ms.md -o ms.pdf --bibliography references.bib --csl ref_style.csl -H ms.sty
```

Note that you will need pandoc installed on your computer (http://pandoc.org).

If you have any troubles please don't hesitate to contact the authors: Daniel Noble and Tom White for more information.

### 2015-04-16 knit_readme.r
Adding the knit file for README

@@ -1,0 +1,4 @@
+
+#Convert README to md to evaluate and integrate code.
+
+knit("READme.Rmd")

### 2015-04-16 knit_readme.r
Minor change to readme

@@ -1,4 +1,4 @@
 
 #Convert README to md to evaluate and integrate code.
 
-knit("READme.Rmd")
+knit("README.Rmd")

### 2015-04-16 knit_readme.r
Actually don't need knit doc. Compiles .Rmd nicely

@@ -1,4 +1,0 @@
-
-#Convert README to md to evaluate and integrate code.
-
-knit("README.Rmd")

### 2015-04-16 README.Rmd
Minor changes to readme

@@
 
-Thomas E. White^1,7^, Rhiannon L. Dalrymple^2,7^, Daniel W. A. Noble^1,2,7^, James C. O’Hanlon ^1,7^, Daniel B. Zurek^3,7^, Kate D. L. Umbers^4,5,6,7^
+Thomas E. White1, Rhiannon L. Dalrymple2, Daniel W. A. Noble1,2, James C. O’Hanlon 1, Daniel B. Zurek3, Kate D. L. Umbers4,5,6
 
-^1^ Department of Biological Sciences, Macquarie University, North Ryde, NSW, Australia, 2109 ^2^ Evolution & Ecology Research Centre, School of Biological, Earth and Environmental Sciences, University of New South Wales, Kensington, NSW, Australia, 2022 ^3^ Department of Biological Sciences, University of Pittsburgh, Pittsburgh, PA, USA, 15260 ^4^ School of Biological Sciences, University of Wollongong, Wollongong, NSW, Austrlaia 2252 ^5^ Centre for Evolutionary Biology, University of Western Australia, Crawley, WA, Australia 6008 ^6^ School of Science and Health, University of Western Sydney Hawkesbury, NSW, Australia 2751
+1 Department of Biological Sciences, Macquarie University, North Ryde, NSW, Australia, 2109 2 Evolution & Ecology Research Centre, School of Biological, Earth and Environmental Sciences, University of New South Wales, Kensington, NSW, Australia, 2022 3 Department of Biological Sciences, University of Pittsburgh, Pittsburgh, PA, USA, 15260 4 School of Biological Sciences, University of Wollongong, Wollongong, NSW, Austrlaia 2252 5 Centre for Evolutionary Biology, University of Western Australia, Crawley, WA, Australia 6008 6 School of Science and Health, University of Western Sydney Hawkesbury, NSW, Australia 2751
 
@@
 source("R/knit_doc.R")
-
 ```

### 2015-04-16 README.Rmd
Testing superscript

@@
 
-Thomas E. White1, Rhiannon L. Dalrymple2, Daniel W. A. Noble1,2, James C. O’Hanlon 1, Daniel B. Zurek3, Kate D. L. Umbers4,5,6
+Thomas E. White<sup>1</sup>, Rhiannon L. Dalrymple2, Daniel W. A. Noble1,2, James C. O’Hanlon 1, Daniel B. Zurek3, Kate D. L. Umbers4,5,6
 
@@
 Our literature search includes studies collected in 2013 from research journals likley to do colour signal research.
-We also detail important methodological deatils that ideally should be reported in studies that quantify colour using both spectrophotometry and camera equipment along with important analytical parameters used for visual modelling.
+We also detail important methodological details that ideally should be reported in studies that quantify colour using both spectrophotometry and camera equipment along with important analytical parameters used for visual modelling.
 

### 2015-04-16 README.Rmd
Added superscripts

@@
 
-Thomas E. White<sup>1</sup>, Rhiannon L. Dalrymple2, Daniel W. A. Noble1,2, James C. O’Hanlon 1, Daniel B. Zurek3, Kate D. L. Umbers4,5,6
+Thomas E. White<sup>1</sup>, Rhiannon L. Dalrymple<sup>2</sup>, Daniel W. A. Noble<sup>1,2</sup>, James C. O’Hanlon<sup>1</sup>, Daniel B. Zurek<sup>3</sup>, Kate D. L. Umbers<sup>4,5,6</sup>
 
-1 Department of Biological Sciences, Macquarie University, North Ryde, NSW, Australia, 2109 2 Evolution & Ecology Research Centre, School of Biological, Earth and Environmental Sciences, University of New South Wales, Kensington, NSW, Australia, 2022 3 Department of Biological Sciences, University of Pittsburgh, Pittsburgh, PA, USA, 15260 4 School of Biological Sciences, University of Wollongong, Wollongong, NSW, Austrlaia 2252 5 Centre for Evolutionary Biology, University of Western Australia, Crawley, WA, Australia 6008 6 School of Science and Health, University of Western Sydney Hawkesbury, NSW, Australia 2751
+<sup>1</sup> Department of Biological Sciences, Macquarie University, North Ryde, NSW, Australia, 2109 <sup>2</sup> Evolution & Ecology Research Centre, School of Biological, Earth and Environmental Sciences, University of New South Wales, Kensington, NSW, Australia, 2022 <sup>3</sup> Department of Biological Sciences, University of Pittsburgh, Pittsburgh, PA, USA, 15260 <sup>4</sup> School of Biological Sciences, University of Wollongong, Wollongong, NSW, Austrlaia 2252 <sup>5</sup> Centre for Evolutionary Biology, University of Western Australia, Crawley, WA, Australia 6008 <sup>6</sup> School of Science and Health, University of Western Sydney Hawkesbury, NSW, Australia 2751
 

### 2015-04-16 R/analysis.r
Corrected figure 1 axis to change angle and integration time numbers. Minor changes throughout to correct spelling and remove & to keep consistent

@@ -86,7 +86,7 @@
 
 		dat     <- t(arrange(data.frame(props, N_props), props, decreasing = TRUE))
 
-		names   <- c("Spec", "Cam", "SpecL", "PixAvg"," CamL"," WtSd", "SpecAvg",  "IntT", "Angle", "Dist", "DrkSd")
+		names   <- c("Spec", "Cam", "SpecL", "PixAvg"," CamL"," WtSd", "SpecAvg",  "Angle", "IntT", "Dist", "DrkSd")
 		colnames(dat) <- names
 
 		barplot(dat[1,], ylim = c(0,100), ylab = "Percentage of studies reporting criteria", xlab = "", col = "gray", space = 0.30, cex.names = 0.72, mgp = c(2.5,0.5,0), cex.axis = 1.2, cex.lab = 1.5) -> bp.out

### 2015-04-16 docs/ms.Rmd
Corrected figure 1 axis to change angle and integration time numbers. Minor changes throughout to correct spelling and remove & to keep consistent

@@
 
-Given that methodological variation may shape results in significant and unpredictable ways (Tables 1 & 2, and references therein), the comprehensive reporting of methods is a simple and crucial step in ensuring research is reproducible.
+Given that methodological variation may shape results in significant and unpredictable ways (Tables 1 and 2, and references therein), the comprehensive reporting of methods is a simple and crucial step in ensuring research is reproducible.
 Accordingly, we developed a list of information about the capture (Table 1) and analysis (Table 2) of colour data that should ideally be reported.
@@
 In such cases, we suggest that details be included in meta-data or in a supplementary file so that the necessary information is available to researchers.
-It is also the case that there is variability in the degree to which each parameter may affect results, and so we have aimed to provide a brief, qualitative outline of the potential effects that variation in each parameter may have on the data (Table 1 & 2).
-These tables are not intended as a guide to the selection of methods, however, for which we refer readers to excellent recent reviews as well as the original publications [@stevens_using_2007; @kelber_animal_2003; @montgomerie_analyzing_2006; @kemp_integrative_2015, and references in Tables 1 & 2].
+It is also the case that there is variability in the degree to which each parameter may affect results, and so we have aimed to provide a brief, qualitative outline of the potential effects that variation in each parameter may have on the data (Table 1 and 2).
+These tables are not intended as a guide to the selection of methods, however, for which we refer readers to excellent recent reviews as well as the original publications [@stevens_using_2007; @kelber_animal_2003; @montgomerie_analyzing_2006; @kemp_integrative_2015, and references in Tables 1 and 2].
 
@@
 1).
-Most studies (`r specs+both`) used a spectrometer to measure colour, yet integration times (`r round(propSpec_fin["prop_int_time"], digits = 0)`% reported), and probe-sample geometry (`r round(propSpec_fin["prop_spec_angle"], digits = 1)`%) and distance (`r round(propSpec_fin["prop_spec_dist"], digits = 0)`%) were often not reported.
-Among studies that used photography (`r cam+both`), `r round(propCam_fin["prop_specpix_avg"], digits = 0)`% detailed the number of pixels averaged, though camera models were more frequently reported (`r round(camModel, digits = 0)`%).
+Most studies (n = `r specs+both`) used a spectrometer to measure colour, yet integration times (`r round(propSpec_fin["prop_int_time"], digits = 0)`%), and probe-sample geometry (`r round(propSpec_fin["prop_spec_angle"], digits = 0)`%) and distance (`r round(propSpec_fin["prop_spec_dist"], digits = 0)`%) were often not reported.
+Among studies that used photography (n = `r cam+both`), `r round(propCam_fin["prop_specpix_avg"], digits = 0)`% detailed the number of pixels averaged, though camera models were more frequently reported (`r round(camModel, digits = 0)`%).
 Light sources were detailed in `r round(propSpec_fin["prop_light_source"], digits = 0)`% and `r round(propCam_fin["prop_light_source"], digits = 0)`% of spectrometer- and camera-based studies, respectively.
@@
 
-Our review of the recent literature highlights some impediments to reproducibility in the study of biological colourationcoloration.
+Our review of the recent literature highlights some impediments to reproducibility in the study of biological coloration.
 The simplest step towards reproducible research is through the comprehensive reporting of methods, yet key information is often unreported (Fig.
 1).
-Here we provide a list of methodological information that we suggest be specified in studies that focus on the spectrometric- or camera-based analysis of colour (Tables 1 & 2).
+Here we provide a list of methodological information that we suggest be specified in studies that focus on the spectrometric- or camera-based analysis of colour (Tables 1 and 2).
 In addition to the suggested information, we also recommend that raw spectra be presented where possible, as these figures allow for the rapid assessment of the nature and quality of the data independent of downstream processing.

### 2015-04-16 docs/ms.Rmd
Made a minor correction on calculation of teh number of studies using colourmetric variables. Excluded some in prior analysis because studies categorized as multiple were not included in the calculation. Fixed

@@
 Light sources were detailed in `r round(propSpec_fin["prop_light_source"], digits = 0)`% and `r round(propCam_fin["prop_light_source"], digits = 0)`% of spectrometer- and camera-based studies, respectively.
-With regard to data analysis, of the `r analysisType[1]` studies that calculated colourimeteric variables, `r round(col_def, digits = 0)`% specifically defined their measure of brightness, hue and/or chroma.
+With regard to data analysis, of the `r colorMet` studies that calculated colourimeteric variables, `r round(col_def, digits = 0)`% specifically defined their measure of brightness, hue and/or chroma.
 The receptor-noise limited model (Vorobyev et al.

### 2015-04-16 R/analysis.r
Added code to calculate total number of studies using colourmetric variables

@@ -60,11 +60,14 @@
 ## How do studies analyse colour?
 	analysisType <- table(coldat$analysis_type)
 
-## How about just spec studies because they can do visual modelling, or colorimetric
-	analysisTypeSpec <- table(DatSpec$analysis_type)
+## How many colourmetric studies
+	colorMet <- length(coldat$colmetric_def[!is.na(coldat$colmetric_def)])
 
 ## Of colormetric studies how many defined their colormetrics?
 	col_def <- (sum(coldat$colmetric_def, na.rm = TRUE)/(length(coldat$colmetric_def[!is.na(coldat$colmetric_def)])))*100
+
+## How about just spec studies because they can do visual modelling, or colorimetric
+	analysisTypeSpec <- table(DatSpec$analysis_type)
 
 ## How did studies fair in their reporting of visual models?
 	vismod_type <- table(DatSpec$vis_mod_type)

### 2015-04-16 R/knit_doc.r
Added source code to run analysis before kniting doc

@@ -1,5 +1,10 @@
 #Read necessary functions and files. There are two functions knit2pdf and knit2html which will produce documents directly, but there are problems with pdf and also problems rendering the markdown file when complete. Best to just use knit first then take md file and convert in ternimal. 
 
-setwd("./docs")
+# First source analysis file 
+source("R/analysis.R")
+
+# knit the .RmD file with R code
+
+setwd("./docs/")
 knit("ms.Rmd")
 setwd("..")

### 2015-04-17 docs/ms.Rmd
Fix dud refs in .Rmd & update legends

@@
 The study of colour in nature has generated insights into fundamental evolutionary and ecological processes, and research into colour traits is a rapidly growing field [@kelber_spectral_2010].
-The ongoing interest in biological coloration has in part been driven by the increased availability of key technologies, including spectrometry and photography, and concurrent advances in methods for analysing colour data, such as visual models [e.g., @endler_comparing_2005; @stevens_using_2007; @kelber_animal_2003].
+The ongoing interest in biological coloration has in part been driven by the increased availability of key technologies, including spectrometry and photography, and concurrent advances in methods for analysing colour data, such as visual models [e.g., @kelber_animal_2003; @endler_comparing_2005; @stevens_using_2007].
 While these developments are positive for the field, the increasingly complex analyses being run on ever-greater amounts of data heighten the need for comprehensive methods reporting and diligent data management [@alsheikh_public_2011; @nekrutenko_next_2012].
@@
 As discussed above, it is also essential that complete methodological details be provided, either in the manuscript or associated meta-data, to ensure data can be used to their full potential.
-This is particularly relevant in the study of colour, where new analysis methods are frequently developed [e.g., @allen_analyzing_2013; @endler_framework_2012; @stoddard_pattern_2014].
-More generally, the provision of open data is likely to foster collaborations as researchers draw upon existing work [@piwowar_sharing_2007; @whitlock_data_2011], and considerably reduce the difficulty of meta-analyses and large scale comparative work [e.g., @burd_distinctive_2014; @maia_key_2013].
+This is particularly relevant in the study of colour, where new analysis methods are frequently developed [e.g., @endler_framework_2012; @allen_analyzing_2013; @stoddard_pattern_2014].
+More generally, the provision of open data is likely to foster collaborations as researchers draw upon existing work [@piwowar_sharing_2007; @whitlock_data_2011], and considerably reduce the difficulty of meta-analyses and large scale comparative work [e.g., @maia_key_2013; @burd_distinctive_2014].
 We thus encourage researchers in our field to publicly archive raw data (e.g., individual reflectance spectra), to maximise its utility.

### 2015-04-17 R/analysis.r
Fixed up figure to meet criteria of editor. y-label changes, cut excess line on x axis and put doubel brakets and letters within graphs

@@ -92,12 +92,11 @@
 		names   <- c("Spec", "Cam", "SpecL", "PixAvg"," CamL"," WtSd", "SpecAvg",  "Angle", "IntT", "Dist", "DrkSd")
 		colnames(dat) <- names
 
-		barplot(dat[1,], ylim = c(0,100), ylab = "Percentage of studies reporting criteria", xlab = "", col = "gray", space = 0.30, cex.names = 0.72, mgp = c(2.5,0.5,0), cex.axis = 1.2, cex.lab = 1.5) -> bp.out
+		barplot(dat[1,], ylim = c(0,120), ylab = "Percentage of studies reporting criteria", xlab = "", col = "gray", space = 0.30, cex.names = 0.72, mgp = c(2.5,0.5,0), cex.axis = 1.2, cex.lab = 1.5, las = 2) -> bp.out
 			abline(h = 0, lwd = 2)
-			text(dat[2,], x = bp.out, y = 98)
+			text(dat[2,], x = bp.out, y = 100)
 			text(paste(round(dat[1,], digits = 0), "%", sep = ""), x = bp.out, y = round(dat[1,], digits = 0)*0.5)
-			mtext("a)", adj = -0.13, padj = -1, cex = 2)
-
+			text(x = bp.out[length(bp.out)], y = 114, "(a)", cex = 2)
 
 		props_analy   <- propSpec_fin[c("prop_irrad_type","prop_vis_mod_sp", "prop_vis_mod_adapt", "prop_vis_mod_qcatch", "prop_vis_mod_bkg", "prop_vis_mod_noise_type")]
 
@@ -108,12 +107,13 @@
 		names_analy   <- c("SpRep", "Bkg", "Irrad", "Qcatch","Noise"," ChrAdapt")
 		colnames(dat_analy) <- names_analy
 
-		barplot(dat_analy[1,], ylim = c(0,100), ylab = "", xlab = "", col = "gray", xlim = c(0,10), cex.names = 0.72, mgp = c(2.5,0.5,0), cex.axis = 1.2, cex.lab = 1.5) -> bp.out
-			abline(h = 0, lwd = 2)
-			text(dat_analy[2,], x = bp.out, y = 98.7)
[…]

### 2015-04-20 README.Rmd
Add final doi for repository
Updated the readme file with a better explanation on how to compile the manuscript and fixes a spelling mistake.

@@
 # Reproducible research in the study of biological coloration
+
+[![DOI](https://zenodo.org/badge/doi/10.5281/zenodo.16949.svg)](http://dx.doi.org/10.5281/zenodo.16949)
 
@@
 
-<sup>1</sup> Department of Biological Sciences, Macquarie University, North Ryde, NSW, Australia, 2109 <sup>2</sup> Evolution & Ecology Research Centre, School of Biological, Earth and Environmental Sciences, University of New South Wales, Kensington, NSW, Australia, 2022 <sup>3</sup> Department of Biological Sciences, University of Pittsburgh, Pittsburgh, PA, USA, 15260 <sup>4</sup> School of Biological Sciences, University of Wollongong, Wollongong, NSW, Austrlaia 2252 <sup>5</sup> Centre for Evolutionary Biology, University of Western Australia, Crawley, WA, Australia 6008 <sup>6</sup> School of Science and Health, University of Western Sydney Hawkesbury, NSW, Australia 2751
+<sup>1</sup> Department of Biological Sciences, Macquarie University, North Ryde, NSW, Australia, 2109 <sup>2</sup> Evolution & Ecology Research Centre, School of Biological, Earth and Environmental Sciences, University of New South Wales, Kensington, NSW, Australia, 2022 <sup>3</sup> Department of Biological Sciences, University of Pittsburgh, Pittsburgh, PA, USA, 15260 <sup>4</sup> School of Biological Sciences, University of Wollongong, Wollongong, NSW, Australia 2252 <sup>5</sup> Centre for Evolutionary Biology, University of Western Australia, Crawley, WA, Australia 6008 <sup>6</sup> School of Science and Health, University of Western Sydney Hawkesbury, NSW, Australia 2751
 
@@
 
-To compile ms, knit to ms.md file. First you need to set working directory to the respository
+To compile the manuscript use 'knitr' to knit the ms.Rmd file to a markdown, ms.md file.
+First you need to set your working directory to this respository and then you can create the markdown file by sourcing the knit_doc.R file in the R folder as follows:
 
@@
 # Set wd to col_sci repo
-setwd("~/col_sci/")
+setwd("[INSERT PATH]/colsci_rep/")
 
@@
 
-This command will generate a markdown file (.md) which can then be used to compile the manuscript in .pdf or .dox using pandoc. In terminal navigate to the "docs/" folder and run:
+This command will generate a markdown file (ms.md) which can then be used to compile the manuscript in .pdf or .dox using pandoc.
+In terminal navigate to the 'docs/' folder in the colsci_rep repository and run:
 
@@
 
-Note that you will need pandoc installed on your computer (http://pandoc.org).
+Note that you will need pandoc and LaTex installed on your computer (http://pandoc.org).
 

### 2015-06-05 README.Rmd
Minor update to readme to include a slightly better explanation. In particular about runnning commands in R
In press

@@
 
-Thomas E. White<sup>1</sup>, Rhiannon L. Dalrymple<sup>2</sup>, Daniel W. A. Noble<sup>1,2</sup>, James C. O’Hanlon<sup>1</sup>, Daniel B. Zurek<sup>3</sup>, Kate D. L. Umbers<sup>4,5,6</sup>
+This paper is now in press, and can be cited as:
 
-<sup>1</sup> Department of Biological Sciences, Macquarie University, North Ryde, NSW, Australia, 2109 <sup>2</sup> Evolution & Ecology Research Centre, School of Biological, Earth and Environmental Sciences, University of New South Wales, Kensington, NSW, Australia, 2022 <sup>3</sup> Department of Biological Sciences, University of Pittsburgh, Pittsburgh, PA, USA, 15260 <sup>4</sup> School of Biological Sciences, University of Wollongong, Wollongong, NSW, Australia 2252 <sup>5</sup> Centre for Evolutionary Biology, University of Western Australia, Crawley, WA, Australia 6008 <sup>6</sup> School of Science and Health, University of Western Sydney Hawkesbury, NSW, Australia 2751
+(White TE, Dalrymple RL, Noble DWA, O’Hanlon JC, Zurek DB, Umbers KDL (2015) Reproducible research in the study of biological colouration.
+Animal Behaviour, in press)[http://dx.doi.org/10.1016/j.anbehav.2015.05.007]
 
@@
 To compile the manuscript use 'knitr' to knit the ms.Rmd file to a markdown, ms.md file.
+This can all be done in the R console.
 First you need to set your working directory to this respository and then you can create the markdown file by sourcing the knit_doc.R file in the R folder as follows:

### 2015-06-05 README.Rmd
Fixed in press link

@@
 
-(White TE, Dalrymple RL, Noble DWA, O’Hanlon JC, Zurek DB, Umbers KDL (2015) Reproducible research in the study of biological colouration.
-Animal Behaviour, in press)[http://dx.doi.org/10.1016/j.anbehav.2015.05.007]
+[White TE, Dalrymple RL, Noble DWA, O’Hanlon JC, Zurek DB, Umbers KDL (2015) Reproducible research in the study of biological colouration.
+*Animal Behaviour*, in press](http://dx.doi.org/10.1016/j.anbehav.2015.05.007)
 

### 2016-08-04 README.Rmd
Published
Updated Zenodo badge to final release

@@
 
-[![DOI](https://zenodo.org/badge/doi/10.5281/zenodo.16949.svg)](http://dx.doi.org/10.5281/zenodo.16949)
+[![DOI](https://zenodo.org/badge/7271/daniel1noble/colsci_rep.svg)](https://zenodo.org/badge/latestdoi/7271/daniel1noble/colsci_rep)
 
-This paper is now in press, and can be cited as:
+This paper is now published, and can be cited as:
 
 [White TE, Dalrymple RL, Noble DWA, O’Hanlon JC, Zurek DB, Umbers KDL (2015) Reproducible research in the study of biological colouration.
-*Animal Behaviour*, in press](http://dx.doi.org/10.1016/j.anbehav.2015.05.007)
+*Animal Behaviour*, 106, 51-57](http://dx.doi.org/10.1016/j.anbehav.2015.05.007)
 
